### 2017-07-20 R/SubGenome Analysis.Rmd
Init Project

---
title: "SubGenome Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-07-20 R/parseForSubgenomes.R

@@ -1,0 +1,31 @@
+library(readr)
+syntelogs.import <- read_delim("~/Dropbox/SynMap Subgenomes/sorghum vs maize.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]
+
+
+library(plyr)
+## Find which sets of chromosomes with syntelogs should be in group 1 or group 2
+syntelogs.aggregate <- ddply(syntelogs.complete, ~block+org_chr1+org_chr2, summarise, median=median(ks), mean=mean(ks), count=length(ks))
+syntelogs.orthologs <- subset(syntelogs.aggregate, count>=12 & median<=0.4234)
+# syntelogs.orthologs[order(syntelogs.orthologs["org_chr1"], syntelogs.orthologs["org_chr2"]),]#view sorted orthologs
+orthologs.counts <- ddply(syntelogs.orthologs, ~org_chr1+org_chr2, summarise, total_genes=sum(count))
+orthologs.sorted <- orthologs.counts[order(orthologs.counts["org_chr1"], -orthologs.counts["total_genes"]),]
+
+# orthologs.sorted[match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1),] #first instance of each group (i.e. biggest=sub1)
+orthologs.sub1rows <- match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1) #not exactly right!!!! how to split sub 1 over multiple chromosomes?
+orthologs.sub1 <- orthologs.sorted[orthologs.sub1rows,]
+orthologs.sub2 <- orthologs.sorted[-orthologs.sub1rows,]
+orthologs.sub1$subGenome="sub1"
+orthologs.sub2$subGenome="sub2"
+
+
+## Now recover which genes are in these sets
+subset(syntelogs.orthologs, count>=12 & median<=0.4234 & org_chr1=="a31607_1" & org_chr2=="b34889_1")
+# orthologs.blocks_groups <- syntelogs.orthologs[syntelogs.orthologs$count>=12 & syntelogs.orthologs$median<=0.4234,]
+subgenome.groups <- rbind(merge(syntelogs.orthologs, orthologs.sub1, all.x = F), merge(syntelogs.orthologs, orthologs.sub2, all.x = F))
+subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
[…]

### 2017-07-24 R/SubGenome Analysis.Rmd
Moved processing from R file to Rmd file.
Moved processing from R file to Rmd file in order to tie the comments related to logic and processing steps into the analysis.  Provides a better visual story and record for reproducability.

@@
 ## R Markdown
+Some preprocessing was needed to generate the input files for this project.
+We begin with SynMap output based on the Schnable 2011 paper (see https://www.evernote.com/shard/s303/nl/52987123/1ce74f11-98a0-4854-b363-3172e79eeae3 for details).
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+This resulted in the file (31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks).
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+We parse this file to a flat format using a perl script (parseKsKnFile.pl)
 
-```{r cars}
-summary(cars)
+(Do this in command line until I figure out how to run it from R)
+
+> perl parseKsKnFile.pl ../Data/31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum\ vs\ maize.tab
+
+```{r parseSynMap, include=FALSE}
+cmd <- paste("perl", "parseKsKnFile.pl", "31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks")
+system(cmd)
 ```
 
-## Including Plots
+The data is now prepared for import and cleaning.  We will import the file and remove any incomplete lines.
 
-You can also embed plots, for example:
+```{r importSynMapData, echo=FALSE}
+library(readr)
+library(plyr)
+library(reshape2)
+library(ggplot2)
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+## Import data and remove incomplete rows
+syntelogs.import <- read_delim("../Data/sorghum vs maize.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]
+
+sprintf("We have imported %d lines from the parsed SynMap file and removed %d incomplete rows resulting in %d syntenic genes.", nrow(syntelogs.import), nrow(syntelogs.import)- nrow(syntelogs.complete), nrow(syntelogs.complete))
 ```
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+We need to select a cutoff ks value to differentiate between the most recent alpha duplication event and the previous beta duplication.
+
+```{r pickKSCutoff, echo=FALSE}
+## Pick a ks value which separates alpha duplication from beta duplication (between the first two peaks)
+log10_ks_cutoff <- -0.575
+qplot(log10(syntelogs.complete$ks), geom="histogram", binwidth=.01)  + geom_vline(mapping = NULL, data = NULL, xintercept = log10_ks_cutoff, na.rm = FALSE, show.legend = NA)
+ks_cutoff <- 10^log10_ks_cutoff
+#ks_cutoff <- 0.4234 #include most of the first 2 peaks
+```
+
+In order to separate subgenome 1 from subgenome 2, we need to find syntenic blocks of 12 or more with a median ks value below the cutoff to determine syntelogs which are related to the alpha duplication event.
+For each chromosome with duplicate genes, the chromosome that has the largest syntenic block is the dominant subgenome, and the lesser sized blocks on other chromosomes are the recessive subgenome.
+
+(but how to group subgenomes over multiple chromosomes?)
+
+```{r findSubgenomes, echo=FALSE}
+## Find which sets of chromosomes with syntelogs should be in group 1 or group 2, where group 1 has larger syntenic blocks
+# First group rows with same syntenic block and chromosome, then summarize each block's ks values with median/mean/count
+syntelogs.aggregate <- ddply(syntelogs.complete, ~block+org_chr1+org_chr2, summarise, median=median(ks), mean=mean(ks), count=length(ks))
+
+# Following the schnable article, syntenic blocks must have 12 genes "The median synonymous substitution rate of all gene pairs in a syntenic block between maize and sorghum can be used to classify syntenic blocks of 12 or more genes unambiguously as orthologous or homoeologous, however" and have a median ks value that discriminates for the alpha duplication event.
+syntelogs.orthologs <- subset(syntelogs.aggregate, count>=12 & median<=ks_cutoff)
+
+# Determine which chromosome has largest syntenic block, the syntelogs on this chromosome are subgenome 1
+orthologs.counts <- ddply(syntelogs.orthologs, ~org_chr1+org_chr2, summarise, total_genes=sum(count))
+orthologs.sorted <- orthologs.counts[order(orthologs.counts["org_chr1"], -orthologs.counts["total_genes"]),]
+
+# orthologs.sorted[match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1),] #first instance of each group (i.e. biggest=sub1)
+orthologs.sub1rows <- match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1) #not exactly right!!!! how to split sub 1 over multiple chromosomes?
+orthologs.sub1 <- orthologs.sorted[orthologs.sub1rows,]
+orthologs.sub2 <- orthologs.sorted[-orthologs.sub1rows,]
+orthologs.sub1$subGenome="sub1"
+orthologs.sub2$subGenome="sub2"
+```
+
+Now that we have recovered the dominant and recessive subgenomes, lets write this to a file.
+
+```{r writeSubgenomesToFile, echo=FALSE}
+## Now recover which genes are in these sets
+# subset(syntelogs.orthologs, count>=12 & median<=ks_cutoff & org_chr1=="a31607_1" & org_chr2=="b34889_1")
+# orthologs.blocks_groups <- syntelogs.orthologs[syntelogs.orthologs$count>=12 & syntelogs.orthologs$median<=ks_cutoff,]
+subgenome.groups <- rbind(merge(syntelogs.orthologs, orthologs.sub1, all.x = F), merge(syntelogs.orthologs, orthologs.sub2, all.x = F))
+subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
+write.table(subgenome, "../Output/recoveredSubGenomes.tab", sep="\t")
+```

### 2017-07-24 Perl/parseKsKnFile.pl

@@ -1,0 +1,37 @@
+use strict;
+use warnings;
+
+my $block = "";
+print "block\tks\tkn\torg_chr1\tstart1\tstop1\tgene1\torg_chr2\tstart2\tstop2\tgene2\n";
+while(<>) {
+  my @words = split /\t/, $_;
+  if ($words[0] =~ /#\d+/) {
+    if ($words[4] =~ /f/) {
+      $block = $words[0] . "f";
+    } else {
+      $block = $words[0] . "r";
+    }
+    #$block = $words[0];
+    #print "$block \n";
+  } elsif ($words[0] =~ /#Ks/ | $words[0] =~ /#This.*/) {
+    #skip
+  } else {
+    my $ks = $words[0];
+    my $kn = $words[1];
+    my $org_chr1 = $words[2];
+    my @org_features1 = split(/\|\|/, $words[3]);
+    my $gene1 = $org_features1[3];
+    my $start1 = $words[4];
+    my $stop1 = $words[5];
+    my $org_chr2 = $words[6];
+    my @org_features2 = split(/\|\|/, $words[7]);
+    my $gene2 = $org_features2[3];
+    my $start2 = $words[8];
+    my $stop2 = $words[9];
+    my $eval = $words[10];
+    my $block_score = $words[11];
+    my $gevo_link = $words[12];
+
+    print $block, "\t", $ks, "\t", $kn, "\t", $org_chr1, "\t", $start1, "\t", $stop1, "\t", $gene1, "\t", $org_chr2, "\t", $start2, "\t", $stop2, "\t", $gene2, "\n"; 
+  }
+}

### 2017-07-26 R/SubGenome Analysis.Rmd
Added data files and separating functions
Starting to separate functions into individual files, preparing to set this up as a single pipeline.

@@
 output: html_document
+params:
+  fileParser: "parseKsKnFile.pl"
+  inputFile: "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks"
 ---
@@
 
-## R Markdown
+## Maize v1 vs. Sorghum v1 (Based on Schnable 2011 paper)
 Some preprocessing was needed to generate the input files for this project.
@@
 
-This resulted in the file (31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks).
+This resulted in the file (6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks) comparing sorghum v1 to maize v1.
 
-We parse this file to a flat format using a perl script (parseKsKnFile.pl)
+We parse this file to a flat format using a perl script (parseKsKnFile.pl) and manually correct the header lines.
 
@@
 
-> perl parseKsKnFile.pl ../Data/31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum\ vs\ maize.tab
+> perl parseKsKnFile.pl ../Data/6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum_v1_vs_maize_v1.tab
 
 ```{r parseSynMap, include=FALSE}
-cmd <- paste("perl", "parseKsKnFile.pl", "31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks")
+cmd <- paste("perl", params$fileParser, params$inputFile)
 system(cmd)
@@
 ## Import data and remove incomplete rows
-syntelogs.import <- read_delim("../Data/sorghum vs maize.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+syntelogs.import <- read_delim("../Data/sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
 syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]
@@
 subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
-write.table(subgenome, "../Output/recoveredSubGenomes.tab", sep="\t")
+write.table(subgenome, "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab", sep="\t")
 ```
+
+We also want to know how we did compared to the paper itself, to see if this method is faithfully reproducing their results.
+```{r compareResultsToPaper, echo=FALSE}
+library(readxl)
+library(splitstackshape)
+#sd01 <- read_excel("~/git/SubGenomes/Data/sd01.xls")
+#sd01$Maize1[grepl("^GRMZM", sd01$Maize1) | grepl("^AC", sd01$Maize1)]
+```
+
+## Maize v4+rejected set vs. Sorghum v3.1 (updated to use up-to-date assemblies and gene models)
+
+```{r parseSynMap2, include=FALSE}
+cmd <- paste("perl", "parseKsKnFile.pl", "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks")
+system(cmd)
+```
+
+```{r importSynMapData2, echo=FALSE}
+## Import data and remove incomplete rows
+syntelogs.import <- read_delim("../Data/sorghum_v3.1_vs_maize_v4+rejected.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]
+
+sprintf("We have imported %d lines from the parsed SynMap file and removed %d incomplete rows resulting in %d syntenic genes.", nrow(syntelogs.import), nrow(syntelogs.import)- nrow(syntelogs.complete), nrow(syntelogs.complete))
+```
+
+We need to select a cutoff ks value to differentiate between the most recent alpha duplication event and the previous beta duplication.
+
+```{r pickKSCutoff2, echo=FALSE}
+## Pick a ks value which separates alpha duplication from beta duplication (between the first two peaks)
+log10_ks_cutoff <- -0.575
+qplot(log10(syntelogs.complete$ks), geom="histogram", binwidth=.01)  + geom_vline(mapping = NULL, data = NULL, xintercept = log10_ks_cutoff, na.rm = FALSE, show.legend = NA)
+ks_cutoff <- 10^log10_ks_cutoff
+#ks_cutoff <- 0.4234 #include most of the first 2 peaks
+```
+
+Find subgenomes
+
+```{r findSubgenomes2, echo=FALSE}
+## Find which sets of chromosomes with syntelogs should be in group 1 or group 2, where group 1 has larger syntenic blocks
+# First group rows with same syntenic block and chromosome, then summarize each block's ks values with median/mean/count
+syntelogs.aggregate <- ddply(syntelogs.complete, ~block+org_chr1+org_chr2, summarise, median=median(ks), mean=mean(ks), count=length(ks))
+
+# Following the schnable article, syntenic blocks must have 12 genes "The median synonymous substitution rate of all gene pairs in a syntenic block between maize and sorghum can be used to classify syntenic blocks of 12 or more genes unambiguously as orthologous or homoeologous, however" and have a median ks value that discriminates for the alpha duplication event.
+syntelogs.orthologs <- subset(syntelogs.aggregate, count>=12 & median<=ks_cutoff)
+
+# Determine which chromosome has largest syntenic block, the syntelogs on this chromosome are subgenome 1
+orthologs.counts <- ddply(syntelogs.orthologs, ~org_chr1+org_chr2, summarise, total_genes=sum(count))
+orthologs.sorted <- orthologs.counts[order(orthologs.counts["org_chr1"], -orthologs.counts["total_genes"]),]
+
+# orthologs.sorted[match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1),] #first instance of each group (i.e. biggest=sub1)
+orthologs.sub1rows <- match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1) #not exactly right!!!! how to split sub 1 over multiple chromosomes?
+orthologs.sub1 <- orthologs.sorted[orthologs.sub1rows,]
+orthologs.sub2 <- orthologs.sorted[-orthologs.sub1rows,]
+orthologs.sub1$subGenome="sub1"
+orthologs.sub2$subGenome="sub2"
+```
+
+Write to file
+
+```{r writeSubgenomesToFile2, echo=FALSE}
+## Now recover which genes are in these sets
+# subset(syntelogs.orthologs, count>=12 & median<=ks_cutoff & org_chr1=="a31607_1" & org_chr2=="b34889_1")
+# orthologs.blocks_groups <- syntelogs.orthologs[syntelogs.orthologs$count>=12 & syntelogs.orthologs$median<=ks_cutoff,]
+subgenome.groups <- rbind(merge(syntelogs.orthologs, orthologs.sub1, all.x = F), merge(syntelogs.orthologs, orthologs.sub2, all.x = F))
+subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
+write.table(subgenome, "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+rejected.tab", sep="\t")
+```

### 2017-07-26 R/parseForSubgenomes.R

@@ -1,3 +1,6 @@
+#Input: data input fileName, kscutoff
+#Output: file with subgenome
+
 library(readr)
 syntelogs.import <- read_delim("~/Dropbox/SynMap Subgenomes/sorghum vs maize.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
 syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]

### 2017-07-26 R/pickKSCutoff.R

@@ -1,0 +1,8 @@
+#Input
+log10_ks_cutoff <- -0.575
+
+#Show
+qplot(log10(syntelogs.complete$ks), geom="histogram", binwidth=.01)  + geom_vline(mapping = NULL, data = NULL, xintercept = log10_ks_cutoff, na.rm = FALSE, show.legend = NA)
+
+#Output
+ks_cutoff <- 10^log10_ks_cutoff

### 2017-07-27 R/parseForSubgenomes.R
Testing some helper functions
These are probably better off not defined as separate functions, but rather handled in the main document with parameters which point to a dataset to analyze.

@@ -1,12 +1,16 @@
-#Input: data input fileName, kscutoff
-#Output: file with subgenome
-
-library(readr)
-syntelogs.import <- read_delim("~/Dropbox/SynMap Subgenomes/sorghum vs maize.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]
-
-
+####################################################################################################
+## Project: Subgenomes project
+## Script purpose: This script should label each gene in the input file as either part of the dominant
+##                subgenome or the recessive subgenome.
+## Input: data input fileName, kscutoff
+## Output: file with subgenome
+## Date: 2017-07-26
+## Author: Jesse R. Walsh
+####################################################################################################
 library(plyr)
+
+syntelogs.complete <- syntelogs.import[!complete.cases(syntelogs.import),]
+
 ## Find which sets of chromosomes with syntelogs should be in group 1 or group 2
 syntelogs.aggregate <- ddply(syntelogs.complete, ~block+org_chr1+org_chr2, summarise, median=median(ks), mean=mean(ks), count=length(ks))
 syntelogs.orthologs <- subset(syntelogs.aggregate, count>=12 & median<=0.4234)
@@ -29,6 +33,4 @@ subgenome.groups <- rbind(merge(syntelogs.orthologs, orthologs.sub1, all.x = F),
 subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
 write.table(subgenome, "~/Desktop/", sep="\t")
 
-library(reshape2)
-msyntelogs.complete <- melt(syntelogs.complete, id.vars = c("org_chr1", "org_chr2", "block"))
-recast(msyntelogs.complete, block + org_chr1 + org_chr2 ~ variable, mean)
+

### 2017-07-27 R/pickKSCutoff.R

@@ -1,8 +1,16 @@
-#Input
-log10_ks_cutoff <- -0.575
+####################################################################################################
+## Project: Subgenomes project
+## Script purpose: Currently the user picks a ks value from the ks (synonymous mutation rate) histogram
+##                which best separates the alpha duplication event from the beta duplication event.
+##                Since the histogram is displayed in log10, the returned value is 10^kscutoff.
+##
+##                Ideally I will find a way to automatically pick the ks cutoff.
+##
+## Input: ks cutoff value in log10 scale
+## Output:
+## Date: 2017-07-26
+## Author: Jesse R. Walsh
+####################################################################################################
+pickKSCutoff <- function(log10_ks_cutoff) {
 
-#Show
-qplot(log10(syntelogs.complete$ks), geom="histogram", binwidth=.01)  + geom_vline(mapping = NULL, data = NULL, xintercept = log10_ks_cutoff, na.rm = FALSE, show.legend = NA)
-
-#Output
-ks_cutoff <- 10^log10_ks_cutoff
+}

### 2017-07-27 R/SubGenome Analysis.Rmd
Basic import done, ready to start analysis portion

--- 2017-07-24 R/SubGenome Analysis.Rmd
@@
 output: html_document
+params:
+  fileParser: "../Perl/parseKsKnFile.pl"
+
+  #v1 vs. v1
+  synMapOutputFile: "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks"
+  inputDataFile: "../Data/sorghum_v1_vs_maize_v1.tab"
+  outputFile: "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab"
+  log10_ks_cutoff: -0.575
+
+  #v3.1 vs v4+rejected
+  #synMapOutputFile: ""
+  #outputFile: "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+rejected.tab"
+  #inputDataFile: "../Data/sorghum_v3.1_vs_maize_v4+rejected.tab"
+  #log10_ks_cutoff: -0.575
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
-```
-
-## R Markdown
-Some preprocessing was needed to generate the input files for this project.
-We begin with SynMap output based on the Schnable 2011 paper (see https://www.evernote.com/shard/s303/nl/52987123/1ce74f11-98a0-4854-b363-3172e79eeae3 for details).
-
-This resulted in the file (31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks).
-
-We parse this file to a flat format using a perl script (parseKsKnFile.pl)
-
-(Do this in command line until I figure out how to run it from R)
-
-> perl parseKsKnFile.pl ../Data/31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum\ vs\ maize.tab
-
-```{r parseSynMap, include=FALSE}
-cmd <- paste("perl", "parseKsKnFile.pl", "31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks")
-system(cmd)
-```
-
-The data is now prepared for import and cleaning.  We will import the file and remove any incomplete lines.
-
-```{r importSynMapData, echo=FALSE}
+library(readr)
 library(readr)
@@
 library(ggplot2)
+library(readxl)
+```
+```{r preamble, echo=FALSE}
+paste("This analysis is using the file", params$inputDataFile)
+```
 
-## Import data and remove incomplete rows
-syntelogs.import <- read_delim("../Data/sorghum vs maize.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+# Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating subgenome assignments in maize
+We begin with SynMap output based on the [Schnable 2011](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3053962/) paper (see [my notes](https://www.evernote.com/shard/s303/nl/52987123/1ce74f11-98a0-4854-b363-3172e79eeae3) for details on parameter and download options).
+
+## Preprocessing to generate a tidy dataset
+We parse the SynMap output file with ks/kn values (i.e.
+a DAGChainer file) to a flat format using a perl script (parseKsKnFile.pl) and import the data into R.
+Some ks and kn values are empty, so we clean the dataset by removing all incomplete observations.
+
+> manually perform the parse, correct the header lines, and run the import lines until pipeline is debugged
+> perl parseKsKnFile.pl ../Data/6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum_v1_vs_maize_v1.tab
+
+```{r importSynMapData, include=FALSE}
+cmd <- paste("perl", params$fileParser, params$synMapOutputFile)
+system(cmd)
+syntelogs.import <- read_delim(params$inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
 syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]
-
+```
+```{r reportSynMapData, echo=FALSE}
+# Report results of import
 sprintf("We have imported %d lines from the parsed SynMap file and removed %d incomplete rows resulting in %d syntenic genes.", nrow(syntelogs.import), nrow(syntelogs.import)- nrow(syntelogs.complete), nrow(syntelogs.complete))
@@
 
-We need to select a cutoff ks value to differentiate between the most recent alpha duplication event and the previous beta duplication.
+We need to select a cutoff value for ks (synonymous mutation rate) to differentiate between the most recent alpha duplication event and the previous beta duplication.
+If the vertical line separates the first two peaks sufficently, it is probably a good enough value.
+If not, we need to pick a new cutoff value for ks.
 
 ```{r pickKSCutoff, echo=FALSE}
-## Pick a ks value which separates alpha duplication from beta duplication (between the first two peaks)
-log10_ks_cutoff <- -0.575
-qplot(log10(syntelogs.complete$ks), geom="histogram", binwidth=.01)  + geom_vline(mapping = NULL, data = NULL, xintercept = log10_ks_cutoff, na.rm = FALSE, show.legend = NA)
-ks_cutoff <- 10^log10_ks_cutoff
-#ks_cutoff <- 0.4234 #include most of the first 2 peaks
+ks_cutoff <- 10^params$log10_ks_cutoff
+qplot(log10(syntelogs.complete$ks), geom="histogram", binwidth=.01)  + geom_vline(mapping = NULL, data = NULL, xintercept = params$log10_ks_cutoff, na.rm = FALSE, show.legend = NA)
 ```
@@
 
-(but how to group subgenomes over multiple chromosomes?)
+> But how to group subgenomes over multiple chromosomes?
 
@@
 ## Now recover which genes are in these sets
-# subset(syntelogs.orthologs, count>=12 & median<=ks_cutoff & org_chr1=="a31607_1" & org_chr2=="b34889_1")
-# orthologs.blocks_groups <- syntelogs.orthologs[syntelogs.orthologs$count>=12 & syntelogs.orthologs$median<=ks_cutoff,]
 subgenome.groups <- rbind(merge(syntelogs.orthologs, orthologs.sub1, all.x = F), merge(syntelogs.orthologs, orthologs.sub2, all.x = F))
 subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
-write.table(subgenome, "../Output/recoveredSubGenomes.tab", sep="\t")
+write.table(subgenome, params$outputFile, sep="\t")
 ```
+
+We also want to know how we did compared to the paper itself, to see if this method is faithfully reproducing their results.
+```{r compareResultsToPaper, echo=FALSE}
+sd01 <- read_excel("../Data/sd01.xls")
+```

### 2017-07-27 R/SubGenome Analysis.Rmd
Report table of matching chromosomal segments
Also, warn if there are unmatched chromosomes.

@@
 library(readr)
-library(readr)
-library(plyr)
-library(reshape2)
+library(readxl)
+library(plyr)#
+library(dplyr)
+library(tidyr)
+library(reshape2)#
 library(ggplot2)
-library(readxl)
 ```
@@
 
-> manually perform the parse, correct the header lines, and run the import lines until pipeline is debugged
+> manually perform the parse and run the import lines until pipeline is debugged
 > perl parseKsKnFile.pl ../Data/6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum_v1_vs_maize_v1.tab
@@
 system(cmd)
-syntelogs.import <- read_delim(params$inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
-syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]
+syntelogs.raw <- read_delim(params$inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
+syntelogs.complete <- syntelogs.raw[complete.cases(syntelogs.raw),]
 ```
-```{r reportSynMapData, echo=FALSE}
+```{r reportSynMapData, echo=FALSE, results="asis"}
 # Report results of import
-sprintf("We have imported %d lines from the parsed SynMap file and removed %d incomplete rows resulting in %d syntenic genes.", nrow(syntelogs.import), nrow(syntelogs.import)- nrow(syntelogs.complete), nrow(syntelogs.complete))
+cat(sprintf("We have imported %d lines from the parsed SynMap file and removed %d incomplete rows resulting in %d syntenic genes.", nrow(syntelogs.raw), nrow(syntelogs.raw)- nrow(syntelogs.complete), nrow(syntelogs.complete)))
 ```
@@
 
-```{r findSubgenomes, echo=FALSE}
+```{r findSubgenomes, include=FALSE}
 ## Find which sets of chromosomes with syntelogs should be in group 1 or group 2, where group 1 has larger syntenic blocks
@@
 # Following the schnable article, syntenic blocks must have 12 genes "The median synonymous substitution rate of all gene pairs in a syntenic block between maize and sorghum can be used to classify syntenic blocks of 12 or more genes unambiguously as orthologous or homoeologous, however" and have a median ks value that discriminates for the alpha duplication event.
-syntelogs.orthologs <- subset(syntelogs.aggregate, count>=12 & median<=ks_cutoff)
+homeologs <- subset(syntelogs.aggregate, count>=12 & median<=ks_cutoff)
 
 # Determine which chromosome has largest syntenic block, the syntelogs on this chromosome are subgenome 1
-orthologs.counts <- ddply(syntelogs.orthologs, ~org_chr1+org_chr2, summarise, total_genes=sum(count))
-orthologs.sorted <- orthologs.counts[order(orthologs.counts["org_chr1"], -orthologs.counts["total_genes"]),]
+homeologs.counts <- ddply(homeologs, ~org_chr1+org_chr2, summarise, total_genes=sum(count))
+homeologs.sorted <- homeologs.counts[order(homeologs.counts["org_chr1"], -homeologs.counts["total_genes"]),]
 
-# orthologs.sorted[match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1),] #first instance of each group (i.e. biggest=sub1)
-orthologs.sub1rows <- match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1) #not exactly right!!!! how to split sub 1 over multiple chromosomes?
-orthologs.sub1 <- orthologs.sorted[orthologs.sub1rows,]
-orthologs.sub2 <- orthologs.sorted[-orthologs.sub1rows,]
-orthologs.sub1$subGenome="sub1"
-orthologs.sub2$subGenome="sub2"
+# homeologs.sorted[match(unique(homeologs.sorted$org_chr1), homeologs.sorted$org_chr1),] #first instance of each group (i.e. biggest=sub1)
+homeologs.sub1rows <- match(unique(homeologs.sorted$org_chr1), homeologs.sorted$org_chr1) #not exactly right!!!! how to split sub 1 over multiple chromosomes?
+homeologs.sub1 <- homeologs.sorted[homeologs.sub1rows,]
+homeologs.sub2 <- homeologs.sorted[-homeologs.sub1rows,]
+homeologs.sub1$subGenome="sub1"
+homeologs.sub2$subGenome="sub2"
+
+# Reconstructed ancestral chromosomes
+homeologs %>% group_by(org_chr1, org_chr2) %>% summarize(genes=sum(count)) %>% spread(org_chr2, genes) -> homeologs.spread
+homeologs.spread$total <- rowSums(homeologs.spread[-1], na.rm = TRUE)
+syntelogs.complete %>% select(org_chr1) %>% distinct() %>% left_join(homeologs.spread) %>% replace(is.na(.), 0) %>% arrange(desc(total)) -> homeologs.totals
+```
+```{r reportFindSubgenomes, echo=FALSE, results="asis"}
+homeologs.totals %>% knitr::kable(caption = "Table")
+homeologs.totals %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0) %>% sprintf("There are %i unaligned sorghum chromosome(s)", .) %>% cat()
+
 ```
@@
 ## Now recover which genes are in these sets
-subgenome.groups <- rbind(merge(syntelogs.orthologs, orthologs.sub1, all.x = F), merge(syntelogs.orthologs, orthologs.sub2, all.x = F))
+subgenome.groups <- rbind(merge(homeologs, homeologs.sub1, all.x = F), merge(homeologs, homeologs.sub2, all.x = F))
 subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)

### 2017-07-27 R/SubGenome Analysis.Rmd
Minor changes

@@
   outputFile: "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab"
-  log10_ks_cutoff: -0.575
+  log10_ks_cutoff: 0
+  #-0.575
 
@@
 homeologs.spread$total <- rowSums(homeologs.spread[-1], na.rm = TRUE)
-syntelogs.complete %>% select(org_chr1) %>% distinct() %>% left_join(homeologs.spread) %>% replace(is.na(.), 0) %>% arrange(desc(total)) -> homeologs.totals
+syntelogs.complete %>% select(org_chr1) %>% distinct() %>% left_join(homeologs.spread) %>% replace(is.na(.), 0) %>% arrange(desc(org_chr1)) -> homeologs.totals
 ```
@@
 homeologs.totals %>% knitr::kable(caption = "Table")
-homeologs.totals %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0) %>% sprintf("There are %i unaligned sorghum chromosome(s)", .) %>% cat()
+homeologs.totals %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0) %>% sprintf("There are %i unaligned sorghum chromosome(s) in the above table.", .) %>% cat()
 

### 2017-07-27 R/SubGenome Analysis.Rmd
Mutate all useful values and summaries upfront

@@
 library(tidyr)
-library(reshape2)#
+#library(reshape2)
 library(ggplot2)
@@
 
-## Preprocessing to generate a tidy dataset
+## Import Data and Preprocessing
 We parse the SynMap output file with ks/kn values (i.e.
 a DAGChainer file) to a flat format using a perl script (parseKsKnFile.pl) and import the data into R.
-Some ks and kn values are empty, so we clean the dataset by removing all incomplete observations.
-
-> manually perform the parse and run the import lines until pipeline is debugged
-> perl parseKsKnFile.pl ../Data/6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum_v1_vs_maize_v1.tab
+Some ks and kn values are empty, which is to be expected.
 
 ```{r importSynMapData, include=FALSE}
-cmd <- paste("perl", params$fileParser, params$synMapOutputFile)
-system(cmd)
 syntelogs.raw <- read_delim(params$inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
-syntelogs.complete <- syntelogs.raw[complete.cases(syntelogs.raw),]
+#syntelogs.complete <- syntelogs.raw[complete.cases(syntelogs.raw),]
+
+syntelogs.raw %>% 
+  group_by(block, org_chr1, org_chr2) %>% 
+  summarise(median_ks=median(ks, na.rm=TRUE), geneCount=n()) %>% 
+  left_join(syntelogs.raw, ., by = c("block", "org_chr1", "org_chr2")) %>% 
+  mutate(gene_length1=abs(stop1-start1)) %>% 
+  mutate(gene_length2=abs(stop2-start2)) %>% 
+  mutate(chr1=regmatches(org_chr1, regexpr("\\d*$",org_chr1))) %>% 
+  mutate(chr2=regmatches(org_chr2, regexpr("\\d*$",org_chr2))) %>% 
+  mutate(id=paste(block, org_chr1, org_chr2, sep="||")) -> syntelogs.mutated
 ```
@@
 homeologs.totals %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0) %>% sprintf("There are %i unaligned sorghum chromosome(s) in the above table.", .) %>% cat()
-
 ```

### 2017-07-28 R/SubGenome Analysis.Rmd
Processing moved to parseForSubgenomes.R
Also converted to dply syntax for readability and simplified the variables needed as well as create useful outputs to compare homeologs by gene, block, and chromosome.

@@
 library(readxl)
-library(plyr)#
 library(dplyr)
 library(tidyr)
-#library(reshape2)
 library(ggplot2)
@@
 ```{r importSynMapData, include=FALSE}
-syntelogs.raw <- read_delim(params$inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
-#syntelogs.complete <- syntelogs.raw[complete.cases(syntelogs.raw),]
-
-syntelogs.raw %>% 
-  group_by(block, org_chr1, org_chr2) %>% 
-  summarise(median_ks=median(ks, na.rm=TRUE), geneCount=n()) %>% 
-  left_join(syntelogs.raw, ., by = c("block", "org_chr1", "org_chr2")) %>% 
-  mutate(gene_length1=abs(stop1-start1)) %>% 
-  mutate(gene_length2=abs(stop2-start2)) %>% 
-  mutate(chr1=regmatches(org_chr1, regexpr("\\d*$",org_chr1))) %>% 
-  mutate(chr2=regmatches(org_chr2, regexpr("\\d*$",org_chr2))) %>% 
-  mutate(id=paste(block, org_chr1, org_chr2, sep="||")) -> syntelogs.mutated
+source("parseForSubgenomes.R")
 ```
@@
 # Report results of import
-cat(sprintf("We have imported %d lines from the parsed SynMap file and removed %d incomplete rows resulting in %d syntenic genes.", nrow(syntelogs.raw), nrow(syntelogs.raw)- nrow(syntelogs.complete), nrow(syntelogs.complete)))
+cat(sprintf("We have imported %d lines from the parsed SynMap file.", nrow(syntelogs.raw)))
 ```
@@
 ```{r pickKSCutoff, echo=FALSE}
-ks_cutoff <- 10^params$log10_ks_cutoff
 qplot(log10(syntelogs.complete$ks), geom="histogram", binwidth=.01)  + geom_vline(mapping = NULL, data = NULL, xintercept = params$log10_ks_cutoff, na.rm = FALSE, show.legend = NA)
@@
 
-> But how to group subgenomes over multiple chromosomes?
+```{r reportFindSubgenomes, echo=FALSE, results="asis"}
+## Display chromosomes with significant syntenic genes between org1 and org2
+homeologs.sorted %>%
+  select(org_chr1, org_chr2, chromosomeGeneCount) %>%
+  spread(org_chr2, chromosomeGeneCount) -> homeologs.spread
 
-```{r findSubgenomes, include=FALSE}
-## Find which sets of chromosomes with syntelogs should be in group 1 or group 2, where group 1 has larger syntenic blocks
-# First group rows with same syntenic block and chromosome, then summarize each block's ks values with median/mean/count
-syntelogs.aggregate <- ddply(syntelogs.complete, ~block+org_chr1+org_chr2, summarise, median=median(ks), mean=mean(ks), count=length(ks))
+homeologs.spread$total <- rowSums(homeologs.spread[-1], na.rm = TRUE)
 
-# Following the schnable article, syntenic blocks must have 12 genes "The median synonymous substitution rate of all gene pairs in a syntenic block between maize and sorghum can be used to classify syntenic blocks of 12 or more genes unambiguously as orthologous or homoeologous, however" and have a median ks value that discriminates for the alpha duplication event.
-homeologs <- subset(syntelogs.aggregate, count>=12 & median<=ks_cutoff)
-
-# Determine which chromosome has largest syntenic block, the syntelogs on this chromosome are subgenome 1
-homeologs.counts <- ddply(homeologs, ~org_chr1+org_chr2, summarise, total_genes=sum(count))
-homeologs.sorted <- homeologs.counts[order(homeologs.counts["org_chr1"], -homeologs.counts["total_genes"]),]
-
-# homeologs.sorted[match(unique(homeologs.sorted$org_chr1), homeologs.sorted$org_chr1),] #first instance of each group (i.e. biggest=sub1)
-homeologs.sub1rows <- match(unique(homeologs.sorted$org_chr1), homeologs.sorted$org_chr1) #not exactly right!!!! how to split sub 1 over multiple chromosomes?
-homeologs.sub1 <- homeologs.sorted[homeologs.sub1rows,]
-homeologs.sub2 <- homeologs.sorted[-homeologs.sub1rows,]
-homeologs.sub1$subGenome="sub1"
-homeologs.sub2$subGenome="sub2"
-
-# Reconstructed ancestral chromosomes
-homeologs %>% group_by(org_chr1, org_chr2) %>% summarize(genes=sum(count)) %>% spread(org_chr2, genes) -> homeologs.spread
-homeologs.spread$total <- rowSums(homeologs.spread[-1], na.rm = TRUE)
-syntelogs.complete %>% select(org_chr1) %>% distinct() %>% left_join(homeologs.spread) %>% replace(is.na(.), 0) %>% arrange(desc(org_chr1)) -> homeologs.totals
-```
-```{r reportFindSubgenomes, echo=FALSE, results="asis"}
-homeologs.totals %>% knitr::kable(caption = "Table")
-homeologs.totals %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0) %>% sprintf("There are %i unaligned sorghum chromosome(s) in the above table.", .) %>% cat()
+homeologs.spread %>% knitr::kable(caption = "Table")
+homeologs.spread %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0) %>% sprintf("There are %i unaligned sorghum chromosome(s) in the above table.", .) %>% cat()
 ```
@@
 ## Now recover which genes are in these sets
-subgenome.groups <- rbind(merge(homeologs, homeologs.sub1, all.x = F), merge(homeologs, homeologs.sub2, all.x = F))
-subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
 write.table(subgenome, params$outputFile, sep="\t")

### 2017-07-28 R/parseForSubgenomes.R

@@ -2,35 +2,67 @@
 ## Project: Subgenomes project
 ## Script purpose: This script should label each gene in the input file as either part of the dominant
 ##                subgenome or the recessive subgenome.
-## Input: data input fileName, kscutoff
-## Output: file with subgenome
+##
+## Input: data input fileName, log10_ks_cutoff
+## Output: not really meant to have defined output, since several intermediary objects can be of use
+##        syntelogs.raw         ->
+##        syntelogs.mutated     ->
+##        homeologs.genes       ->
+##        homeologs.block       ->
+##        homeologs.chromosome  ->
+##        subgenome             ->
+##        ks_cutoff             ->
 ## Date: 2017-07-26
 ## Author: Jesse R. Walsh
 ####################################################################################################
-library(plyr)
 
-syntelogs.complete <- syntelogs.import[!complete.cases(syntelogs.import),]
+inputDataFile <- params$inputDataFile
+log10_ks_cutoff <- params$log10_ks_cutoff
 
-## Find which sets of chromosomes with syntelogs should be in group 1 or group 2
-syntelogs.aggregate <- ddply(syntelogs.complete, ~block+org_chr1+org_chr2, summarise, median=median(ks), mean=mean(ks), count=length(ks))
-syntelogs.orthologs <- subset(syntelogs.aggregate, count>=12 & median<=0.4234)
-# syntelogs.orthologs[order(syntelogs.orthologs["org_chr1"], syntelogs.orthologs["org_chr2"]),]#view sorted orthologs
-orthologs.counts <- ddply(syntelogs.orthologs, ~org_chr1+org_chr2, summarise, total_genes=sum(count))
-orthologs.sorted <- orthologs.counts[order(orthologs.counts["org_chr1"], -orthologs.counts["total_genes"]),]
+## Import the raw data from the parsed SynMap output
+syntelogs.raw <- read_delim(inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
 
-# orthologs.sorted[match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1),] #first instance of each group (i.e. biggest=sub1)
[…]

### 2017-08-04 R/SubGenome Analysis.Rmd
Updated plot function
*Not in a complete state yet

@@
   #v1 vs. v1
-  synMapOutputFile: "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks"
-  inputDataFile: "../Data/sorghum_v1_vs_maize_v1.tab"
-  outputFile: "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab"
+  # synMapOutputFile: "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks"
+  # inputDataFile: "../Data/sorghum_v1_vs_maize_v1.tab"
+  # outputFile: "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab"
+  # log10_ks_cutoff: 0
+
+  #v3.1 vs v4+rejected
+  synMapOutputFile: "31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords"
+  inputDataFile: "../Data/sorghum_v3.1_vs_maize_v4+rejected.tab"
+  outputFile: "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+rejected.tab"
   log10_ks_cutoff: 0
   #-0.575
-
-  #v3.1 vs v4+rejected
-  #synMapOutputFile: ""
-  #outputFile: "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+rejected.tab"
-  #inputDataFile: "../Data/sorghum_v3.1_vs_maize_v4+rejected.tab"
-  #log10_ks_cutoff: -0.575
 ---
@@
 If not, we need to pick a new cutoff value for ks.
+Note that looking at ks values by gene produces no signal.
+Looking at median ks values by syntenic block (kindof) shows two peaks, first for orthologs and second for homeologs from "pregrass tetraploidy".
 
 ```{r pickKSCutoff, echo=FALSE}
-qplot(log10(syntelogs.complete$ks), geom="histogram", binwidth=.01)  + geom_vline(mapping = NULL, data = NULL, xintercept = params$log10_ks_cutoff, na.rm = FALSE, show.legend = NA)
+qplot(
+  syntelogs.mutated$ks[syntelogs.mutated$ks < 2.5], 
+  geom="histogram", binwidth=.05, na.rm=TRUE
+  ) + geom_vline(
+    mapping = NULL, 
+    data = NULL, 
+    xintercept = 10^params$log10_ks_cutoff, 
+    na.rm = FALSE, 
+    show.legend = NA
+  ) + labs(
+    title="Frequency of KS values by syntenic block (only KS values < 2.5)", 
+    x="KS value of syntenic block", 
+    y="Count (# of genes)"
+  )
+
+qplot(syntelogs.mutated$median_ks[syntelogs.mutated$median_ks < 2.5], geom="histogram", binwidth=.05, na.rm=TRUE)  + geom_vline(mapping = NULL, data = NULL, xintercept = 10^params$log10_ks_cutoff, na.rm = FALSE, show.legend = NA) + labs(title="Frequency of median KS values by syntenic block (only median KS values < 2.5)", x="Median KS value of syntenic block", y="Count (# of genes)")
+
+qplot(log10(syntelogs.raw$ks), geom="histogram", binwidth=.01, na.rm=TRUE)  + geom_vline(mapping = NULL, data = NULL, xintercept = params$log10_ks_cutoff, na.rm = FALSE, show.legend = NA) + labs(title="Frequency of log10 KS values", x="log10(KS) value of genes", y="Count (# of genes)")
 ```
@@
 
-```{r reportFindSubgenomes, echo=FALSE, results="asis"}
-## Display chromosomes with significant syntenic genes between org1 and org2
-homeologs.sorted %>%
+```{r reportAncestralChromosomes, echo=FALSE, results="asis"}
+## Display chromosomes with homeologous genes between org1 and org2
+homeologs.chromosome %>%
   select(org_chr1, org_chr2, chromosomeGeneCount) %>%
@@
 ## Now recover which genes are in these sets
-write.table(subgenome, params$outputFile, sep="\t")
+#write.table(subgenome, params$outputFile, sep="\t")
 ```
@@
 ```{r compareResultsToPaper, echo=FALSE}
-sd01 <- read_excel("../Data/sd01.xls")
+#sd01 <- read_excel("../Data/sd01.xls")
 ```

### 2017-08-04 R/SubGenome Analysis.Rmd
Minor

@@
 knitr::opts_chunk$set(echo = TRUE)
-library(readr)
 library(readxl)
-library(dplyr)
 library(tidyr)

### 2017-08-07 R/Test Files/junkCode.R
Moving unused code to separate folder
Not ready to delete yet.  Also tested some iRange and gRange package a little in junkCode.R

@@ -75,3 +75,26 @@ write.table(subgenome, "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+r
 library(reshape2)
 msyntelogs.complete <- melt(syntelogs.complete, id.vars = c("org_chr1", "org_chr2", "block"))
 recast(msyntelogs.complete, block + org_chr1 + org_chr2 ~ variable, mean)
+
+
+
+
+
+
+
+
+
+#https://www.bioconductor.org/help/course-materials/2015/Uruguay2015/V3-GenomicRanges.html
+# Testing bioconductor IRanges
+source("https://bioconductor.org/biocLite.R")
+biocLite()
+biocLite("GenomicRanges")
+
+library(IRanges)
+ir1 <- IRanges(start = syntelogs.raw$start1, end = syntelogs.raw$stop1)
+ir1
+library(GenomicRanges)
+gr <- GRanges(seqnames = syntelogs.raw$gene1, strand = rep("*",nrows(syntelogs.raw)), start = syntelogs.raw$start1, end = syntelogs.raw$stop1)
+
+
+install.packages("rPython")

### 2017-08-07 R/SubGenome Analysis.Rmd
Updated to point to renamed file

@@
 ```{r importSynMapData, include=FALSE}
-source("parseForSubgenomes.R")
+source("parseSubgenomes.R")
 ```

### 2017-08-09 R/parseSubgenomes.R
Updated subgenome detection

@@ -66,27 +66,28 @@ homeologs.chromosome <-
 # Reconstructed ancestral chromosomes
 # ! This isn't quite right, since it assumes the single largest homologous.chromosome is the sub1 genome, but sometimes (half the time?)
 # ! the sub1 genome is split over multiple chromosomes
-subgenome <-
-  match(unique(homeologs.chromosome$chr1), homeologs.chromosome$chr1) %>%
-  slice(homeologs.chromosome, .) %>%
-  select(org_chr1, org_chr2) %>%
-  mutate(subgenome="sub1") %>%
-  left_join(homeologs.genes, ., by=c("org_chr1", "org_chr2"))
-
-subgenome$subgenome[is.na(subgenome$subgenome)] <- "sub2"
+# subgenome <-
+#   match(unique(homeologs.chromosome$chr1), homeologs.chromosome$chr1) %>%
+#   slice(homeologs.chromosome, .) %>%
+#   select(org_chr1, org_chr2) %>%
+#   mutate(subgenome="sub1") %>%
+#   left_join(homeologs.genes, ., by=c("org_chr1", "org_chr2"))
+#
+# subgenome$subgenome[is.na(subgenome$subgenome)] <- "sub2"
 
 ## Add chromosome wide start/stop values, output for greedy algorithm processing in external lanquage
 # ! this is a better method than above chunk
-# homeologs.chromosomeStopStart <-
-#   syntelogs.raw %>%
-#   inner_join(homeologs.block, by=c("block", "org_chr1", "org_chr2")) %>%
-#   mutate(low = pmin(start1, stop1), high = pmax(start1, stop1)) %>%
-#   group_by(org_chr1, org_chr2) %>%
-#   summarise(chrLow = min(low,high), chrHigh = max(low,high)) %>%
-#   left_join(homeologs.chromosome, ., by=c("org_chr1", "org_chr2")) %>%
-#   select(org_chr1, org_chr2, chromosomeGeneCount, chrLow, chrHigh)
+homeologs.chromosomeStopStart <-
+  syntelogs.raw %>%
+  inner_join(homeologs.block, by=c("block", "org_chr1", "org_chr2")) %>%
+  mutate(low = pmin(start1, stop1), high = pmax(start1, stop1)) %>%
+  group_by(org_chr1, org_chr2) %>%
+  summarise(chrLow = min(low,high), chrHigh = max(low,high)) %>%
[…]

### 2017-08-09 R/addGOTerms.R
Updated go annotation assignment

@@ -37,11 +37,64 @@ subgenome %>%
   distinct() -> subgenome.sub2
 
 # Rows from go annotation file annotating a gene from sub1 or sub2
-goAnnotations[goAnnotations$`MaizeCyc2.2 Accession-1` %in% subgenome.sub1$gene2,]
-goAnnotations[goAnnotations$`MaizeCyc2.2 Accession-1` %in% subgenome.sub2$gene2,]
+goAnnotations.sub1 <-
+  goAnnotations[goAnnotations$`MaizeCyc2.2 Accession-1` %in% subgenome.sub1$gene2,] %>%
+  select(`MaizeCyc2.2 Accession-1`, `GO Term`)
+goAnnotations.sub2 <-
+  goAnnotations[goAnnotations$`MaizeCyc2.2 Accession-1` %in% subgenome.sub2$gene2,] %>%
+  select(`MaizeCyc2.2 Accession-1`, `GO Term`)
+
+# write.table(goAnnotations.sub1, "goAnnotationsSub1.tab", sep="\t")
+# write.table(goAnnotations.sub2, "goAnnotationsSub2.tab", sep="\t")
 
 #
-goAnnotations[goAnnotations$`MaizeCyc2.2 Accession-1` %in% subgenome.sub1$gene2,] %>%
+# goAnnotations[goAnnotations$`MaizeCyc2.2 Accession-1` %in% subgenome.sub1$gene2,] %>%
+#   select(`GO Term`) %>%
+#   group_by(`GO Term`) %>%
+#   summarise(count(`GO Term`))
+
+
+### GO Analysis -> grouping into useful slices
+goAnnotations.sub1.aggr <-
+  goAnnotations.sub1 %>%
   select(`GO Term`) %>%
   group_by(`GO Term`) %>%
-  summarise(count(`GO Term`))
+  summarise(n=n())
+goAnnotations.sub2.aggr <-
+  goAnnotations.sub2 %>%
+  select(`GO Term`) %>%
+  group_by(`GO Term`) %>%
+  summarise(n=n())
+goAnnotations.diff <-
+  full_join(goAnnotations.sub1.aggr, goAnnotations.sub2.aggr, by="GO Term", is.na=0) %>%
+  rename(sub1_n=n.x, sub2_n=n.y) %>%
+  subset(is.na(sub1_n) | is.na(sub2_n))
+goAnnotations.diff.counts <-
[…]

### 2017-08-09 R/SubGenome Analysis.Rmd
Minor

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(readr)
 library(readxl)
+library(ggplot2)
 library(tidyr)
-library(ggplot2)
+library(dplyr)
 ```

### 2017-08-10 R/SubGenome Analysis.Rmd
Reformatting
Fixed comments, titles, captions, plot sizes, and reorganized for better readability.  Added some basic stats on the input data.

--- 2017-08-07 R/SubGenome Analysis.Rmd
@@
 ---
-title: "SubGenome Analysis"
+title: "Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating subgenome assignments in maize"
 output: html_document
@@
   #v1 vs. v1
-  # synMapOutputFile: "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks"
-  # inputDataFile: "../Data/sorghum_v1_vs_maize_v1.tab"
-  # outputFile: "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab"
-  # log10_ks_cutoff: 0
+  synMapOutputFile: "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks"
+  inputDataFile: "../Data/sorghum_v1_vs_maize_v1.tab"
+  outputFile: "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab"
+  log10_ks_cutoff: 0
 
   #v3.1 vs v4+rejected
-  synMapOutputFile: "31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords"
-  inputDataFile: "../Data/sorghum_v3.1_vs_maize_v4+rejected.tab"
-  outputFile: "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+rejected.tab"
-  log10_ks_cutoff: 0
+  # synMapOutputFile: "31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords"
+  # inputDataFile: "../Data/sorghum_v3.1_vs_maize_v4+rejected.tab"
+  # outputFile: "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+rejected.tab"
+  # log10_ks_cutoff: 0
   #-0.575
@@
 knitr::opts_chunk$set(echo = TRUE)
+library(readr)
 library(readxl)
+library(ggplot2)
 library(tidyr)
-library(ggplot2)
-```
-```{r preamble, echo=FALSE}
-paste("This analysis is using the file", params$inputDataFile)
+library(dplyr)
 ```
 
-# Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating subgenome assignments in maize
+This analysis is based on the file: **`r params$inputDataFile`**
+
 We begin with SynMap output based on the [Schnable 2011](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3053962/) paper (see [my notes](https://www.evernote.com/shard/s303/nl/52987123/1ce74f11-98a0-4854-b363-3172e79eeae3) for details on parameter and download options).
 
-## Import Data and Preprocessing
+# Import Data and Preprocessing
+***
 We parse the SynMap output file with ks/kn values (i.e.
@@
 Some ks and kn values are empty, which is to be expected.
+These are ignored for the purposes of finding the median, but included in the results if they are in a block that meets the cutoff criteria.
+GO terms related to Maize gene models are imported as well.
 
 ```{r importSynMapData, include=FALSE}
-source("parseSubgenomes.R")
-```
-```{r reportSynMapData, echo=FALSE, results="asis"}
-# Report results of import
-cat(sprintf("We have imported %d lines from the parsed SynMap file.", nrow(syntelogs.raw)))
+source("~/git/SubGenomes/R/parseSubgenomes.R")
+# source("~/git/SubGenomes/R/addGOTerms.R")
 ```
 
+Lets take a look at what we just imported.
+
+***  
+ * There were **`r syntelogs.raw %>% nrow()`** rows of data in the SynMap file.
+ * There were **`r syntelogs.raw %>% select(gene2) %>% distinct() %>% nrow()`** distinct Maize genes with synteny to Sorghum.
+ * Summary of the KS values
+     + **`r syntelogs.raw %>% select(ks) %>% summary()`**
+ * Out of the `r syntelogs.raw %>% select(gene2) %>% distinct() %>% nrow()` distinct Maize genes with synteny to sorghum, we keep **`r homeologs.genes %>% select(gene2) %>% distinct() %>% nrow()`** after applying the KS cutoff.
+ *
+
+***
+
+# Choosing a KS value cutoff
+***  
 We need to select a cutoff value for ks (synonymous mutation rate) to differentiate between the most recent alpha duplication event and the previous beta duplication.
+Looking at ks value frequency (left), we should not see any obvious cutoff point.
+By transforming ks values to block median values, we should see two (or more) peaks (center).
+In this (center) graph, the first peak represents orthologs and the second represents homeologs from "pregrass tetraploidy".
 If the vertical line separates the first two peaks sufficently, it is probably a good enough value.
 If not, we need to pick a new cutoff value for ks.
-Note that looking at ks values by gene produces no signal.
-Looking at median ks values by syntenic block (kindof) shows two peaks, first for orthologs and second for homeologs from "pregrass tetraploidy".
+Viewing the log10 transform of the ks values helps differentiate types of homologs (better version of this graph with color can be generated in CoGe).
 
-```{r pickKSCutoff, echo=FALSE}
+You have selected the log10(ks) cutoff value to be **`r params$log10_ks_cutoff`**.  Again, this should ideally separate two peaks in the middle graph.
+
+```{r pickKSCutoff, echo=FALSE, fig.width=3, fig.height=3}
 qplot(
@@
 ) + labs(
-    title="Frequency of KS values by syntenic block (only KS values < 2.5)", 
+  title="Frequency of KS values by \nsyntenic block (KS < 2.5)", 
   x="KS value of syntenic block", 
@@
 
-qplot(syntelogs.mutated$median_ks[syntelogs.mutated$median_ks < 2.5], geom="histogram", binwidth=.05, na.rm=TRUE)  + geom_vline(mapping = NULL, data = NULL, xintercept = 10^params$log10_ks_cutoff, na.rm = FALSE, show.legend = NA) + labs(title="Frequency of median KS values by syntenic block (only median KS values < 2.5)", x="Median KS value of syntenic block", y="Count (# of genes)")
+qplot(
+  syntelogs.mutated$median_ks[syntelogs.mutated$median_ks < 2.5], 
+  geom="histogram", 
+  binwidth=.05, 
+  na.rm=TRUE
+) + geom_vline(
+  mapping = NULL, 
+  data = NULL, 
+  xintercept = 10^params$log10_ks_cutoff, 
+  na.rm = FALSE, 
+  show.legend = NA
+) + labs(
+  title="Frequency of median KS \nvalues by syntenic block \n(median KS < 2.5)", 
+  x="Median KS value of syntenic block", 
+  y="Count (# of genes)"
+)
 
-qplot(log10(syntelogs.raw$ks), geom="histogram", binwidth=.01, na.rm=TRUE)  + geom_vline(mapping = NULL, data = NULL, xintercept = params$log10_ks_cutoff, na.rm = FALSE, show.legend = NA) + labs(title="Frequency of log10 KS values", x="log10(KS) value of genes", y="Count (# of genes)")
+qplot(
+  log10(syntelogs.raw$ks), 
+  geom="histogram", 
+  binwidth=.01, 
+  na.rm=TRUE
+) + geom_vline(
+  mapping = NULL, 
+  data = NULL, 
+  xintercept = params$log10_ks_cutoff, 
+  na.rm = FALSE, 
+  show.legend = NA
+) + labs(
+  title="Frequency of log10 KS \nvalues", 
+  x="log10(KS) value of genes", 
+  y="Count (# of genes)"
+)
 ```
 
-In order to separate subgenome 1 from subgenome 2, we need to find syntenic blocks of 12 or more with a median ks value below the cutoff to determine syntelogs which are related to the alpha duplication event.
-For each chromosome with duplicate genes, the chromosome that has the largest syntenic block is the dominant subgenome, and the lesser sized blocks on other chromosomes are the recessive subgenome.
+# Major Syntenic Blocks
+***
+Regions of synteny are not confined to a 1:1 relationship between Sorghum and Maize chromosomes.
+Where multiple regions in Maize are syntenic to the same region on Sorghum, we can assume this is a result of the whole genome duplication even.
+These are the regions which will be sorted into subgenome 1 and subgenome 2.
+Here we report on the results of the subgenome determination.
 
-```{r reportAncestralChromosomes, echo=FALSE, results="asis"}
+~~In order to separate subgenome 1 from subgenome 2, we need to find syntenic blocks of 12 or more with a median ks value below the cutoff to determine syntelogs which are related to the alpha duplication event.
+For each chromosome with duplicate genes, the chromosome that has the largest syntenic block is the dominant subgenome, and the lesser sized blocks on other chromosomes are the recessive subgenome.~~ greedy algorithm
+
+```{r reportSyntenicRegions, echo=FALSE, results="asis"}
 ## Display chromosomes with homeologous genes between org1 and org2
@@
 
-homeologs.spread %>% knitr::kable(caption = "Table")
-homeologs.spread %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0) %>% sprintf("There are %i unaligned sorghum chromosome(s) in the above table.", .) %>% cat()
+# De-emphasize na's by turning them light grey
+homeologs.spread.formattedTable <- homeologs.spread
+homeologs.spread.formattedTable[is.na(homeologs.spread.formattedTable)] <- "<font color=\"lightgrey\">na</font>"
+homeologs.spread.formattedTable %>% 
+  knitr::kable(caption = "Table", 
+               align = "lccccccccccr")
 ```
 
-Now that we have recovered the dominant and recessive subgenomes, lets write this to a file.
+There are **<font color=\"darkgreen\">`r homeologs.spread %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0)`</font>** unaligned sorghum chromosome(s) in the above table.
+If there are any unaligned sorghum chromosomes, consider picking a new ks cutoff.
+
+# SubGenomes
+***
+```{r reportSubGenomes, echo=FALSE, results="asis"}
+
+```
+
+## Compared to the original paper.
+
+```{r compareResultsToPaper, echo=FALSE}
+```
+
+# GOTerms
+***
+
+***
+
+The results of this report are available at **`r params$outputFile`**.
 
@@
 ```
-
-We also want to know how we did compared to the paper itself, to see if this method is faithfully reproducing their results.
-```{r compareResultsToPaper, echo=FALSE}
-#sd01 <- read_excel("../Data/sd01.xls")
-```

### 2017-08-10 R/parseSubgenomes.R

@@ -40,8 +40,10 @@ syntelogs.mutated <-
 ks_cutoff <- 10^log10_ks_cutoff
 
 ## Find which sets of chromosomes with syntelogs should be in group 1 or group 2, where group 1 has larger syntenic blocks
-# First group rows with same syntenic block and chromosome, then summarize each block's ks values with median/mean/count
-# Following the schnable article, syntenic blocks must have 12 genes "The median synonymous substitution rate of all gene pairs in a syntenic block between maize and sorghum can be used to classify syntenic blocks of 12 or more genes unambiguously as orthologous or homoeologous, however" and have a median ks value that discriminates for the alpha duplication event.
+## First group rows with same syntenic block and chromosome, then summarize each block's ks values with median/mean/count
+## Following the schnable article, syntenic blocks must have 12 genes "The median synonymous substitution rate of all gene
+## pairs in a syntenic block between maize and sorghum can be used to classify syntenic blocks of 12 or more genes unambiguously
+## as orthologous or homoeologous, however" and have a median ks value that discriminates for the alpha duplication event.
 homeologs.block <-
   syntelogs.mutated %>%
   select(block, chr1, org_chr1, chr2, org_chr2, median_ks, blockGeneCount) %>%
@@ -82,7 +84,7 @@ subgenome <-
   select(org_chr1, org_chr2, subgenome) %>%
   left_join(homeologs.genes, ., by=c("org_chr1", "org_chr2"))
 
-# Also, since we know this comparison is 1 sorghum gene = 2 maize genes, lets group the homeologous pairs
+## Also, since we know this comparison is 1 sorghum gene = 2 maize genes, lets group the homeologous pairs
 homeologs.pairs <-
   subgenome %>%
   select(gene1, gene2, subgenome) %>%
@@ -94,7 +96,7 @@ homeologs.pairs <-
   group_by(gene1) %>%
   summarise(Maize1=trimws(toString(na.omit(sub1))), Maize2=trimws(toString(na.omit(sub2))))
 
-# Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2
+## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2
 subgenome.homeologs <-
   homeologs.pairs %>%
   subset(Maize1 != "" & Maize2 != "")

### 2017-08-10 R/SubGenome Analysis.Rmd
Added additional basic stats and comparison of subgenomes to paper

@@
 
-# Import Data and Preprocessing
+# Import SynMap and GO Term Data
 ***
@@
 source("~/git/SubGenomes/R/parseSubgenomes.R")
-# source("~/git/SubGenomes/R/addGOTerms.R")
+source("~/git/SubGenomes/R/addGOTerms.R")
 ```
@@
  * Out of the `r syntelogs.raw %>% select(gene2) %>% distinct() %>% nrow()` distinct Maize genes with synteny to sorghum, we keep **`r homeologs.genes %>% select(gene2) %>% distinct() %>% nrow()`** after applying the KS cutoff.
- *
+ * SubGenome 1
+     + **`r subgenome %>% filter(subgenome == "sub1") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 1
+     + **`r goAnnotations.sub1 %>% nrow()`** GO assignments using **`r goAnnotations.sub1 %>% select('GO Term') %>% distinct() %>% nrow()`** GO Terms
+ * SubGenome 2
+     + **`r subgenome %>% filter(subgenome == "sub2") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 2
+     + **`r goAnnotations.sub2 %>% nrow()`** GO assignments using **`r goAnnotations.sub2 %>% select('GO Term') %>% distinct() %>% nrow()`** GO Terms
+ * Rejected
+     + **`r subgenome %>% filter(subgenome == "rejected:small") %>% select(gene2) %>% distinct() %>% nrow()`** genes were rejected during subgenome parsing
+     (Too few syntenic genes across chromosome)
 
-***
-
-# Choosing a KS value cutoff
+# Choosing a KS cutoff value
 ***  
@@
 These are the regions which will be sorted into subgenome 1 and subgenome 2.
-Here we report on the results of the subgenome determination.
-
-~~In order to separate subgenome 1 from subgenome 2, we need to find syntenic blocks of 12 or more with a median ks value below the cutoff to determine syntelogs which are related to the alpha duplication event.
-For each chromosome with duplicate genes, the chromosome that has the largest syntenic block is the dominant subgenome, and the lesser sized blocks on other chromosomes are the recessive subgenome.~~ greedy algorithm
 
@@
 ***
+In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size.
+
 ```{r reportSubGenomes, echo=FALSE, results="asis"}
-
+# library(venneuler)
+# i <- intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
+# t <- setdiff(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
+# p <- setdiff(subgenome.truth, subgenome.chromosomes.test) %>% nrow()
+# vd <- venneuler(c(Paper=p, Test=t, "Paper&Test"=i))
+# vd$labels <- c(
+#   paste("Paper\n",p+i),
+#   paste("Test\n",t+i),
+#   paste("Both\n",i)
+# )
+# plot(vd)
 ```
@@
 ## Compared to the original paper.
+How many chromosomes with synteny to sorghum were correctly placed in subgenome 1 or subgenome 2?
+Items in "Paper" only are missed subgenome assignments.
+Items in "Test" only are false assignments.
+Items in center are correct.
+(this is an unfair comparison if I sorted by hand)
 
-```{r compareResultsToPaper, echo=FALSE}
+```{r compareResultsToPaper, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=3, fig.align='center'}
+library(VennDiagram)
+# i <- intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
+# t <- setdiff(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
+# p <- setdiff(subgenome.truth, subgenome.chromosomes.test) %>% nrow()
+subgenome.chromosomes.test <- 
+  subgenome.chromosomes %>% 
+  mutate(chr1=as.numeric(regmatches(org_chr1, regexpr("\\d*$",org_chr1)))) %>%
+  mutate(chr2=as.numeric(regmatches(org_chr2, regexpr("\\d*$",org_chr2)))) %>%
+  select(chr1, chr2, subgenome) %>%
+  subset(subgenome == "sub1" | subgenome == "sub2")
+
+draw.pairwise.venn(
+  area1 = subgenome.truth%>%nrow(), 
+  area2 = subgenome.chromosomes.test%>%nrow(), 
+  cross.area = intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow(), 
+  category = c("Paper", "Test"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0
+)
 ```

### 2017-08-11 R/SubGenome Analysis.Rmd
Hardcoded support for windows file paths

@@
   fileParser: "../Perl/parseKsKnFile.pl"
+  goDataFile: "../Data/GO from maizecyc.tab"
+  subGenomeFile: "/home/jesse/Dropbox/deleteme/outfile_processed_byHand.tab"
 
@@
 ```{r importSynMapData, include=FALSE}
-source("~/git/SubGenomes/R/parseSubgenomes.R")
-source("~/git/SubGenomes/R/addGOTerms.R")
+## source("~/git/SubGenomes/R/parseSubgenomes.R")
+## source("~/git/SubGenomes/R/addGOTerms.R")
+
+source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseSubgenomes.R")
+source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\addGOTerms.R")
 ```

### 2017-08-11 R/addGOTerms.R

@@ -11,9 +11,12 @@ library(dplyr)
 ## Author: Jesse R. Walsh
 ####################################################################################################
 
+goDataFile <- params$goDataFile
+
+goDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO from maizecyc.tab"
+
 ## Read in GO Annotation data for maize genes
-goAnnotations <- read_delim("../Data/GO from maizecyc.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-# goAnnotations <- read_delim("./Data/GO from maizecyc.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+goAnnotations <- read_delim(goDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
 
 ## Merge v3 homeolog gene IDs to the GO term data, each line has a unique Gene -> GO term pair, where genes and go terms can be reused
 ## This list is only genes that have homeologs and have GO annotation

### 2017-08-11 R/parseSubgenomes.R

@@ -20,7 +20,11 @@ library(dplyr)
 ####################################################################################################
 
 inputDataFile <- params$inputDataFile
+subGenomeFile <- params$subGenomeFile
 log10_ks_cutoff <- params$log10_ks_cutoff
+
+inputDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\sorghum_v1_vs_maize_v1.tab"
+subGenomeFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\deleteme\\outfile_processed_byHand.tab"
 
 ## Import the raw data from the parsed SynMap output
 syntelogs.raw <- read_delim(inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
@@ -86,7 +90,7 @@ homeologs.chromosomeStopStart <-
 
 ## Currently need to perform the greedy sorting by hand... need to automate this step.
 # write.table(homeologs.chromosomeStopStart, "outfile.tab", sep="\t")
-subgenome.chromosomes <- read_delim("/home/jesse/Dropbox/deleteme/outfile_processed_byHand.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+subgenome.chromosomes <- read_delim(subGenomeFile, "\t", escape_double = FALSE, trim_ws = TRUE)
 
 subgenome <-
   subgenome.chromosomes %>%

### 2017-08-11 R/SubGenome Analysis.Rmd
Provided support for a basic GO analysis
Cleaner data slices, parsed Citation field, but needs more cleanup, we have some unused and obsolete objects generated here.

Can do a basic GO analysis on the GO terms themselves, but not yet on their assignment to genes.

@@
 library(dplyr)
+library(VennDiagram)
 ```
@@
 ***
-In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size.
+In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size. (TODO)
 
 ```{r reportSubGenomes, echo=FALSE, results="asis"}
-# library(venneuler)
-# i <- intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
-# t <- setdiff(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
-# p <- setdiff(subgenome.truth, subgenome.chromosomes.test) %>% nrow()
-# vd <- venneuler(c(Paper=p, Test=t, "Paper&Test"=i))
-# vd$labels <- c(
-#   paste("Paper\n",p+i),
-#   paste("Test\n",t+i),
-#   paste("Both\n",i)
-# )
-# plot(vd)
+#Positions on each chr with sub 1 or sub 2
+#Total genes for each subgenome by chr and overall
 ```
 
-## Compared to the original paper.
+## Compared to Paper
 How many chromosomes with synteny to sorghum were correctly placed in subgenome 1 or subgenome 2?
@@
 
-```{r compareResultsToPaper, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=3, fig.align='center'}
-library(VennDiagram)
-# i <- intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
-# t <- setdiff(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
-# p <- setdiff(subgenome.truth, subgenome.chromosomes.test) %>% nrow()
+```{r compareResultsToPaper_chromosomes, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=3, fig.align='center'}
 subgenome.chromosomes.test <- 
@@
 
+grid.newpage()
 draw.pairwise.venn(
-  area1 = subgenome.truth%>%nrow(), 
-  area2 = subgenome.chromosomes.test%>%nrow(), 
+  area1 = subgenome.truth %>% nrow(), 
+  area2 = subgenome.chromosomes.test %>% nrow(), 
   cross.area = intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow(), 
@@
 
+Genes placed in the "correct" subgenome (TODO)
+
+```{r compareResultsToPaper_genes, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=3, fig.align='center'}
+#need to import "truth" and compare
+#visualize as venn
+```
+
 # GOTerms
 ***
+GO Term types by EVCode
 
+```{r GOTerms_by_EVCode, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+Type <- c("EV-EXP","EV-COMP")
+Count <- c(goAnnotations.mutated %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
+      goAnnotations.mutated %>% filter(grepl("EV-COMP", EVCode)) %>% nrow())
+df = data.frame(Type,Count)
+ggplot(data=df, aes(x=Type, y=Count)) +
+  geom_bar(stat="identity", fill="steelblue") +
+  geom_text(aes(label=Count), vjust=-0.3, size=3.5)
+```
+
+GO Term by subgenome (All EV-Codes)
+
+```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+# All EV-Codes, Unique GO Terms
+grid.newpage()
+draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% select("GO Term") %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% select("GO Term") %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% select("GO Term"), goAnnotations.sub2 %>% select("GO Term")) %>% nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0
+)
+
+#EXP only, Unique GO Terms
+grid.newpage()
+draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term"), 
+                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term")
+                         ) %>% nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0
+)
+```
+
+GOTerm assignments to genes by subgenome (All EV-Codes)
+(broken - doesn't take into account homeologous genes - try replacing equivalent genes from homeolog list)
+
+```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+# All EV-Codes, all assignments
+grid.newpage()
+draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% nrow(),
+  area2 = goAnnotations.sub2 %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% select(gene2, "GO Term"), goAnnotations.sub2 %>% select(gene2, "GO Term")) %>% nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0
+)
+
+#EXP only, all assignments
+grid.newpage()
+draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
+  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term"), 
+                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term")
+                         ) %>% nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0
+)
+```
+
+# Wrapup
 ***

### 2017-08-11 R/addGOTerms.R

@@ -17,6 +17,104 @@ goDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO 
 
 ## Read in GO Annotation data for maize genes
 goAnnotations <- read_delim(goDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
+goAnnotations.raw <- read_delim(goDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
+
+## Parse out citation string
+goAnnotations.mutated <-
+  goAnnotations.raw %>%
+  separate(Citation, c("Publication", "EVCode","TimeStamp","Curator"), sep=":", extra="drop")
+
+
+## GO Analysis -> grouping into useful slices
+goAnnotations.sub1 <-
+  subgenome %>%
+  subset(subgenome == "sub1") %>%
+  distinct() %>%
+  inner_join(goAnnotations.mutated, by = c("gene2" = "MaizeCyc2.2 Accession-1")) %>%
+  select(gene2, "GO Term", EVCode) %>%
+  distinct()
+
+goAnnotations.sub2 <-
+  subgenome %>%
+  subset(subgenome == "sub2") %>%
+  distinct() %>%
+  inner_join(goAnnotations.mutated, by = c("gene2" = "MaizeCyc2.2 Accession-1")) %>%
+  select(gene2, "GO Term", EVCode) %>%
+  distinct()
+
+goAnnotations.sub1.aggr <-
+  goAnnotations.sub1 %>%
+  select(`GO Term`) %>%
+  group_by(`GO Term`) %>%
+  summarise(n=n())
+
+goAnnotations.sub2.aggr <-
+  goAnnotations.sub2 %>%
+  select(`GO Term`) %>%
+  group_by(`GO Term`) %>%
+  summarise(n=n())
+
+#
+# setdiff(goAnnotations.sub1.aggr %>% select("GO Term"), goAnnotations.sub2.aggr %>% select("GO Term")) %>% nrow()
+# setdiff(goAnnotations.sub2.aggr %>% select("GO Term"), goAnnotations.sub1.aggr %>% select("GO Term"))
+#
+# full_join(goAnnotations.sub1.aggr, goAnnotations.sub2.aggr, by="GO Term", is.na=0) %>%
+#   rename(sub1_n=n.x, sub2_n=n.y) %>%
+#   subset(is.na(sub2_n)) %>%
+#   distinct() %>%
[…]

### 2017-08-11 R/addGOTerms.R
Minor
Back to Linux paths

@@ -13,7 +13,7 @@ library(dplyr)
 
 goDataFile <- params$goDataFile
 
-goDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO from maizecyc.tab"
+# goDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO from maizecyc.tab"
 
 ## Read in GO Annotation data for maize genes
 goAnnotations <- read_delim(goDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)

### 2017-08-11 R/parseSubgenomes.R

@@ -23,8 +23,8 @@ inputDataFile <- params$inputDataFile
 subGenomeFile <- params$subGenomeFile
 log10_ks_cutoff <- params$log10_ks_cutoff
 
-inputDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\sorghum_v1_vs_maize_v1.tab"
-subGenomeFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\deleteme\\outfile_processed_byHand.tab"
+# inputDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\sorghum_v1_vs_maize_v1.tab"
+# subGenomeFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\deleteme\\outfile_processed_byHand.tab"
 
 ## Import the raw data from the parsed SynMap output
 syntelogs.raw <- read_delim(inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)

### 2017-08-11 R/SubGenome Analysis.Rmd
Minor
Back to Linux paths
Manually fixed table
Moved chr10 for both a and b into last spot in table and gave the org_chr1 column a pretty name.

@@
 ```{r importSynMapData, include=FALSE}
-## source("~/git/SubGenomes/R/parseSubgenomes.R")
-## source("~/git/SubGenomes/R/addGOTerms.R")
+source("~/git/SubGenomes/R/parseSubgenomes.R")
+source("~/git/SubGenomes/R/addGOTerms.R")
 
-source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseSubgenomes.R")
-source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\addGOTerms.R")
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseSubgenomes.R")
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\addGOTerms.R")
 ```
@@
 # De-emphasize na's by turning them light grey
-homeologs.spread.formattedTable <- homeologs.spread
+homeologs.spread.formattedTable <- data.frame(rbind(homeologs.spread[-c(2),], homeologs.spread[c(2),]))
+homeologs.spread.formattedTable <- cbind(homeologs.spread.formattedTable[,-c(3, 12)], homeologs.spread.formattedTable[,c(3, 12)])
+homeologs.spread.formattedTable <- rename(homeologs.spread.formattedTable, Chromosome=org_chr1)
 homeologs.spread.formattedTable[is.na(homeologs.spread.formattedTable)] <- "<font color=\"lightgrey\">na</font>"
@@
 In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size. (TODO)
+
+foreach sorghum chromosome, set the largest syntenic chromosome (by gene) as sub1.
+Foreach additional syntenic chromosome if doesn't overlap what is already in sub1, add it to sub1.
+Else, if it does not overlap anything in sub2, add to sub2.
+Else toss.
+After all syntenic chromosomes are process for this sorghum chromosome, count the number of genes in sub1 and sub2.
+If sub2 is bigger, switch them.
+*Toss out "small" syntenic chromosomes (51 genes or less?), and consider a tolerance for how much overlap must occur to consider it a true overlap.
+Also consider switching to gene level overlap rather than whole-chromosome start/stop values.
 

### 2017-08-11 R/SubGenome Analysis.Rmd
Added titles to plots

@@
 library(VennDiagram)
+library(gridExtra)
 ```
@@
 ***
-In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size. (TODO)
+In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size. 
+**(TODO)**
 
@@
 Items in "Paper" only are missed subgenome assignments.
-Items in "Test" only are false assignments.
+Items in "Greedy" only are false assignments.
 Items in center are correct.
@@
 grid.newpage()
-draw.pairwise.venn(
+g = draw.pairwise.venn(
   area1 = subgenome.truth %>% nrow(), 
@@
   cross.area = intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow(), 
-  category = c("Paper", "Test"), 
+  category = c("Paper", "Greedy"), 
   lty = rep("blank", 2), 
@@
   sep.dist = 0.3, 
-  rotation.degree = 0
+  rotation.degree = 0,
+  ind = FALSE
 )
+grid.arrange(gTree(children=g), top="Overlap of Subgenome Assignments vs. \"Truth\"")
 ```
 
-Genes placed in the "correct" subgenome (TODO)
+Genes placed in the "correct" subgenome
+
+**(TODO)**
 
@@
 ***
-GO Term types by EVCode
 
-```{r GOTerms_by_EVCode, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+```{r GOTerms_by_EVCode, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4, fig.align='center'}
 Type <- c("EV-EXP","EV-COMP")
@@
   geom_bar(stat="identity", fill="steelblue") +
-  geom_text(aes(label=Count), vjust=-0.3, size=3.5)
+  geom_text(aes(label=Count), vjust=-0.3, size=3.5
+) + labs(
+  title="GOTerms by EVCode"
+)
 ```
 
-GO Term by subgenome (All EV-Codes)
+***
+
+GO Terms by subgenome.  If the GO Term is assigned to any gene in the subgenome, it is counted once for that subgenome.  This is a presence/absence test.
 
@@
 grid.newpage()
-draw.pairwise.venn(
+g <- draw.pairwise.venn(
   area1 = goAnnotations.sub1 %>% select("GO Term") %>% distinct() %>% nrow(),
@@
   sep.dist = 0.3, 
-  rotation.degree = 0
+  rotation.degree = 0,
+  ind = FALSE
 )
+grid.arrange(gTree(children=g), top="GO Term by subgenome (All EV-Codes)")
 
@@
 grid.newpage()
-draw.pairwise.venn(
+g <- draw.pairwise.venn(
   area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
@@
   sep.dist = 0.3, 
-  rotation.degree = 0
+  rotation.degree = 0,
+  ind = FALSE
 )
+grid.arrange(gTree(children=g), top="GO Term by subgenome (EV-EXP only)")
 ```
 
-GOTerm assignments to genes by subgenome (All EV-Codes)
+***
+
+GOTerm assignments to genes by subgenome, which takes into account the gene-GOTerm pairings and considers homeologs equivalent for the purposes of determining overlap in assignments.
+
 (broken - doesn't take into account homeologous genes - try replacing equivalent genes from homeolog list)
@@
 grid.newpage()
-draw.pairwise.venn(
+g <- draw.pairwise.venn(
   area1 = goAnnotations.sub1 %>% nrow(),
@@
   sep.dist = 0.3, 
-  rotation.degree = 0
+  rotation.degree = 0,
+  ind = FALSE
 )
+grid.arrange(gTree(children=g), top="GO Term assignments by subgenome (EV-EXP only)")
 
@@
 grid.newpage()
-draw.pairwise.venn(
+g <- draw.pairwise.venn(
   area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
@@
   sep.dist = 0.3, 
-  rotation.degree = 0
+  rotation.degree = 0,
+  ind = FALSE
 )
+grid.arrange(gTree(children=g), top="GO Term assignments by subgenome (EV-EXP only)")
 ```

### 2017-08-11 R/SubGenome Analysis.Rmd
Implemented a high level GO Term comparison by gene assignment
Accounts for homeolgous genes.

@@
 
-(broken - doesn't take into account homeologous genes - try replacing equivalent genes from homeolog list)
+```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2
+converted <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  right_join(., goAnnotations.sub2,by=c("Maize2"="gene2"))
+converted$Maize2[!is.na(converted$Maize1)] <- as.character(converted$Maize1[!is.na(converted$Maize1)])
+converted <- 
+  converted %>%
+  select(Maize2, "GO Term", EVCode) %>%
+  rename(gene2=Maize2)
 
-```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
 # All EV-Codes, all assignments
@@
   area1 = goAnnotations.sub1 %>% nrow(),
-  area2 = goAnnotations.sub2 %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% select(gene2, "GO Term"), goAnnotations.sub2 %>% select(gene2, "GO Term")) %>% nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term assignments by subgenome (EV-EXP only)")
-
-#EXP only, all assignments
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
-  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term"), 
-                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term")
+  area2 = converted %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% select(gene2, "GO Term") %>% distinct(), 
+                         converted %>% select(gene2, "GO Term") %>% distinct()
                          ) %>% nrow(),
@@
 )
-grid.arrange(gTree(children=g), top="GO Term assignments by subgenome (EV-EXP only)")
+grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(EV-EXP only)")
+
+#EXP only, all assignments
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
+  area2 = converted %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term") %>% distinct(), 
+                         converted %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term") %>% distinct()
+                         ) %>% nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(EV-EXP only)")
 ```
+
+<!-- ## GO Analysis -> grouping into useful slices -->
+<!-- goAnnotations.diff <- -->
+<!--   full_join(goAnnotations.sub1.aggr, goAnnotations.sub2.aggr, by="GO Term", is.na=0) %>% -->
+<!--   rename(sub1_n=n.x, sub2_n=n.y) %>% -->
+<!--   subset(is.na(sub1_n) | is.na(sub2_n)) -->
+<!-- goAnnotations.diff.counts <- -->
+<!--   full_join(goAnnotations.sub1.aggr, goAnnotations.sub2.aggr, by="GO Term", is.na=0) %>% -->
+<!--   rename(sub1_n=n.x, sub2_n=n.y) %>% -->
+<!--   replace_na(list(sub1_n=0, sub2_n=0)) %>% -->
+<!--   transmute(`GO Term`=`GO Term`,diff=sub1_n-sub2_n) %>% -->
+<!--   subset(diff != 0) -->
+<!-- goAnnotations.same <- -->
+<!--   full_join(goAnnotations.sub1.aggr, goAnnotations.sub2.aggr, by="GO Term") %>% -->
+<!--   rename(sub1_n=n.x, sub2_n=n.y) %>% -->
+<!--   replace_na(list(sub1_n=0, sub2_n=0)) %>% -->
+<!--   transmute(`GO Term`=`GO Term`,diff=sub1_n-sub2_n) %>% -->
+<!--   subset(diff == 0) -->
+
+<!-- print("Number of GO terms in sub1 not in sub2") -->
+<!-- print(nrow(goAnnotations.diff %>% subset(is.na(sub2_n)))) -->
+<!-- print("Number of GO terms in sub2 not in sub1") -->
+<!-- print(nrow(goAnnotations.diff %>% subset(is.na(sub1_n)))) -->
 

### 2017-08-11 R/SubGenome Analysis.Rmd
Added GO Analysis segment including GOSlim_Plant filter

@@
 ---
-title: "Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating subgenome assignments in maize"
-output: html_document
+title: 'Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating
+  subgenome assignments in maize'
 params:
-  fileParser: "../Perl/parseKsKnFile.pl"
-  goDataFile: "../Data/GO from maizecyc.tab"
-  subGenomeFile: "/home/jesse/Dropbox/deleteme/outfile_processed_byHand.tab"
-
-  #v1 vs. v1
-  synMapOutputFile: "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks"
-  inputDataFile: "../Data/sorghum_v1_vs_maize_v1.tab"
-  outputFile: "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab"
+  fileParser: ../Perl/parseKsKnFile.pl
+  goDataFile: ../Data/GO from maizecyc.tab
+  inputDataFile: ../Data/sorghum_v1_vs_maize_v1.tab
   log10_ks_cutoff: 0
-
-  #v3.1 vs v4+rejected
-  # synMapOutputFile: "31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords"
-  # inputDataFile: "../Data/sorghum_v3.1_vs_maize_v4+rejected.tab"
-  # outputFile: "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+rejected.tab"
-  # log10_ks_cutoff: 0
-  #-0.575
+  outputFile: ../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab
+  subGenomeFile: /home/jesse/Dropbox/deleteme/outfile_processed_byHand.tab
+  synMapOutputFile: 6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks
+output:
+  html_document: default
+  pdf_document: default
 ---
@@
 )
-grid.arrange(gTree(children=g), top="GO Term by subgenome (All EV-Codes)")
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes)")
 
@@
 )
-grid.arrange(gTree(children=g), top="GO Term by subgenome (EV-EXP only)")
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP only)")
 ```
@@
 )
-grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(EV-EXP only)")
+grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(All EV-Codes)")
 
@@
 grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(EV-EXP only)")
+```
+
+GO Terms by subgenome, only counting GO Terms that are in GOSlimPlant
+```{r GOTerms_by_subgenome_goslim_import, echo=FALSE}
+goSlim <- read_delim("../Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+goSlim$GOSlimTerm <- paste("|",goSlim$GOSlimTerm,"|",sep = "")
+```
+```{r GOTerms_by_subgenome_goslim, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+# All EV-Codes, Unique GO Terms
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term") %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term") %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term"), 
+                         goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term")) %>% 
+              nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim only)")
+
+#EXP only, Unique GO Terms
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm"))%>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term"), 
+                         goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term")
+                         ) %>% nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP only, GOSlim only)")
 ```

### 2017-08-11 Perl/parseGOSlimFile.pl

@@ -1,0 +1,9 @@
+use strict;
+use warnings;
+
+while(<>) {
+  if ($_ =~ /id: GO:.*$/) {
+    my @words = split / /, $_;
+    print $words[1];
+  }
+}

### 2017-08-11 R/parseGOSlim.R

@@ -1,0 +1,12 @@
+goSlim <- read_delim("./Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+goSlim$GOSlimTerm <- paste("|",goSlim$GOSlimTerm,"|",sep = "")
+
+
+
+goAssignmentsIn2not1_withSlim <-
+  converted %>%
+  inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>%
+  filter(grepl("EV-EXP", EVCode)) %>%
+  setdiff(., goAnnotations.sub1 %>%
+              inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>%
+              filter(grepl("EV-EXP", EVCode)))

### 2017-08-17 R/SubGenome Analysis.Rmd
Spelling

@@
 
-# GOTerms
+# GO Terms
 ***
@@
 ) + labs(
-  title="GOTerms by EVCode"
+  title="GO Terms by EVCode"
 )
@@
 
-GOTerm assignments to genes by subgenome, which takes into account the gene-GOTerm pairings and considers homeologs equivalent for the purposes of determining overlap in assignments.
+GO Term assignments to genes by subgenome, which takes into account the gene-GO Term pairings and considers homeologs equivalent for the purposes of determining overlap in assignments.
 

### 2017-08-18 R/SubGenome Analysis.Rmd
Minor

@@
   outputFile: ../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab
-  subGenomeFile: /home/jesse/Dropbox/deleteme/outfile_processed_byHand.tab
+  subGenomeFile: ../Data/outfile_processed_byHand.tab
   synMapOutputFile: 6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks
 output:
+  pdf_document: default
   html_document: default
-  pdf_document: default
 ---

### 2017-08-18 R/parseGOSlim.R

@@ -1,3 +1,5 @@
+# http://www.geneontology.org/ontology/subsets/goslim_plant.obo
+
 goSlim <- read_delim("./Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
 goSlim$GOSlimTerm <- paste("|",goSlim$GOSlimTerm,"|",sep = "")
 

### 2017-08-18 R/SubGenome Analysis.Rmd
Don't have bars (|) in go terms

@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 goSlim <- read_delim("../Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-goSlim$GOSlimTerm <- paste("|",goSlim$GOSlimTerm,"|",sep = "")
+# goSlim$GOSlimTerm <- paste("|",goSlim$GOSlimTerm,"|",sep = "")
 ```

### 2017-08-18 R/addGOTerms.R

@@ -23,6 +23,8 @@ goAnnotations.mutated <-
   goAnnotations.raw %>%
   separate(Citation, c("Publication", "EVCode","TimeStamp","Curator"), sep=":", extra="drop")
 
+## Remove |'s from GO Terms
+goAnnotations.mutated$`GO Term` <- gsub(goAnnotations.mutated$`GO Term`, pattern = "\\|", replacement = "")
 
 ## GO Analysis -> grouping into useful slices
 goAnnotations.sub1 <-

### 2017-08-18 R/parseGOSlim.R

@@ -1,7 +1,7 @@
 # http://www.geneontology.org/ontology/subsets/goslim_plant.obo
 
 goSlim <- read_delim("./Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-goSlim$GOSlimTerm <- paste("|",goSlim$GOSlimTerm,"|",sep = "")
+# goSlim$GOSlimTerm <- paste("|",goSlim$GOSlimTerm,"|",sep = "")
 
 
 

### 2017-08-18 R/SubGenome Analysis.Rmd
Tied GOSlim import to main report

@@
 source("~/git/SubGenomes/R/addGOTerms.R")
+source("~/git/SubGenomes/R/parseGOSlim.R")
 
@@
 GO Terms by subgenome, only counting GO Terms that are in GOSlimPlant
-```{r GOTerms_by_subgenome_goslim_import, echo=FALSE}
-goSlim <- read_delim("../Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-# goSlim$GOSlimTerm <- paste("|",goSlim$GOSlimTerm,"|",sep = "")
-```
+
 ```{r GOTerms_by_subgenome_goslim, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}

### 2017-08-18 R/parseGOSlim.R

@@ -1,14 +1,14 @@
 # http://www.geneontology.org/ontology/subsets/goslim_plant.obo
 
-goSlim <- read_delim("./Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+goSlim <- read_delim("../Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
 # goSlim$GOSlimTerm <- paste("|",goSlim$GOSlimTerm,"|",sep = "")
 
 
 
-goAssignmentsIn2not1_withSlim <-
-  converted %>%
-  inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>%
-  filter(grepl("EV-EXP", EVCode)) %>%
-  setdiff(., goAnnotations.sub1 %>%
-              inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>%
-              filter(grepl("EV-EXP", EVCode)))
+# goAssignmentsIn2not1_withSlim <-
+#   converted %>%
+#   inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>%
+#   filter(grepl("EV-EXP", EVCode)) %>%
+#   setdiff(., goAnnotations.sub1 %>%
+#               inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>%
+#               filter(grepl("EV-EXP", EVCode)))

### 2017-08-18 R/SubGenome Analysis.Rmd
Minor

@@
   goDataFile: ../Data/GO from maizecyc.tab
+  log10_ks_cutoff: 0
+  synMapOutputFile: 6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks
   inputDataFile: ../Data/sorghum_v1_vs_maize_v1.tab
-  log10_ks_cutoff: 0
-  outputFile: ../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab
   subGenomeFile: ../Data/outfile_processed_byHand.tab
-  synMapOutputFile: 6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks
+
+  # inputDataFile: ../Data/sorghum_v3.1_vs_maize_v4+rejected.tab
+  # subGenomeFile: ../Data/outfile_processed_byHand.tab
 output:
@@
 
+```{r aggregates, include=FALSE}
+goAnnotations.sub1.aggr <-
+  goAnnotations.sub1 %>%
+  select(`GO Term`) %>%
+  group_by(`GO Term`) %>%
+  summarise(n=n())
+
+goAnnotations.sub2.aggr <-
+  goAnnotations.sub2 %>%
+  select(`GO Term`) %>%
+  group_by(`GO Term`) %>%
+  summarise(n=n())
+```
+
 <!-- ## GO Analysis -> grouping into useful slices -->

### 2017-08-18 R/addGOTerms.R

@@ -42,29 +42,3 @@ goAnnotations.sub2 <-
   inner_join(goAnnotations.mutated, by = c("gene2" = "MaizeCyc2.2 Accession-1")) %>%
   select(gene2, "GO Term", EVCode) %>%
   distinct()
-
-goAnnotations.sub1.aggr <-
-  goAnnotations.sub1 %>%
-  select(`GO Term`) %>%
-  group_by(`GO Term`) %>%
-  summarise(n=n())
-
-goAnnotations.sub2.aggr <-
-  goAnnotations.sub2 %>%
-  select(`GO Term`) %>%
-  group_by(`GO Term`) %>%
-  summarise(n=n())
-
-
-## A simple list of genes in subgenome 1
-subgenome.sub1 <-
-  subgenome %>%
-  filter(subgenome == "sub1") %>%
-  select(gene2) %>%
-  distinct()
-## A simple list of genes in subgenome 2
-subgenome.sub2 <-
-  subgenome %>%
-  filter(subgenome == "sub2") %>%
-  select(gene2) %>%
-  distinct()

### 2017-08-18 R/SubGenome Analysis.Rmd
Updated analysis to use the v4 SynMap data

@@
   synMapOutputFile: 6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks
-  inputDataFile: ../Data/sorghum_v1_vs_maize_v1.tab
+  # inputDataFile: ../Data/sorghum_v1_vs_maize_v1.tab
+  # subGenomeFile: ../Data/outfile_processed_byHand.tab
+
+  inputDataFile: ../Data/sorghum_v3.1_vs_maize_v4+rejected.tab
   subGenomeFile: ../Data/outfile_processed_byHand.tab
-
-  # inputDataFile: ../Data/sorghum_v3.1_vs_maize_v4+rejected.tab
-  # subGenomeFile: ../Data/outfile_processed_byHand.tab
 output:

### 2017-08-18 R/addGOTerms.R

@@ -23,6 +23,12 @@ goAnnotations.mutated <-
   goAnnotations.raw %>%
   separate(Citation, c("Publication", "EVCode","TimeStamp","Curator"), sep=":", extra="drop")
 
+## Rename column (for v2 id's use MaizeCyc2.2 Accession-1, for v4 use V4_ID)
+goAnnotations.mutated <-
+  goAnnotations.mutated %>%
+  rename("geneID" = "V4_ID")
+  #rename("geneID" = "MaizeCyc2.2 Accession-1")
+
 ## Remove |'s from GO Terms
 goAnnotations.mutated$`GO Term` <- gsub(goAnnotations.mutated$`GO Term`, pattern = "\\|", replacement = "")
 
@@ -31,7 +37,7 @@ goAnnotations.sub1 <-
   subgenome %>%
   subset(subgenome == "sub1") %>%
   distinct() %>%
-  inner_join(goAnnotations.mutated, by = c("gene2" = "MaizeCyc2.2 Accession-1")) %>%
+  inner_join(goAnnotations.mutated, by = c("gene2" = "geneID")) %>%
   select(gene2, "GO Term", EVCode) %>%
   distinct()
 
@@ -39,6 +45,6 @@ goAnnotations.sub2 <-
   subgenome %>%
   subset(subgenome == "sub2") %>%
   distinct() %>%
-  inner_join(goAnnotations.mutated, by = c("gene2" = "MaizeCyc2.2 Accession-1")) %>%
+  inner_join(goAnnotations.mutated, by = c("gene2" = "geneID")) %>%
   select(gene2, "GO Term", EVCode) %>%
   distinct()

### 2017-08-18 R/parseSubgenomes.R

@@ -29,6 +29,10 @@ log10_ks_cutoff <- params$log10_ks_cutoff
 ## Import the raw data from the parsed SynMap output
 syntelogs.raw <- read_delim(inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
 
+## For v4 imports, need to remove the CDS: and _T00# parts of the gene2 column
+syntelogs.raw$gene2 <- gsub(syntelogs.raw$gene2, pattern = "CDS:", replacement = "")
+syntelogs.raw$gene2 <- gsub(syntelogs.raw$gene2, pattern = "_T\\d\\d\\d", replacement = "")
+
 ## Import "True" sugenomes as reported by Schnable 2011
 subgenome.truth <- setNames(data.frame(
   c(1,1,1,2,2,3,3,4,4,5,5,6,6,7,7,7,7,8,8,8,9,9,9,10,10,10),

### 2017-08-21 R/SubGenome Analysis.Rmd
Added venn filtered by MF terms

@@
   log10_ks_cutoff: 0
-  synMapOutputFile: 6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks
+  # synMapOutputFile: 6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks
   # inputDataFile: ../Data/sorghum_v1_vs_maize_v1.tab
@@
 knitr::opts_chunk$set(echo = TRUE)
+library(topGO)
 library(readr)
@@
 source("~/git/SubGenomes/R/parseGOSlim.R")
+# source("~/git/SubGenomes/R/createTopGO.R")
+
+MFterms <- ls(GOMFTerm)
 
@@
 
+GO Terms by subgenome, only counting GO Terms that are in GOSlimPlant and are Molecular Function terms.
+
+```{r GOTerms_by_subgenome_goslim_MF, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+# All EV-Codes, Unique GO Terms
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term") %>% 
+    filter(`GO Term` %in% MFterms) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term") %>% 
+    filter(`GO Term` %in% MFterms) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term") %>% filter(`GO Term` %in% MFterms), 
+                         goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term") %>% filter(`GO Term` %in% MFterms)) %>%
+                          nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim, MF only)")
+
+#EXP only, Unique GO Terms
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% 
+    filter(`GO Term` %in% MFterms) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% 
+    filter(`GO Term` %in% MFterms) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm"))%>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% 
+                           filter(`GO Term` %in% MFterms), 
+                         goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% 
+                           filter(`GO Term` %in% MFterms)
+                         ) %>% nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim, MF only)")
+```
+
 ```{r aggregates, include=FALSE}

### 2017-08-25 R/createTopGO.R
Minor

@@ -5,11 +5,16 @@ library(tidyr)
 ## Project:
 ## Script purpose:
 ##
-## Input:
+## Input: goAnnotations.mutated from the addGOTerms.R script
 ## Output:
+##
 ## Date: 2017-08-18
 ## Author: Jesse R. Walsh
 ####################################################################################################
+## GOBPTerm, GOMFTerm and GOCCTerm are part of the topGO package
+MFterms <- ls(GOMFTerm)
+# head(MFterms)
+
 
 # goDataFile <- params$goDataFile
 # goDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO from maizecyc.tab"
@@ -46,9 +51,7 @@ goAnnotations.sub1.temp$`GO Term` <- gsub(goAnnotations.sub1.temp$`GO Term`, pat
 goAnnotations.sub1.temp %>%
   subset(`GO Term` %in% MFterms)
 
-## GOBPTerm, GOMFTerm and GOCCTerm are part of the topGO package
-# MFterms <- ls(GOMFTerm)
-# head(MFterms)
+
 
 # geneID2GO.temp$GO <- as.character(strsplit(geneID2GO.temp$GO, ","))
 # geneID2GO <- as.character(geneID2GO.temp$GO)

### 2017-08-25 R/SubGenome Analysis.Rmd
Added some analysis for expression data

@@
   goDataFile: ../Data/GO from maizecyc.tab
+  inputDataFile: ../Data/sorghum_v3.1_vs_maize_v4+rejected.tab
   log10_ks_cutoff: 0
-  # synMapOutputFile: 6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks
-  # inputDataFile: ../Data/sorghum_v1_vs_maize_v1.tab
-  # subGenomeFile: ../Data/outfile_processed_byHand.tab
-
-  inputDataFile: ../Data/sorghum_v3.1_vs_maize_v4+rejected.tab
   subGenomeFile: ../Data/outfile_processed_byHand.tab
@@
 source("~/git/SubGenomes/R/parseGOSlim.R")
-# source("~/git/SubGenomes/R/createTopGO.R")
-
+source("~/git/SubGenomes/R/createTopGO.R")
+source("~/git/SubGenomes/R/parseExpressionData.R")
 MFterms <- ls(GOMFTerm)
@@
      + **`r subgenome %>% filter(subgenome == "sub1") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 1
+     + **`r subgenome %>% filter(subgenome == "sub1") %>% filter(gene2 %in% expressedGenes$geneID) %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned            to subgenome 1 that are expressed significantly on average over 68 samples
      + **`r goAnnotations.sub1 %>% nrow()`** GO assignments using **`r goAnnotations.sub1 %>% select('GO Term') %>% distinct() %>% nrow()`** GO Terms
@@
      + **`r subgenome %>% filter(subgenome == "sub2") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 2
+     + **`r subgenome %>% filter(subgenome == "sub2") %>% filter(gene2 %in% expressedGenes$geneID) %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned            to subgenome 2 that are expressed significantly on average over 68 samples
      + **`r goAnnotations.sub2 %>% nrow()`** GO assignments using **`r goAnnotations.sub2 %>% select('GO Term') %>% distinct() %>% nrow()`** GO Terms
@@
 ```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
-## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2
+## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2.  The converted data.frame will hold maize2 genes with the names replaced with maize1 homologs
 converted <-
@@
 
+# Expression
+
+Cases where the maize1 homolog is more or less expressed than the maize2 homolog
+
+```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+## Attach log2(FPKM_mean) values to homeologous pairs
+expressedPairs <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  inner_join(expressedGenes, by=c("Maize1"="geneID")) %>%
+  rename(FPKM_mean1=FPKM_mean) %>%
+  inner_join(expressedGenes, by=c("Maize2"="geneID")) %>%
+  rename(FPKM_mean2=FPKM_mean)
+
+# Expression Diffs
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2) %>% nrow(),
+  area2 = expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2) %>% nrow(),
+  cross.area = intersect(expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2),
+                         expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2)) %>%
+                          nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="Genes with 2-fold higher\nexpression by subgenome")
+```
+
 # Wrapup

### 2017-08-31 R/cleanData.R
Massive renaming and reorganizing
Clear load and clean scripts.  Processing left in parse* scripts. Naming following clearer conventions and more descriptive names, including both objects and column names.

@@ -1,0 +1,9 @@
+####################################################################################################
+## Project:
+## Script purpose:
+##
+## Input:
+## Output:
+## Date: 2017-08-25
+## Author: Jesse R. Walsh
+####################################################################################################

### 2017-08-31 R/createTopGO.R

@@ -28,10 +28,10 @@ MFterms <- ls(GOMFTerm)
 
 ## Create a topGO custom geneID2GO annotation object, mapping gene ids to go terms
 geneID2GO.temp <-
-  goAnnotations.mutated %>%
-  select(geneID, `GO Term`) %>%
+  go.maize.clean %>%
+  select(geneID, goTerm) %>%
   group_by(geneID) %>%
-  summarise(GO = paste(`GO Term`, collapse = ", "))
+  summarise(GO = paste(goTerm, collapse = ", "))
 geneID2GO <- strsplit(geneID2GO.temp$GO, ", ")
 names(geneID2GO) <- geneID2GO.temp$geneID
 

### 2017-08-31 R/loadData.R

@@ -1,0 +1,30 @@
+library(readr)
+####################################################################################################
+## Project:
+## Script purpose:
+##
+## Input:
+## Output:
+## Date: 2017-08-25
+## Author: Jesse R. Walsh
+####################################################################################################
+
+## Import the raw data from the parsed SynMap output
+syntelogs.v1.raw <- read_delim("./Data/sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+syntelogs.v4.raw <- read_delim("./Data/sorghum_v3.1_vs_maize_v4+rejected.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+
+subgenome.chromosomes <- read_delim("./Data/outfile_processed_byHand.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+
+## Read in GO Annotation data for maize genes
+goAnnotations.raw <- read_delim("./Data/GO from maizecyc.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+
+sorghum.go <- read_delim("./Data/gramene_sorghumv2_goterms.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
+
+goSlim <- read_delim("./Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+
+GSE50191_FPKM <- read_delim("./Data/GSE50191_FPKM.tsv", "\t", trim_ws = TRUE)
+
+
+
+#--------------------------------------------------------------------------------------------------#
+detach(readr)

### 2017-08-31 R/parseExpressionData.R

@@ -1,11 +1,11 @@
 library(readr)
-GSE50191_FPKM <- read_delim("~/git/SubGenomes/Data/GSE50191_FPKM.tsv",
-                            "\t",
-                            trim_ws = TRUE)
-
-## Remove low FPKM values, calculate row means
-GSE50191_FPKM[GSE50191_FPKM < 1] <- NA
-expressedGenes <- data.frame(ID=GSE50191_FPKM[,1], Means=rowMeans(GSE50191_FPKM[,-1], na.rm = TRUE))
+# GSE50191_FPKM <- read_delim("~/git/SubGenomes/Data/GSE50191_FPKM.tsv",
+#                             "\t",
+#                             trim_ws = TRUE)
+#
+# ## Remove low FPKM values, calculate row means
+# GSE50191_FPKM[GSE50191_FPKM < 1] <- NA
+expressedGenes <- data.frame(ID=maize.expression.clean[,1], Means=rowMeans(maize.expression.clean[,-1], na.rm = TRUE))
 
 ## Remove rows with NA
 expressedGenes <-
@@ -17,9 +17,9 @@ expressedGenes <-
 
 ## Convert to v4 ids (this is a bad conversion table, only has CornCyc Genes in it.... get a better one!)
 geneIDConvertTable <-
-  goAnnotations.mutated %>%
-  select(geneID, `MaizeCyc2.2 Accession-1`) %>%
-  rename(v4ID=geneID, v3ID=`MaizeCyc2.2 Accession-1`) %>%
+  go.maize.raw %>%
+  select(V4_ID, `MaizeCyc2.2 Accession-1`) %>%
+  rename(v4ID=V4_ID, v3ID=`MaizeCyc2.2 Accession-1`) %>%
   distinct()
 
 expressedGenes <-

### 2017-08-31 R/parseGOSlim.R

@@ -11,7 +11,7 @@ library(readr)
 ## DataSource: http://www.geneontology.org/ontology/subsets/goslim_plant.obo
 ####################################################################################################
 
-goSlim <- read_delim("../Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+# goSlim <- read_delim("../Data/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
 
 
 # goSlim$GOSlimTerm <- paste("|",goSlim$GOSlimTerm,"|",sep = "")

### 2017-08-31 R/parseSorghumGO.R

@@ -1,22 +1,22 @@
-sorghum.go <- read_delim("./Data/gramene_sorghumv2_goterms.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
-
-sorghum.go <-
-  sorghum.go %>%
-  rename(sorghumID=`Gene stable ID`, goTerm=`GO term accession`) %>%
-  select(sorghumID, goTerm)
-
-sorghum.go$sorghumID <- gsub(sorghum.go$sorghumID, pattern = "SORBI_", replacement = "Sobic.")
+# sorghum.go <- read_delim("./Data/gramene_sorghumv2_goterms.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
+#
+# sorghum.go <-
+#   sorghum.go %>%
+#   rename(sorghumID=`Gene stable ID`, goTerm=`GO term accession`) %>%
+#   select(sorghumID, goTerm)
+#
+# sorghum.go$sorghumID <- gsub(sorghum.go$sorghumID, pattern = "SORBI_", replacement = "Sobic.")
 # sorghum.go$sorghumID <- gsub(sorghum.go$sorghumID, pattern = "$", replacement = ".1.v3.1")
 
 maizeWithSorghumGO <-
   syntelogs.mutated %>%
   select(gene1, gene2)
 
-maizeWithSorghumGO$gene1 <- gsub(maizeWithSorghumGO$gene1, pattern = ".\\d.v\\d.\\d", replacement = "")
+# maizeWithSorghumGO$gene1 <- gsub(maizeWithSorghumGO$gene1, pattern = ".\\d.v\\d.\\d", replacement = "")
 
 maizeWithSorghumGO <-
   maizeWithSorghumGO %>%
-  left_join(sorghum.go, by=c("gene1"="sorghumID"))
+  left_join(go.sorghum.clean, by=c("gene1"="sorghumID"))
 
 maizeWithSorghumGO <-
   maizeWithSorghumGO %>%

### 2017-08-31 SubGenome Analysis.Rmd
Massive renaming and reorganizing
Clear load and clean scripts.  Processing left in parse* scripts. Naming following clearer conventions and more descriptive names, including both objects and column names.

--- 2017-08-18 R/SubGenome Analysis.Rmd
@@
 params:
-  fileParser: ../Perl/parseKsKnFile.pl
-  goDataFile: ../Data/GO from maizecyc.tab
   log10_ks_cutoff: 0
-  synMapOutputFile: 6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks
-  # inputDataFile: ../Data/sorghum_v1_vs_maize_v1.tab
-  # subGenomeFile: ../Data/outfile_processed_byHand.tab
-
-  inputDataFile: ../Data/sorghum_v3.1_vs_maize_v4+rejected.tab
-  subGenomeFile: ../Data/outfile_processed_byHand.tab
 output:
@@
 ```
-
-This analysis is based on the file: **`r params$inputDataFile`**
 
@@
 ```{r importSynMapData, include=FALSE}
+source("~/git/SubGenomes/R/loadData.R")
+source("~/git/SubGenomes/R/cleanData.R")
 source("~/git/SubGenomes/R/parseSubgenomes.R")
@@
 source("~/git/SubGenomes/R/parseGOSlim.R")
+# source("~/git/SubGenomes/R/createTopGO.R")
+source("~/git/SubGenomes/R/parseExpressionData.R")
+
+MFterms <- ls(GOMFTerm)
 
@@
 ***  
- * There were **`r syntelogs.raw %>% nrow()`** rows of data in the SynMap file.
- * There were **`r syntelogs.raw %>% select(gene2) %>% distinct() %>% nrow()`** distinct Maize genes with synteny to Sorghum.
+ * There were **`r syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean %>% nrow()`** rows of data in the SynMap file.
+ * There were **`r syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean %>% select(gene2) %>% distinct() %>% nrow()`** distinct Maize genes with synteny to Sorghum.
  * Summary of the KS values
-     + **`r syntelogs.raw %>% select(ks) %>% summary()`**
- * Out of the `r syntelogs.raw %>% select(gene2) %>% distinct() %>% nrow()` distinct Maize genes with synteny to sorghum, we keep **`r homeologs.genes %>% select(gene2) %>% distinct() %>% nrow()`** after applying the KS cutoff.
+     + **`r syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean %>% select(ks) %>% summary()`**
+ * Out of the `r syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean %>% select(gene2) %>% distinct() %>% nrow()` distinct Maize genes with synteny to sorghum, we keep **`r homeologs.genes %>% select(gene2) %>% distinct() %>% nrow()`** after applying the KS cutoff.
  * SubGenome 1
      + **`r subgenome %>% filter(subgenome == "sub1") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 1
-     + **`r goAnnotations.sub1 %>% nrow()`** GO assignments using **`r goAnnotations.sub1 %>% select('GO Term') %>% distinct() %>% nrow()`** GO Terms
+     + **`r subgenome %>% filter(subgenome == "sub1") %>% filter(gene2 %in% expressedGenes$geneID) %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned            to subgenome 1 that are expressed significantly on average over 68 samples
+     + **`r goAnnotations.sub1 %>% nrow()`** GO assignments using **`r goAnnotations.sub1 %>% select(goTerm) %>% distinct() %>% nrow()`** GO Terms
  * SubGenome 2
      + **`r subgenome %>% filter(subgenome == "sub2") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 2
-     + **`r goAnnotations.sub2 %>% nrow()`** GO assignments using **`r goAnnotations.sub2 %>% select('GO Term') %>% distinct() %>% nrow()`** GO Terms
+     + **`r subgenome %>% filter(subgenome == "sub2") %>% filter(gene2 %in% expressedGenes$geneID) %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned            to subgenome 2 that are expressed significantly on average over 68 samples
+     + **`r goAnnotations.sub2 %>% nrow()`** GO assignments using **`r goAnnotations.sub2 %>% select(goTerm) %>% distinct() %>% nrow()`** GO Terms
  * Rejected
@@
 qplot(
-  log10(syntelogs.raw$ks), 
+  log10(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$ks), 
   geom="histogram", 
@@
 subgenome.chromosomes.test <- 
-  subgenome.chromosomes %>% 
+  subgenome.assignments %>% 
   mutate(chr1=as.numeric(regmatches(org_chr1, regexpr("\\d*$",org_chr1)))) %>%
@@
 Type <- c("EV-EXP","EV-COMP")
-Count <- c(goAnnotations.mutated %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
-      goAnnotations.mutated %>% filter(grepl("EV-COMP", EVCode)) %>% nrow())
+Count <- c(go.maize.clean %>% filter(grepl("EV-EXP", evCode)) %>% nrow(),
+      go.maize.clean %>% filter(grepl("EV-COMP", evCode)) %>% nrow())
 df = data.frame(Type,Count)
@@
 g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% select("GO Term") %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% select("GO Term") %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% select("GO Term"), goAnnotations.sub2 %>% select("GO Term")) %>% nrow(), 
+  area1 = goAnnotations.sub1 %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% select(goTerm), goAnnotations.sub2 %>% select(goTerm)) %>% nrow(), 
   category = c("SubGenome 1", "SubGenome 2"), 
@@
 g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term"), 
-                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term")
+  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm), 
+                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm)
                          ) %>% nrow(), 
@@
   converted %>%
-  select(Maize2, "GO Term", EVCode) %>%
+  select(Maize2, goTerm, evCode) %>%
   rename(gene2=Maize2)
@@
   area2 = converted %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% select(gene2, "GO Term") %>% distinct(), 
-                         converted %>% select(gene2, "GO Term") %>% distinct()
+  cross.area = intersect(goAnnotations.sub1 %>% select(gene2, goTerm) %>% distinct(), 
+                         converted %>% select(gene2, goTerm) %>% distinct()
                          ) %>% nrow(),
@@
 g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
-  area2 = converted %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term") %>% distinct(), 
-                         converted %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term") %>% distinct()
+  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", evCode)) %>% nrow(),
+  area2 = converted %>% filter(grepl("EV-EXP", evCode)) %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", evCode)) %>% select(gene2, goTerm) %>% distinct(), 
+                         converted %>% filter(grepl("EV-EXP", evCode)) %>% select(gene2, goTerm) %>% distinct()
                          ) %>% nrow(), 
@@
 g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term") %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term") %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term"), 
-                         goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% select("GO Term")) %>% 
+  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm), 
+                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm)) %>% 
               nrow(), 
@@
 g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm"))%>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term"), 
-                         goAnnotations.sub2 %>% inner_join(goSlim, by=c("GO Term" = "GOSlimTerm")) %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term")
+  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm), 
+                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm)
                          ) %>% nrow(), 
@@
 
+GO Terms by subgenome, only counting GO Terms that are in GOSlimPlant and are Molecular Function terms.
+
+```{r GOTerms_by_subgenome_goslim_MF, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+# All EV-Codes, Unique GO Terms
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% 
+    filter(goTerm %in% MFterms) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% 
+    filter(goTerm %in% MFterms) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% filter(goTerm %in% MFterms), 
+                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% filter(goTerm %in% MFterms)) %>%
+                          nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim, MF only)")
+
+#EXP only, Unique GO Terms
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% 
+    filter(goTerm %in% MFterms) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% 
+    filter(goTerm %in% MFterms) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% 
+                           filter(goTerm %in% MFterms), 
+                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% 
+                           filter(goTerm %in% MFterms)
+                         ) %>% nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim, MF only)")
+```
+
 ```{r aggregates, include=FALSE}
@@
   goAnnotations.sub1 %>%
-  select(`GO Term`) %>%
-  group_by(`GO Term`) %>%
+  select(goTerm) %>%
+  group_by(goTerm) %>%
   summarise(n=n())
@@
   goAnnotations.sub2 %>%
-  select(`GO Term`) %>%
-  group_by(`GO Term`) %>%
+  select(goTerm) %>%
+  group_by(goTerm) %>%
   summarise(n=n())
@@
 
-<!-- ## GO Analysis -> grouping into useful slices -->
-<!-- goAnnotations.diff <- -->
-<!--   full_join(goAnnotations.sub1.aggr, goAnnotations.sub2.aggr, by="GO Term", is.na=0) %>% -->
-<!--   rename(sub1_n=n.x, sub2_n=n.y) %>% -->
-<!--   subset(is.na(sub1_n) | is.na(sub2_n)) -->
-<!-- goAnnotations.diff.counts <- -->
-<!--   full_join(goAnnotations.sub1.aggr, goAnnotations.sub2.aggr, by="GO Term", is.na=0) %>% -->
-<!--   rename(sub1_n=n.x, sub2_n=n.y) %>% -->
-<!--   replace_na(list(sub1_n=0, sub2_n=0)) %>% -->
-<!--   transmute(`GO Term`=`GO Term`,diff=sub1_n-sub2_n) %>% -->
-<!--   subset(diff != 0) -->
-<!-- goAnnotations.same <- -->
-<!--   full_join(goAnnotations.sub1.aggr, goAnnotations.sub2.aggr, by="GO Term") %>% -->
-<!--   rename(sub1_n=n.x, sub2_n=n.y) %>% -->
-<!--   replace_na(list(sub1_n=0, sub2_n=0)) %>% -->
-<!--   transmute(`GO Term`=`GO Term`,diff=sub1_n-sub2_n) %>% -->
-<!--   subset(diff == 0) -->
+# Expression
 
-<!-- print("Number of GO terms in sub1 not in sub2") -->
-<!-- print(nrow(goAnnotations.diff %>% subset(is.na(sub2_n)))) -->
-<!-- print("Number of GO terms in sub2 not in sub1") -->
-<!-- print(nrow(goAnnotations.diff %>% subset(is.na(sub1_n)))) -->
+Cases where the maize1 homolog is more or less expressed than the maize2 homolog
+
+```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+## Attach log2(FPKM_mean) values to homeologous pairs
+expressedPairs <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  inner_join(expressedGenes, by=c("Maize1"="geneID")) %>%
+  rename(FPKM_mean1=FPKM_mean) %>%
+  inner_join(expressedGenes, by=c("Maize2"="geneID")) %>%
+  rename(FPKM_mean2=FPKM_mean)
+
+# Expression Diffs
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2) %>% nrow(),
+  area2 = expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2) %>% nrow(),
+  cross.area = intersect(expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2),
+                         expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2)) %>%
+                          nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="Genes with 2-fold higher\nexpression by subgenome")
+```
 

### 2017-08-31 R/addGOTerms.R

@@ -11,40 +11,40 @@ library(dplyr)
 ## Author: Jesse R. Walsh
 ####################################################################################################
 
-goDataFile <- params$goDataFile
-
-# goDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO from maizecyc.tab"
-
-## Read in GO Annotation data for maize genes
-goAnnotations.raw <- read_delim(goDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
-
-## Parse out citation string
-goAnnotations.mutated <-
-  goAnnotations.raw %>%
-  separate(Citation, c("Publication", "EVCode","TimeStamp","Curator"), sep=":", extra="drop")
-
-## Rename column (for v2 id's use MaizeCyc2.2 Accession-1, for v4 use V4_ID)
-goAnnotations.mutated <-
-  goAnnotations.mutated %>%
-  rename("geneID" = "V4_ID")
-  #rename("geneID" = "MaizeCyc2.2 Accession-1")
-
-## Remove |'s from GO Terms
-goAnnotations.mutated$`GO Term` <- gsub(goAnnotations.mutated$`GO Term`, pattern = "\\|", replacement = "")
+# goDataFile <- params$goDataFile
+#
+# # goDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO from maizecyc.tab"
+#
+# ## Read in GO Annotation data for maize genes
+# goAnnotations.raw <- read_delim(goDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
+#
+# ## Parse out citation string
+# goAnnotations.mutated <-
+#   goAnnotations.raw %>%
+#   separate(Citation, c("Publication", "EVCode","TimeStamp","Curator"), sep=":", extra="drop")
+#
+# ## Rename column (for v2 id's use MaizeCyc2.2 Accession-1, for v4 use V4_ID)
+# goAnnotations.mutated <-
+#   goAnnotations.mutated %>%
+#   rename("geneID" = "V4_ID")
+#   #rename("geneID" = "MaizeCyc2.2 Accession-1")
+#
+# ## Remove |'s from GO Terms
+# goAnnotations.mutated$`GO Term` <- gsub(goAnnotations.mutated$`GO Term`, pattern = "\\|", replacement = "")
 
[…]

### 2017-08-31 R/cleanData.R

@@ -1,0 +1,84 @@
+library(tidyr)
+library(dplyr)
+####################################################################################################
+## Project:         Subgenomes project
+## Script purpose:  Clean up raw datafiles needed for this project, including renaming column and
+##        selecting relevant columns, handling missing data and incomplete rows, and any global
+##        conversions or modification to data to make it usable.
+##
+## Input:
+##        syntelogs.sorghum.v1.maize.v1.raw
+##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw
+##        maize.expression.raw
+##        go.maize.raw
+##        go.sorghum.raw
+## Output:
+##        syntelogs.sorghum.v1.maize.v1.clean
+##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean
+##        maize.expression.clean
+##        go.maize.clean
+##        go.sorghum.clean
+##
+## Date: 2017-08-25
+## Author: Jesse R. Walsh
+####################################################################################################
+#==================================================================================================#
+## syntelogs.sorghum.v1.maize.v1.raw
+#--------------------------------------------------------------------------------------------------#
+syntelogs.sorghum.v1.maize.v1.clean <- syntelogs.sorghum.v1.maize.v1.raw
+
+#==================================================================================================#
+## syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw
+#--------------------------------------------------------------------------------------------------#
+syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean <- syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw
+
+## For v4 imports, need to remove the CDS: and _T00# parts of the gene2 column
+syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene2 <- gsub(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene2, pattern = "CDS:", replacement = "")
+syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene2 <- gsub(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene2, pattern = "_T\\d\\d\\d", replacement = "")
+
+## For sorghum v3.1, need to change identifier to SORBI_ format
+syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene1 <- gsub(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene1, pattern = "Sobic.", replacement = "SORBI_")
[…]

### 2017-08-31 R/loadData.R

@@ -1,0 +1,50 @@
+library(readr)
+####################################################################################################
+## Project:         Subgenomes project
+## Script purpose:  Import raw dataset files relevant to this project
+##
+## Output:
+##        syntelogs.sorghum.v1.maize.v1.raw
+##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw
+##        maize.expression.raw
+##        go.maize.raw
+##        go.sorghum.raw
+##        go.goSlim.plant
+##        subgenome.assignments
+##        subgenome.truth
+##
+## Date: 2017-08-25
+## Author: Jesse R. Walsh
+####################################################################################################
+
+## Import the raw data from the parsed SynMap output
+syntelogs.sorghum.v1.maize.v1.raw <- read_delim("./Data/SynMap/sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw <- read_delim("./Data/SynMap/sorghum_v3.1_vs_maize_v4+rejected.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+
+## Expression data from the Walley 2016 paper in FPKM over 68 tissues
+maize.expression.raw <- read_delim("./Data/Expression/GSE50191_FPKM.tsv", "\t", trim_ws = TRUE)
+
+## Read in GO Annotation data for maize genes
+go.maize.raw <- read_delim("./Data/GO/go_from_ maizecyc.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+
+## Read in GO Annotation data for sorghum genes
+go.sorghum.raw <- read_delim("./Data/GO/gramene_sorghumv2_goterms.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
+
+## Read in list of plant GO-slim subset
+go.goSlim.plant <- read_delim("./Data/GO/goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+
+## This represents subgenome assignments given using a greedy approach (currently performed by hand)
+subgenome.assignments <- read_delim("./Data/outfile_processed_byHand.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+
+## This represents the subgenome assignments given in the paper Schnable 2011
[…]

### 2017-08-31 R/parseSubgenomes.R

@@ -6,7 +6,7 @@ library(dplyr)
 ## Script purpose: This script should label each gene in the input file as either part of the dominant
 ##                subgenome or the recessive subgenome.
 ##
-## Input: data input fileName, log10_ks_cutoff
+## Input: data input fileName, log10_ks_cutoff, syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean
 ## Output:
 ##        syntelogs.raw         ->
 ##        syntelogs.mutated     ->
@@ -19,35 +19,36 @@ library(dplyr)
 ## Author: Jesse R. Walsh
 ####################################################################################################
 
-inputDataFile <- params$inputDataFile
-subGenomeFile <- params$subGenomeFile
+# inputDataFile <- params$inputDataFile
+# subGenomeFile <- params$subGenomeFile
 log10_ks_cutoff <- params$log10_ks_cutoff
+# log10_ks_cutoff <- 0
 
 # inputDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\sorghum_v1_vs_maize_v1.tab"
 # subGenomeFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\deleteme\\outfile_processed_byHand.tab"
 
-## Import the raw data from the parsed SynMap output
-syntelogs.raw <- read_delim(inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
-
-## For v4 imports, need to remove the CDS: and _T00# parts of the gene2 column
-syntelogs.raw$gene2 <- gsub(syntelogs.raw$gene2, pattern = "CDS:", replacement = "")
-syntelogs.raw$gene2 <- gsub(syntelogs.raw$gene2, pattern = "_T\\d\\d\\d", replacement = "")
-
-## Import "True" sugenomes as reported by Schnable 2011
-subgenome.truth <- setNames(data.frame(
-  c(1,1,1,2,2,3,3,4,4,5,5,6,6,7,7,7,7,8,8,8,9,9,9,10,10,10),
-  c(1,5,9,7,2,3,8,5,4,4,2,2,10,1,6,10,4,1,10,3,6,10,8,5,9,6),
-  c("sub1","sub2","sub2","sub1","sub2","sub1","sub2","sub1","sub2",
-    "sub1","sub2","sub1","sub2","sub1","sub1","sub1","sub2","sub1",
[…]

### 2017-09-01 SubGenome Analysis.Rmd
Added windows paths for load/source

@@
 ```{r importSynMapData, include=FALSE}
-source("~/git/SubGenomes/R/loadData.R")
-source("~/git/SubGenomes/R/cleanData.R")
-source("~/git/SubGenomes/R/parseSubgenomes.R")
-source("~/git/SubGenomes/R/addGOTerms.R")
-source("~/git/SubGenomes/R/parseGOSlim.R")
-# source("~/git/SubGenomes/R/createTopGO.R")
-source("~/git/SubGenomes/R/parseExpressionData.R")
+# source("~/git/SubGenomes/R/loadData.R")
+# source("~/git/SubGenomes/R/cleanData.R")
+# source("~/git/SubGenomes/R/parseSubgenomes.R")
+# source("~/git/SubGenomes/R/addGOTerms.R")
+# source("~/git/SubGenomes/R/parseGOSlim.R")
+# # source("~/git/SubGenomes/R/createTopGO.R")
+# source("~/git/SubGenomes/R/parseExpressionData.R")
 
@@
 
-# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseSubgenomes.R")
-# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\addGOTerms.R")
+source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\loadData.R")
+source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\cleanData.R")
+source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseSubgenomes.R")
+source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\addGOTerms.R")
+source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseGOSlim.R")
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\createTopGO.R")
+source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseExpressionData.R")
 ```

### 2017-09-01 R/loadData.R

@@ -1,4 +1,5 @@
 library(readr)
+library(readxl)
 ####################################################################################################
 ## Project:         Subgenomes project
 ## Script purpose:  Import raw dataset files relevant to this project
@@ -12,29 +13,39 @@ library(readr)
 ##        go.goSlim.plant
 ##        subgenome.assignments
 ##        subgenome.truth
+##        maize.genes.v3_to_v4_map
 ##
 ## Date: 2017-08-25
 ## Author: Jesse R. Walsh
 ####################################################################################################
 
-## Import the raw data from the parsed SynMap output
-syntelogs.sorghum.v1.maize.v1.raw <- read_delim("./Data/SynMap/sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw <- read_delim("./Data/SynMap/sorghum_v3.1_vs_maize_v4+rejected.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+syntelogs.sorghum.v1.maize.v1.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\SynMap\\sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\SynMap\\sorghum_v3.1_vs_maize_v4+rejected.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+maize.expression.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\Expression\\GSE50191_FPKM.tsv", "\t", trim_ws = TRUE)
+go.maize.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO\\go_from_ maizecyc.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+go.sorghum.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO\\gramene_sorghumv2_goterms.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
[…]

### 2017-09-01 SubGenome Analysis.Rmd
Minor

--- 2017-08-31 SubGenome Analysis.Rmd
@@
 
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\loadData.R")
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\cleanData.R")
 # source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseSubgenomes.R")
 # source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\addGOTerms.R")
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseGOSlim.R")
+# # source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\createTopGO.R")
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseExpressionData.R")
 ```

### 2017-09-01 R/loadData.R

@@ -19,33 +19,33 @@ library(readxl)
 ## Author: Jesse R. Walsh
 ####################################################################################################
 
-syntelogs.sorghum.v1.maize.v1.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\SynMap\\sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\SynMap\\sorghum_v3.1_vs_maize_v4+rejected.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-maize.expression.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\Expression\\GSE50191_FPKM.tsv", "\t", trim_ws = TRUE)
-go.maize.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO\\go_from_ maizecyc.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-go.sorghum.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO\\gramene_sorghumv2_goterms.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
-go.goSlim.plant <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO\\goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-subgenome.assignments <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\outfile_processed_byHand.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-maize.genes.v3_to_v4_map.raw <- read_xlsx("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\MaizeGDB_v3_v4.genes.xlsx")
+# syntelogs.sorghum.v1.maize.v1.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\SynMap\\sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
[…]

### 2017-09-13 SubGenome Analysis.Rmd
Large renaming and reorganization

--- 2017-08-31 SubGenome Analysis.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
-library(topGO)
-library(readr)
-library(readxl)
-library(ggplot2)
-library(tidyr)
-library(dplyr)
-library(VennDiagram)
-library(gridExtra)
 ```
@@
 source("~/git/SubGenomes/R/cleanData.R")
-source("~/git/SubGenomes/R/parseSubgenomes.R")
-source("~/git/SubGenomes/R/addGOTerms.R")
-source("~/git/SubGenomes/R/parseGOSlim.R")
+source("~/git/SubGenomes/R/functions.R")
+source("~/git/SubGenomes/R/do.R")
+
+# source("~/git/SubGenomes/R/parseSubgenomes.R")
+# source("~/git/SubGenomes/R/addGOTerms.R")
+# source("~/git/SubGenomes/R/parseGOSlim.R")
 # source("~/git/SubGenomes/R/createTopGO.R")
-source("~/git/SubGenomes/R/parseExpressionData.R")
+# source("~/git/SubGenomes/R/parseExpressionData.R")
 
-MFterms <- ls(GOMFTerm)
-
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\loadData.R")
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\cleanData.R")
 # source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseSubgenomes.R")
 # source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\addGOTerms.R")
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseGOSlim.R")
+# # source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\createTopGO.R")
+# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseExpressionData.R")
+
+library(topGO)
+library(readr)
+library(readxl)
+library(ggplot2)
+library(tidyr)
+library(dplyr)
+library(VennDiagram)
+library(gridExtra)
+
+# topGO::groupGOTerms()
+MFterms <- ls(GOMFTerm)
 ```

### 2017-09-13 R/do.R

@@ -1,0 +1,79 @@
+library(tidyr)
+library(dplyr)
+#==================================================================================================#
+## parseSubgenomes.R
+#--------------------------------------------------------------------------------------------------#
+source("~/git/SubGenomes/R/parseSubgenomes.R")
+
+#==================================================================================================#
+## addGOTerms.R
+#--------------------------------------------------------------------------------------------------#
+## GO Analysis -> grouping into useful slices
+goAnnotations.sub1 <-
+  subgenome %>%
+  subset(subgenome == "sub1") %>%
+  distinct() %>%
+  inner_join(go.maize.clean, by = c("gene2" = "geneID")) %>%
+  select(gene2, goTerm, evCode) %>%
+  distinct()
+
+goAnnotations.sub2 <-
+  subgenome %>%
+  subset(subgenome == "sub2") %>%
+  distinct() %>%
+  inner_join(go.maize.clean, by = c("gene2" = "geneID")) %>%
+  select(gene2, goTerm, evCode) %>%
+  distinct()
+
+#==================================================================================================#
+## parseSorghumGO.R
+#--------------------------------------------------------------------------------------------------#
+maizeWithSorghumGO <-
+  syntelogs.mutated %>%
+  select(gene1, gene2)
+
+maizeWithSorghumGO <-
+  maizeWithSorghumGO %>%
+  left_join(go.sorghum.clean, by=c("gene1"="sorghumID"))
+
+maizeWithSorghumGO <-
+  maizeWithSorghumGO %>%
+  filter(!is.na(goTerm)) %>%
+  select(gene2, goTerm) %>%
+  distinct()
+
+#==================================================================================================#
+## parseExpressionData.R
+#--------------------------------------------------------------------------------------------------#
+expressedGenes <- data.frame(ID=maize.expression.clean[,1], Means=rowMeans(maize.expression.clean[,-1], na.rm = TRUE))
+
+## Remove rows with NA
+expressedGenes <-
+  expressedGenes %>%
+  mutate(FPKM_mean = Means) %>%
+  select(tracking_id, FPKM_mean) %>%
+  rename(geneID = tracking_id) %>%
+  subset(!is.na(FPKM_mean))
+
+## Convert to v4 ids (this is a bad conversion table, only has CornCyc Genes in it.... get a better one!)
+geneIDConvertTable <- maize.genes.v3_to_v4_map.clean
+
[…]

### 2017-09-13 R/parseSubgenomes.R

@@ -18,30 +18,8 @@ library(dplyr)
 ## Date: 2017-07-26
 ## Author: Jesse R. Walsh
 ####################################################################################################
-
-# inputDataFile <- params$inputDataFile
-# subGenomeFile <- params$subGenomeFile
 log10_ks_cutoff <- params$log10_ks_cutoff
-# log10_ks_cutoff <- 0
-
-# inputDataFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\sorghum_v1_vs_maize_v1.tab"
-# subGenomeFile <- "C:\\Users\\Jesse\\Dropbox (Personal)\\deleteme\\outfile_processed_byHand.tab"
-
-# ## Import the raw data from the parsed SynMap output
-# syntelogs.raw <- read_delim(inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
-#
-# ## For v4 imports, need to remove the CDS: and _T00# parts of the gene2 column
-# syntelogs.raw$gene2 <- gsub(syntelogs.raw$gene2, pattern = "CDS:", replacement = "")
-# syntelogs.raw$gene2 <- gsub(syntelogs.raw$gene2, pattern = "_T\\d\\d\\d", replacement = "")
-#
-# ## Import "True" sugenomes as reported by Schnable 2011
-# subgenome.truth <- setNames(data.frame(
-#   c(1,1,1,2,2,3,3,4,4,5,5,6,6,7,7,7,7,8,8,8,9,9,9,10,10,10),
-#   c(1,5,9,7,2,3,8,5,4,4,2,2,10,1,6,10,4,1,10,3,6,10,8,5,9,6),
-#   c("sub1","sub2","sub2","sub1","sub2","sub1","sub2","sub1","sub2",
-#     "sub1","sub2","sub1","sub2","sub1","sub1","sub1","sub2","sub1",
-#     "sub1","sub2","sub1","sub1","sub2","sub1","sub1","sub2"),
-#   stringsAsFactors=FALSE), c("chr1","chr2","subgenome"))
+ks_cutoff <- 10^log10_ks_cutoff
 
 ## Add median and geneCount values (aggregated by block/org_chr1/org_chr2) to each row.  Calculate gene sizes.  Add chromosome ids as numbers.
 syntelogs.mutated <-
@@ -53,9 +31,6 @@ syntelogs.mutated <-
   mutate(gene_length2=abs(stop2-start2)) %>%
[…]

### 2017-09-14 SubGenome Analysis.Rmd
Another pass at data cleaning
Particularly GO terms for maize.  Added go term type and simplified ev codes to either exp or comp.

@@
 # topGO::groupGOTerms()
-MFterms <- ls(GOMFTerm)
+# MFterms <- ls(GOMFTerm)
+# BPterms <- ls(GOBPTerm)
+# CCterms <- ls(GOCCTerm)
 ```
@@
 Type <- c("EV-EXP","EV-COMP")
-Count <- c(go.maize.clean %>% filter(grepl("EV-EXP", evCode)) %>% nrow(),
-      go.maize.clean %>% filter(grepl("EV-COMP", evCode)) %>% nrow())
+Count <- c(go.maize.clean %>% filter(evCode=="EXP") %>% nrow(),
+      go.maize.clean %>% filter(evCode=="COMP") %>% nrow())
 df = data.frame(Type,Count)
@@
 ) + labs(
-  title="GO Terms by EVCode"
+  title="GO Term Assignments by Evidence Code"
 )
@@
 g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm), 
-                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm)
+  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% select(goTerm), 
+                         goAnnotations.sub2 %>% filter(evCode=="EXP") %>% select(goTerm)
                          ) %>% nrow(), 
@@
 grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP only)")
+```
+
+GO Terms by subgenome, only counting GO Terms that are in GOSlimPlant
+
+```{r GOTerms_by_subgenome_goslim, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+# All EV-Codes, Unique GO Terms
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm), 
+                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm)) %>% 
+              nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim only)")
+
+#EXP only, Unique GO Terms
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(evCode=="EXP") %>% select(goTerm), 
+                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm)
+                         ) %>% nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP only, GOSlim only)")
+```
+
+GO Terms by subgenome, only counting GO Terms that are in GOSlimPlant and are Molecular Function terms.
+
+```{r GOTerms_by_subgenome_goslim_MF, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+# All EV-Codes, Unique GO Terms
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(type=="MF") %>% select(goTerm) %>% 
+    distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(type=="MF")  %>% select(goTerm) %>% 
+    distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(type=="MF") %>% select(goTerm), 
+                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(type=="MF") %>% select(goTerm)) %>%
+                          nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim, MF only)")
+
+#EXP only, Unique GO Terms
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>%
+    distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>%
+    distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(evCode=="EXP") %>% filter(type=="MF") %>%
+                           select(goTerm), 
+                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% filter(type=="MF") %>%
+                           select(goTerm)
+                         ) %>% nrow(), 
+  category = c("SubGenome 1", "SubGenome 2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim, MF only)")
 ```
@@
 g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", evCode)) %>% nrow(),
-  area2 = converted %>% filter(grepl("EV-EXP", evCode)) %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", evCode)) %>% select(gene2, goTerm) %>% distinct(), 
-                         converted %>% filter(grepl("EV-EXP", evCode)) %>% select(gene2, goTerm) %>% distinct()
+  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% nrow(),
+  area2 = converted %>% filter(evCode=="EXP") %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% select(gene2, goTerm) %>% distinct(), 
+                         converted %>% filter(evCode=="EXP") %>% select(gene2, goTerm) %>% distinct()
                          ) %>% nrow(), 
@@
 grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(EV-EXP only)")
-```
-
-GO Terms by subgenome, only counting GO Terms that are in GOSlimPlant
-
-```{r GOTerms_by_subgenome_goslim, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
-# All EV-Codes, Unique GO Terms
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm), 
-                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm)) %>% 
-              nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim only)")
-
-#EXP only, Unique GO Terms
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm), 
-                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm)
-                         ) %>% nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP only, GOSlim only)")
-```
-
-GO Terms by subgenome, only counting GO Terms that are in GOSlimPlant and are Molecular Function terms.
-
-```{r GOTerms_by_subgenome_goslim_MF, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
-# All EV-Codes, Unique GO Terms
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% 
-    filter(goTerm %in% MFterms) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% 
-    filter(goTerm %in% MFterms) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% filter(goTerm %in% MFterms), 
-                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% filter(goTerm %in% MFterms)) %>%
-                          nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim, MF only)")
-
-#EXP only, Unique GO Terms
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% 
-    filter(goTerm %in% MFterms) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% 
-    filter(goTerm %in% MFterms) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% 
-                           filter(goTerm %in% MFterms), 
-                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(grepl("EV-EXP", evCode)) %>% select(goTerm) %>% 
-                           filter(goTerm %in% MFterms)
-                         ) %>% nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim, MF only)")
 ```
@@
 # Expression
+
+***
 

### 2017-09-14 R/cleanData.R

@@ -1,3 +1,4 @@
+library(topGO)
 library(tidyr)
 library(dplyr)
 startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
@@ -30,6 +31,8 @@ startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
 #--------------------------------------------------------------------------------------------------#
 syntelogs.sorghum.v1.maize.v1.clean <- syntelogs.sorghum.v1.maize.v1.raw
 
+## ks and kn values are sometimes NA, this will be allowed
+
 #==================================================================================================#
 ## syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw
 #--------------------------------------------------------------------------------------------------#
@@ -42,6 +45,8 @@ syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene2 <- gsub(syntelogs.sorgh
 ## For sorghum v3.1, need to change identifier to SORBI_ format
 syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene1 <- gsub(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene1, pattern = "Sobic.", replacement = "SORBI_")
 syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene1 <- gsub(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene1, pattern = ".\\d.v\\d.\\d", replacement = "")
+
+## ks and kn values are sometimes NA, this will be allowed
 
 #==================================================================================================#
 ## maize.expression.raw
@@ -69,8 +74,27 @@ go.maize.clean <-
   select(geneID, goTerm, Citation) %>%
   separate(Citation, c("publication", "evCode","timeStamp","curator"), sep=":", extra="drop")
 
+## Publication is uncommon, replace blanks with NA's
+go.maize.clean$publication[go.maize.clean$publication == ""] <- NA
+
+## Simplify evidence codes, assume missing is computationally derived.
+go.maize.clean$evCode[grepl("EV-EXP", go.maize.clean$evCode)] <- "EXP"
+go.maize.clean$evCode[grepl("EV-AS", go.maize.clean$evCode)] <- "EXP" #trust author statements
[…]

### 2017-09-14 R/do.R

@@ -14,7 +14,7 @@ goAnnotations.sub1 <-
   subset(subgenome == "sub1") %>%
   distinct() %>%
   inner_join(go.maize.clean, by = c("gene2" = "geneID")) %>%
-  select(gene2, goTerm, evCode) %>%
+  select(gene2, goTerm, evCode, type) %>%
   distinct()
 
 goAnnotations.sub2 <-
@@ -22,7 +22,7 @@ goAnnotations.sub2 <-
   subset(subgenome == "sub2") %>%
   distinct() %>%
   inner_join(go.maize.clean, by = c("gene2" = "geneID")) %>%
-  select(gene2, goTerm, evCode) %>%
+  select(gene2, goTerm, evCode, type) %>%
   distinct()
 
 #==================================================================================================#

### 2017-09-14 SubGenome Analysis.Rmd
More GO comparison, cleaning up text and plots
Changed venn's of goslim+MF filters to just MF filters, since using both at the same time filtered too much.

Added boxplots for GO by ontology type.

In an effort to make the report more readable, shrank plot size, reorganized in a more intuitive way, added more prose.

@@
 In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size. 
-**(TODO)**
-
-foreach sorghum chromosome, set the largest syntenic chromosome (by gene) as sub1.
+Psuedocode of the algorithm is as follows: foreach sorghum chromosome, set the largest syntenic chromosome (by gene) as sub1.
 Foreach additional syntenic chromosome if doesn't overlap what is already in sub1, add it to sub1.
@@
 Items in center are correct.
-(this is an unfair comparison if I sorted by hand)
+(currently sorted by hand)
 
@@
 
-Genes placed in the "correct" subgenome
+<!-- print("Missed subgenome assignments:") -->
+<!-- setdiff(subgenome.truth, subgenome.chromosomes.test) -->
 
-**(TODO)**
+<!-- print("False positive subgenome assignments:") -->
+<!-- setdiff(subgenome.chromosomes.test, subgenome.truth) -->
 
-```{r compareResultsToPaper_genes, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=3, fig.align='center'}
-#need to import "truth" and compare
-#visualize as venn
-```
-
-# GO Terms
+# GO Terms in Maize
 ***
 
-```{r GOTerms_by_EVCode, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4, fig.align='center'}
-Type <- c("EV-EXP","EV-COMP")
-Count <- c(go.maize.clean %>% filter(evCode=="EXP") %>% nrow(),
-      go.maize.clean %>% filter(evCode=="COMP") %>% nrow())
+First we show both how many unique GO terms have been assigned using a computational ev-code vs.
+an experimental ev-code.
+A few terms have been annotated using both or were assigned multiple times based on different publications, which is why n is greater than the number of unique GO terms in our set.
+We also break down the GO terms by type.
+A few terms may not have a type, typically because they have been made obsolete.
+
+Second, we look at the same divisions but this time for all gene annotations (i.e.
+gene + GO Term + Ev-Code).
+Again, since some terms are assigned computationally and experimentally to the same gene or assigned multiple times based on different publications, n is greater than the number of unique GO term assignments in out set.
+The main takeaways here are that computationally assigned GO terms vastly outnumber experimental annotations, most annotations use MF terms, and there are a few terms falling through the cracks most likely due to being outdated.
+
+```{r GOTerms_in_Maize, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=2.3, fig.height=3.6}
+Evidence <- c("EV-EXP","EV-COMP")
+Count <- c(go.maize.clean %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+      go.maize.clean %>% filter(evCode=="COMP") %>% select(goTerm) %>% distinct() %>% nrow())
+df = data.frame(Evidence,Count)
+ggplot(data=df, aes(x=Evidence, y=Count)) +
+  geom_bar(stat="identity", fill="steelblue") +
+  geom_text(aes(label=Count), vjust=-0.3, size=3.5
+) + labs(
+  title=paste("GO Terms by EV Code\n(n=", nrow(unique(go.maize.clean[,c("goTerm", "evCode")])),")", sep = "")
+) + theme(plot.title = element_text(size = rel(.8)))
+
+Type <- c("MF","BP","CC","<NA>")
+Count <- c(go.maize.clean %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
+           go.maize.clean %>% filter(type=="BP") %>% select(goTerm) %>% distinct() %>% nrow(),
+           go.maize.clean %>% filter(type=="CC") %>% select(goTerm) %>% distinct() %>% nrow(),
+           go.maize.clean %>% filter(is.na(type)) %>% select(goTerm) %>% distinct() %>% nrow())
 df = data.frame(Type,Count)
@@
 ) + labs(
-  title="GO Term Assignments by Evidence Code"
-)
+  title=paste("GO Terms by Type\n(n=", nrow(unique(go.maize.clean[,c("goTerm", "type")])),")", sep = "")
+) + theme(plot.title = element_text(size = rel(.8)))
+
+Evidence <- c("EV-EXP","EV-COMP")
+Count <- c(go.maize.clean %>% filter(evCode=="EXP") %>% nrow(),
+      go.maize.clean %>% filter(evCode=="COMP") %>% nrow())
+df = data.frame(Evidence,Count)
+ggplot(data=df, aes(x=Evidence, y=Count)) +
+  geom_bar(stat="identity", fill="steelblue") +
+  geom_text(aes(label=Count), vjust=-0.3, size=3.5
+) + labs(
+  title=paste("GO Term Annotations by\nEV Code (n=", nrow(go.maize.clean),")", sep = "")
+) + theme(plot.title = element_text(size = rel(.8)))
+
+Type <- c("MF","BP","CC","<NA>")
+Count <- c(go.maize.clean %>% filter(type=="MF") %>% select(geneID, goTerm, evCode) %>% distinct() %>% nrow(),
+           go.maize.clean %>% filter(type=="BP") %>% select(geneID, goTerm, evCode) %>% distinct() %>% nrow(),
+           go.maize.clean %>% filter(type=="CC") %>% select(geneID, goTerm, evCode) %>% distinct() %>% nrow(),
+           go.maize.clean %>% filter(is.na(type)) %>% select(geneID, goTerm, evCode) %>% distinct() %>% nrow())
+df = data.frame(Type,Count)
+ggplot(data=df, aes(x=Type, y=Count)) +
+  geom_bar(stat="identity", fill="steelblue") +
+  geom_text(aes(label=Count), vjust=-0.3, size=3.5
+) + labs(
+  title=paste("GO Term Annotations by\nType (n=", nrow(unique(go.maize.clean[,c("geneID","goTerm","evCode","type")])),")", sep = "")
+) + theme(plot.title = element_text(size = rel(.8)))
 ```
 
+# GO Term Comparison by Subgenome
 ***
@@
 
-```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
-# All EV-Codes, Unique GO Terms
+The main takeaways here are that 1) there are differences in functional assignments between each subgenome, regardless of which filters I apply.
+2) the GOSlim for plants is extremely restrictive, and probably not useful.
+3) Sub1 has more functional assignments than sub2, although this may be a direct correlation to the larger size (by definition) of sub1.
+
+```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=3, fig.height=3}
+# All EV-Codes
 grid.newpage()
@@
   cross.area = intersect(goAnnotations.sub1 %>% select(goTerm), goAnnotations.sub2 %>% select(goTerm)) %>% nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
+  category = c("Sub1", "Sub2"), 
   lty = rep("blank", 2), 
@@
 
-#EXP only, Unique GO Terms
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% select(goTerm), 
-                         goAnnotations.sub2 %>% filter(evCode=="EXP") %>% select(goTerm)
-                         ) %>% nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP only)")
-```
-
-GO Terms by subgenome, only counting GO Terms that are in GOSlimPlant
-
-```{r GOTerms_by_subgenome_goslim, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
-# All EV-Codes, Unique GO Terms
+# All EV-Codes, GOSlim only
 grid.newpage()
@@
               nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
+  category = c("Sub1", "Sub2"), 
   lty = rep("blank", 2), 
@@
 
-#EXP only, Unique GO Terms
+# All EV-Codes, MF terms only
 grid.newpage()
 g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(evCode=="EXP") %>% select(goTerm), 
-                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm)
-                         ) %>% nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
+  area1 = goAnnotations.sub1 %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(type=="MF") %>% select(goTerm), 
+                         goAnnotations.sub2 %>% filter(type=="MF") %>% select(goTerm)) %>%
+                          nrow(), 
+  category = c("Sub1", "Sub2"), 
   lty = rep("blank", 2), 
@@
 )
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP only, GOSlim only)")
-```
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, MF only)")
 
-GO Terms by subgenome, only counting GO Terms that are in GOSlimPlant and are Molecular Function terms.
-
-```{r GOTerms_by_subgenome_goslim_MF, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
-# All EV-Codes, Unique GO Terms
+#EXP only
 grid.newpage()
 g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(type=="MF") %>% select(goTerm) %>% 
-    distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(type=="MF")  %>% select(goTerm) %>% 
-    distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(type=="MF") %>% select(goTerm), 
-                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(type=="MF") %>% select(goTerm)) %>%
-                          nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
+  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% select(goTerm), 
+                         goAnnotations.sub2 %>% filter(evCode=="EXP") %>% select(goTerm)
+                         ) %>% nrow(), 
+  category = c("Sub1", "Sub2"), 
   lty = rep("blank", 2), 
@@
 )
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim, MF only)")
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP only)")
 
-#EXP only, Unique GO Terms
+#EXP only, GOSlim only
 grid.newpage()
 g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>%
-    distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>%
-    distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(evCode=="EXP") %>% filter(type=="MF") %>%
-                           select(goTerm), 
-                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% filter(type=="MF") %>%
-                           select(goTerm)
+  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(evCode=="EXP") %>% select(goTerm), 
+                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm)
                          ) %>% nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
+  category = c("Sub1", "Sub2"), 
   lty = rep("blank", 2), 
@@
 )
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim, MF only)")
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim only)")
+
+#EXP only, MF terms only
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
+  area2 = goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm), 
+                         goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm)
+                         ) %>% nrow(), 
+  category = c("Sub1", "Sub2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, MF only)")
 ```
@@
 
-```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+The main takeaways here mirror the assessment from the GO Term presence/absence check above.
+There are many differences in assignments, although we can assume that the same terms are reused very often based on the differenced between unique terms and gene annotations.
+
+```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=3, fig.height=3}
 ## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2.  The converted data.frame will hold maize2 genes with the names replaced with maize1 homologs
@@
   converted %>%
-  select(Maize2, goTerm, evCode) %>%
+  select(Maize2, goTerm, evCode, type) %>%
   rename(gene2=Maize2)
@@
                          ) %>% nrow(),
-  category = c("SubGenome 1", "SubGenome 2"), 
+  category = c("Sub1", "Sub2"), 
   lty = rep("blank", 2), 
@@
                          ) %>% nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
+  category = c("Sub1", "Sub2"), 
   lty = rep("blank", 2), 
@@
 grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(EV-EXP only)")
-```
 
-```{r aggregates, include=FALSE}
-goAnnotations.sub1.aggr <-
-  goAnnotations.sub1 %>%
-  select(goTerm) %>%
-  group_by(goTerm) %>%
-  summarise(n=n())
-
-goAnnotations.sub2.aggr <-
-  goAnnotations.sub2 %>%
-  select(goTerm) %>%
-  group_by(goTerm) %>%
-  summarise(n=n())
+#EXP only, MF only
+grid.newpage()
+g <- draw.pairwise.venn(
+  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% nrow(),
+  area2 = converted %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% nrow(),
+  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(gene2, goTerm) %>% distinct(), 
+                         converted %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(gene2, goTerm) %>% distinct()
+                         ) %>% nrow(), 
+  category = c("Sub1", "Sub2"), 
+  lty = rep("blank", 2), 
+  fill = c("light blue", "green"), 
+  alpha = rep(0.5, 2), 
+  cat.pos = c(0, 180), 
+  scaled = FALSE,
+  euler.d = FALSE, 
+  sep.dist = 0.3, 
+  rotation.degree = 0,
+  ind = FALSE
+)
+grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(EV-EXP, MF only)")
 ```
@@
                           nrow(), 
-  category = c("SubGenome 1", "SubGenome 2"), 
+  category = c("Sub1", "Sub2"), 
   lty = rep("blank", 2), 

### 2017-09-14 R/do.R
Minor
Unused code, moved it here for now.

@@ -25,6 +25,18 @@ goAnnotations.sub2 <-
   select(gene2, goTerm, evCode, type) %>%
   distinct()
 
+goAnnotations.sub1.aggr <-
+  goAnnotations.sub1 %>%
+  select(goTerm) %>%
+  group_by(goTerm) %>%
+  summarise(n=n())
+
+goAnnotations.sub2.aggr <-
+  goAnnotations.sub2 %>%
+  select(goTerm) %>%
+  group_by(goTerm) %>%
+  summarise(n=n())
+
 #==================================================================================================#
 ## parseSorghumGO.R
 #--------------------------------------------------------------------------------------------------#

### 2017-09-15 SubGenome Analysis.Rmd
Load, filter, and plot isoform frequencies

@@
 #Total genes for each subgenome by chr and overall
+```
+
+In order to see if one subgenome tends to have more isoforms, we plot the density of isoform counts from each subgenome based on isoforms in the v4 release 32 GFF file.
+We scale the x axis using log10, as we have a very long tail on this data (i.e.
+most counts are 1, but a few range to nearly 600).
+The densities track each other very strongly, so there isn't likely anything to look for here.
+
+```{r subgenome_isoform_counts, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=6, fig.height=4, fig.align='center'}
+## Plot frequency of isoform counts for each gene
+data <- geneTranscript.counts %>% inner_join(subgenome, by=c("gene"="gene2")) %>% select(gene, n, subgenome) %>% filter(subgenome=="sub1" | subgenome=="sub2")
+ggplot(data, aes(n,color=subgenome)) + 
+  geom_density() + 
+  scale_x_log10() +
+  labs(y = "Density (Isoform Counts)", 
+       x = "log10(Isoform Counts)", 
+       title = "Comparison of Density of Isoform Counts between\nGenes in Subgenome 1 and Subgenome 2"
+  )
 ```
@@
 ```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
-## Attach log2(FPKM_mean) values to homeologous pairs
-expressedPairs <-
-  homeologs.pairs %>%
-  subset(Maize1 != "" & Maize2 != "") %>%
-  select(Maize1, Maize2) %>%
-  inner_join(expressedGenes, by=c("Maize1"="geneID")) %>%
-  rename(FPKM_mean1=FPKM_mean) %>%
-  inner_join(expressedGenes, by=c("Maize2"="geneID")) %>%
-  rename(FPKM_mean2=FPKM_mean)
-
 # Expression Diffs

### 2017-09-15 R/do.R

@@ -1,3 +1,4 @@
+library(GenomicFeatures)
 library(tidyr)
 library(dplyr)
 ####################################################################################################
@@ -76,14 +77,22 @@ expressedGenes <-
   rename(geneID = tracking_id) %>%
   subset(!is.na(FPKM_mean))
 
-## Convert to v4 ids (this is a bad conversion table, only has CornCyc Genes in it.... get a better one!)
-geneIDConvertTable <- maize.genes.v3_to_v4_map.clean
-
+## Convert to v4 ids
 expressedGenes <-
   expressedGenes %>%
-  inner_join(geneIDConvertTable, by=c("geneID" = "v3_id")) %>%
+  inner_join(maize.genes.v3_to_v4_map.clean, by=c("geneID" = "v3_id")) %>%
   select(v4_id, FPKM_mean) %>%
   rename(geneID=v4_id)
+
+## Attach log2(FPKM_mean) values to homeologous pairs
+expressedPairs <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  inner_join(expressedGenes, by=c("Maize1"="geneID")) %>%
+  rename(FPKM_mean1=FPKM_mean) %>%
+  inner_join(expressedGenes, by=c("Maize2"="geneID")) %>%
+  rename(FPKM_mean2=FPKM_mean)
 
 #==================================================================================================#
 ## analyzeGODiffs.R
@@ -95,6 +104,32 @@ expressedGenes <-
 #--------------------------------------------------------------------------------------------------#
 # source("~/git/SubGenomes/R/createTopGO.R")
 
+#==================================================================================================#
+## txdb -> geneTranscript.map
 #--------------------------------------------------------------------------------------------------#
+## Only work with chromosomes, ignore unplaced contigs
+seqlevels(txdb) <- c("1","2","3","4","5","6","7","8","9","10")
+
+## Get gene/transcript names
+geneTranscript.map <- data.frame(transcripts(txdb)$tx_name)
+# GRList <- exonsBy(txdb, by = "tx")
+# tx_ids <- names(GRList)
+# head(select(txdb, keys=tx_ids, columns=c("GENEID","TXNAME"), keytype="TXID"))
+
+## Clean geneTranscript.map
+geneTranscript.map <-
+  geneTranscript.map %>%
[…]

### 2017-09-15 R/loadData.R

@@ -1,5 +1,6 @@
 library(readr)
 library(readxl)
+library(GenomicFeatures)
 ####################################################################################################
 ## Project:         Subgenomes project
 ## Script purpose:  Import raw dataset files relevant to this project
@@ -14,6 +15,7 @@ library(readxl)
 ##        subgenome.assignments
 ##        subgenome.truth
 ##        maize.genes.v3_to_v4_map.raw
+##        txdb
 ##
 ## Date: 2017-08-25
 ## Author: Jesse R. Walsh
@@ -59,5 +61,10 @@ subgenome.truth <- setNames(data.frame(
 ## Mapping data provided by Maggie, based on synteny from SynMap
 maize.genes.v3_to_v4_map.raw <- read_xlsx("./Data/MaizeGDB_v3_v4.genes.xlsx")
 
+## Load maize GFF data
+txdb <- makeTxDbFromGFF("./Data/Zea_mays.AGPv4.32.gff3.gz", format="gff3")
+
 #--------------------------------------------------------------------------------------------------#
 detach("package:readr", unload=TRUE)
+detach("package:readxl", unload=TRUE)
+detach("package:GenomicFeatures", unload=TRUE)

### 2017-09-15 SubGenome Analysis.Rmd
Comparing based on expression

@@
 
-```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4, fig.align='center'}
 # Expression Diffs
@@
 
+Overall expression patterns between subgenome 1 and subgenome 2 are highly similar.
+Below we show to views of the same data.
+In the density graph, we can see that subgenome 1 is slightly shifted to the right of subgenome 2 expression, but otherwise follows the same shape.
+Viewed as a boxplot, you can again see that the quartiles in subgenome 1 and 2 are largely the same.
+This also shows a few high expressing outliers on subgenome 1.
+
+```{r subgenome_expression_frequency, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4.5, fig.height=4}
+## Plot frequency of isoform counts for each gene
+data <- expressedGenes %>% inner_join(subgenome, by=c("geneID"="gene2")) %>% select(geneID, FPKM_mean, subgenome) %>% filter(subgenome=="sub1" | subgenome=="sub2")
+ggplot(data, aes(FPKM_mean,color=subgenome)) + 
+  geom_density() + 
+  scale_x_log10() +
+  labs(y = "Density (FPKM)", 
+       x = "log10(FPKM)", 
+       title = "Comparison of FPKM Expression between\nGenes in Subgenome 1 and Subgenome 2"
+  )
+
+ggplot(data, aes(subgenome,log10(FPKM_mean))) +
+  geom_boxplot() +
+  labs(y = "log10(FPKM)",
+       x = "Subgenome",
+       title = "Comparison of FPKM Expression between\nGenes in Subgenome 1 and Subgenome 2"
+  )
+```
+
 # Wrapup

### 2017-09-15 R/test_gene_transcripts.R

@@ -26,31 +26,44 @@ nrow(gene_transcripts_v4) - nrow(gene_transcripts_v4_gff)
 #==================================================================================================#
 ## geneTranscript.map and geneTranscript.counts
 #--------------------------------------------------------------------------------------------------#
-## Load the data
-gff_file <- system.file("extdata", "GFF3_files", "Zea_mays.AGPv4.32.gff3.gz", package="GenomicFeatures")
+# data <- geneTranscript.counts[geneTranscript.counts$gene %in% subgenome$gene2[subgenome$subgenome=="sub1"],]
+data <- geneTranscript.counts %>% inner_join(subgenome, by=c("gene"="gene2")) %>% select(gene, n, subgenome) %>% filter(subgenome=="sub1" | subgenome=="sub2")
+ggplot(data, aes(n,color=subgenome)) + geom_density() + scale_x_log10()
+ggplot(data, aes(n,color=subgenome)) + geom_histogram(bins = 50)
 
-## Create the txdb object
-txdb <- makeTxDbFromGFF("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\Zea_mays.AGPv4.32.gff3.gz", format="gff3")
+# data <- geneTranscript.counts[geneTranscript.counts$gene %in% subgenome$gene2[subgenome$subgenome=="sub2"],]
+# ggplot(data, aes(n)) + geom_bar()
 
-## Only work with chromosomes, ignore unplaced contigs
-seqlevels(txdb) <- c("1","2","3","4","5","6","7","8","9","10")
+#==================================================================================================#
+##
+#--------------------------------------------------------------------------------------------------#
+expressedGenes
+expressedPairs
 
-## Get gene/transcript names
-geneTranscript.map <- data.frame(transcripts(txdb)$tx_name)
-# GRList <- exonsBy(txdb, by = "tx")
-# tx_ids <- names(GRList)
-# head(select(txdb, keys=tx_ids, columns=c("GENEID","TXNAME"), keytype="TXID"))
+## Plot frequency of isoform counts for each gene
+data <- expressedGenes %>% inner_join(subgenome, by=c("geneID"="gene2")) %>% select(geneID, FPKM_mean, subgenome) %>% filter(subgenome=="sub1" | subgenome=="sub2")
[…]

### 2017-09-19 SubGenome Analysis.Rmd
Large knit speed improvement by storing cleaned data objects in file

@@
 ```{r importSynMapData, include=FALSE}
-source("~/git/SubGenomes/R/loadData.R")
-source("~/git/SubGenomes/R/cleanData.R")
+load("~/git/SubGenomes/Data/SavedObjects/loadedData.RData")
+
+# source("~/git/SubGenomes/R/loadData.R")
+# source("~/git/SubGenomes/R/cleanData.R")
 source("~/git/SubGenomes/R/functions.R")
@@
 ```{r subgenome_expression_frequency, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4.5, fig.height=4}
+# Paired t-test suggests that there is no difference in means between FPKM in sub1 vs. sub2:
+# t.test(x=expressedPairs$FPKM_mean1, y=expressedPairs$FPKM_mean2, paired = TRUE)
+
 ## Plot frequency of isoform counts for each gene

### 2017-09-19 R/cleanData.R

@@ -1,3 +1,4 @@
+library(GenomicFeatures)
 library(topGO)
 library(tidyr)
 library(dplyr)
@@ -15,6 +16,7 @@ startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
 ##        go.maize.raw
 ##        go.sorghum.raw
 ##        maize.genes.v3_to_v4_map.raw
+##        txdb
 ## Output:
 ##        syntelogs.sorghum.v1.maize.v1.clean
 ##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean
@@ -22,6 +24,7 @@ startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
 ##        go.maize.clean
 ##        go.sorghum.clean
 ##        maize.genes.v3_to_v4_map.clean
+##        geneTranscript.map
 ##
 ## Date: 2017-08-25
 ## Author: Jesse R. Walsh
@@ -133,8 +136,35 @@ maize.genes.v3_to_v4_map.clean <-
 ## v3 to v4 is 1 to many mapping
 # maize.genes.v3_to_v4_map.clean$v4_id[duplicated(maize.genes.v3_to_v4_map.clean$v4_id)]
 
+#==================================================================================================#
+## txdb -> geneTranscript.map
+#--------------------------------------------------------------------------------------------------#
+## Only work with chromosomes, ignore unplaced contigs
+seqlevels(txdb) <- c("1","2","3","4","5","6","7","8","9","10")
+
+## Get gene/transcript names
+geneTranscript.map <- data.frame(transcripts(txdb)$tx_name)
+# GRList <- exonsBy(txdb, by = "tx")
+# tx_ids <- names(GRList)
+# head(select(txdb, keys=tx_ids, columns=c("GENEID","TXNAME"), keytype="TXID"))
+
+## Clean geneTranscript.map
+geneTranscript.map <-
+  geneTranscript.map %>%
+  rename(transcript=transcripts.txdb..tx_name)
+geneTranscript.map$transcript <- sub("transcript:", "", geneTranscript.map$transcript)
+geneTranscript.map$gene <- sub("(Zm[0-9]{5}d[0-9]{6}).*", "\\1", geneTranscript.map$transcript)
+geneTranscript.map <- geneTranscript.map[!startsWith(geneTranscript.map$transcript, "MI"),]
+geneTranscript.counts <-
+  geneTranscript.map %>%
+  select(gene) %>%
[…]

### 2017-09-19 R/do.R

@@ -1,4 +1,3 @@
-library(GenomicFeatures)
 library(tidyr)
 library(dplyr)
 ####################################################################################################
@@ -72,9 +71,7 @@ expressedGenes <- data.frame(ID=maize.expression.clean[,1], Means=rowMeans(maize
 ## Remove rows with NA
 expressedGenes <-
   expressedGenes %>%
-  mutate(FPKM_mean = Means) %>%
-  select(tracking_id, FPKM_mean) %>%
-  rename(geneID = tracking_id) %>%
+  rename(geneID = tracking_id, FPKM_mean = Means) %>%
   subset(!is.na(FPKM_mean))
 
 ## Convert to v4 ids
@@ -83,6 +80,12 @@ expressedGenes <-
   inner_join(maize.genes.v3_to_v4_map.clean, by=c("geneID" = "v3_id")) %>%
   select(v4_id, FPKM_mean) %>%
   rename(geneID=v4_id)
+
+## Converting from v3 to v4 will give duplicate values (when gene models are merged, etc.), assume they are they same length so we can add their FPKM together
+expressedGenes <-
+  expressedGenes %>%
+  group_by(geneID) %>%
+  summarise(FPKM_mean=sum(FPKM_mean))
 
 ## Attach log2(FPKM_mean) values to homeologous pairs
 expressedPairs <-
@@ -104,32 +107,6 @@ expressedPairs <-
 #--------------------------------------------------------------------------------------------------#
 # source("~/git/SubGenomes/R/createTopGO.R")
 
-#==================================================================================================#
-## txdb -> geneTranscript.map
 #--------------------------------------------------------------------------------------------------#
-## Only work with chromosomes, ignore unplaced contigs
-seqlevels(txdb) <- c("1","2","3","4","5","6","7","8","9","10")
-
-## Get gene/transcript names
-geneTranscript.map <- data.frame(transcripts(txdb)$tx_name)
-# GRList <- exonsBy(txdb, by = "tx")
-# tx_ids <- names(GRList)
-# head(select(txdb, keys=tx_ids, columns=c("GENEID","TXNAME"), keytype="TXID"))
-
-## Clean geneTranscript.map
-geneTranscript.map <-
-  geneTranscript.map %>%
-  rename(transcript=transcripts.txdb..tx_name)
-geneTranscript.map$transcript <- sub("transcript:", "", geneTranscript.map$transcript)
[…]

### 2017-09-19 R/functions.R

@@ -1,5 +1,5 @@
 ####################################################################################################
-## Project:         Subgenomes project
+## Project:         SubGenomes project
 ## Script purpose:  Create functions to be sourced and used for this project
 ##
 ## Input:
@@ -7,3 +7,34 @@
 ## Date: 2017-09-14
 ## Author: Jesse R. Walsh
 ####################################################################################################
+
+saveMyObjects <- function() {
+  rm(list = ls())
+  source("~/git/SubGenomes/R/loadData.R")
+  source("~/git/SubGenomes/R/cleanData.R")
+  save.image("~/git/SubGenomes/Data/SavedObjects/loadedData.RData")
+  # save(syntelogs.sorghum.v1.maize.v1.clean, file = "~/git/SubGenomes/Data/SavedObjects/syntelogs.sorghum.v1.maize.v1.clean")
+  # save(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean, file = "~/git/SubGenomes/Data/SavedObjects/syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean")
+  # save(maize.expression.clean, file = "~/git/SubGenomes/Data/SavedObjects/maize.expression.clean")
+  # save(go.maize.clean, file = "~/git/SubGenomes/Data/SavedObjects/go.maize.clean")
+  # save(go.sorghum.clean, file = "~/git/SubGenomes/Data/SavedObjects/go.sorghum.clean")
+  # save(maize.genes.v3_to_v4_map.clean, file = "~/git/SubGenomes/Data/SavedObjects/maize.genes.v3_to_v4_map.clean")
+  # save(go.goSlim.plant, file = "~/git/SubGenomes/Data/SavedObjects/go.goSlim.plant")
+  # save(subgenome.assignments, file = "~/git/SubGenomes/Data/SavedObjects/subgenome.assignments")
+  # save(subgenome.truth, file = "~/git/SubGenomes/Data/SavedObjects/subgenome.truth")
+  # save(txdb, file = "~/git/SubGenomes/Data/SavedObjects/txdb")
+}
+
+loadMyObjects <- function() {
+  load("~/git/SubGenomes/Data/SavedObjects/loadedData.RData")
+  # load(file = "~/git/SubGenomes/Data/SavedObjects/syntelogs.sorghum.v1.maize.v1.clean")
[…]

### 2017-09-22 SubGenome Analysis.Rmd
Analysis of Subgenomes by expression data

--- 2017-09-15 SubGenome Analysis.Rmd
@@
 
+Expression data was retrieved from [NCBI: GSE50191](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE50191) and was reported in [Walley 2016](http://science.sciencemag.org/content/353/6301/814.full).
+This set includes 69 experiments (3 biological replicates, 23 tissues) for 62,547 mRNAs aligned to Zea mays B73 RefGen_v2 5a WGS annotation set.
+
 # Import SynMap and GO Term Data
@@
 
+Paired t-test suggests that there is no difference in means between FPKM of homeologs in sub1 vs. sub2 (p-val > 0.05, so we accept null hypothesis of equal means):
+```{r subgenome_fpkm_ttest, echo=FALSE}
+t.test(x=expressedPairs$FPKM_mean1, y=expressedPairs$FPKM_mean2, paired = TRUE)
+```
+
 ```{r subgenome_expression_frequency, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4.5, fig.height=4}
@@
 
+I am curious if our definition of sub 1 and sub 2 (i.e.
+the larger syntenic block is defined as sub 1) is really the best way to decide which genes are in sub 1 and sub 2.
+It seems to me that if one could resonably expect that, given 2 identically duplicate genes, if there was no benefit to having duplicated function then it would be a matter of random chance which of the two duplicate genes would loose function over time.
+It also seems to me that the quickest way to "degrade" an unnecessary gene (evolutionarily speaking) is to reduce expression.
+So what happens if we pick the higher expressing gene to be in sub 1 and the lower expressing gene to be in sub 2?
+Not surprising, we can find differences in populations between high and low expressing pairs.
+I wonder if this holds across different maize lines?
+
+```{r subgenome_expression_high_v_low, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4.5, fig.height=4, fig.align='center'}
+data <- expressedPairs
+data$LowExpressingGene <- apply(data[,c(3,4)], 1, FUN=min)
+data$HighExpressingGene <- apply(data[,c(3,4)], 1, FUN=max)
+data <- data %>% gather(key = type, value = value, 5:6)
+ggplot(data, aes(type, log10(value))) +
+  geom_boxplot() +
+  labs(y = "log10(FPKM)",
+       x = "Types",
+       title = "Comparison of FPKM Expression between\nGenes in high express vs. low express groups"
+  )
+```
+
+```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=4, fig.align='center'}
+plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 1, TRUE)
+# plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 4, FALSE)
+plot(plot)
+```
+
 # Wrapup

### 2017-09-22 R/do.R

@@ -64,6 +64,26 @@ maizeWithSorghumGO <-
   distinct()
 
 #==================================================================================================#
+## maize.expression.all
+#--------------------------------------------------------------------------------------------------#
+maize.expression.all <- maize.expression.clean
+
+## Convert to v4 ids
+maize.expression.all <-
+  maize.expression.all %>%
+  inner_join(maize.genes.v3_to_v4_map.clean, by=c("tracking_id" = "v3_id")) %>%
+  rename(geneID=v4_id)
+
+## Reorder columns
+maize.expression.all <- maize.expression.all[,c(70,2:69)]
+
+## Merge duplicate rows by adding FPKM values together
+maize.expression.all <-
+  maize.expression.all %>%
+  group_by(geneID) %>%
+  summarise_all(funs(sum))
+
+#==================================================================================================#
 ## parseExpressionData.R
 #--------------------------------------------------------------------------------------------------#
 expressedGenes <- data.frame(ID=maize.expression.clean[,1], Means=rowMeans(maize.expression.clean[,-1], na.rm = TRUE))

### 2017-09-22 R/functions.R

@@ -8,33 +8,90 @@
 ## Author: Jesse R. Walsh
 ####################################################################################################
 
+## Slicing data
+#--------------------------------------------------------------------------------------------------#
+getExpressionByExperiment <- function(maize.expression.all, homeologs.pairs, experiment) {
+  if (experiment <= 1 | experiment >= 70) {
+    return(NA)
+  }
+  expressedGenes <- maize.expression.all[,c(1,experiment)]
+  homeolog.expression <-
+    homeologs.pairs %>%
+    subset(Maize1 != "" & Maize2 != "") %>%
+    select(Maize1, Maize2) %>%
+    inner_join(expressedGenes, by=c("Maize1"="geneID")) %>%
+    inner_join(expressedGenes, by=c("Maize2"="geneID"))
+
+    names(homeolog.expression)[3] <- "FPKM_maize1"
+    names(homeolog.expression)[4] <- "FPKM_maize2"
+
+    return(homeolog.expression)
+}
+
+## experimentList = list of column positions from maize.expression.all (the dataset used currently has data in columns 2:69, but only about 10 display well at a time)
+## Factor = how many times greater expression has to be to be considered "dominant" -> (i.e. is A >= factor*B)
+## omitNA = if FALSE, NA values in the expression dataset for either homeolog result in a "Not Expressed",
+##          if TRUE, NA values are converted to zero's, so any expression for one homeolog will automatically dominate the other and NA in both results in a "Neither"
+##          this setting does not stop NAs in the case where the gene pair doesn't exist in the expression set
+graphGenePairExpressionsByExperiement <- function(maize.expression.all, homeologs.pairs, experimentList, factor, includeNA) {
+  genePairs <-
+    homeologs.pairs %>%
+    subset(Maize1 != "" & Maize2 != "") %>%
+    select(Maize1, Maize2) %>%
+    distinct()
+
+  nPairs <- nrow(genePairs)
+
+  for (i in experimentList) {
+    data <- getExpressionByExperiment(maize.expression.all, homeologs.pairs, i)
+    if (includeNA) {
[…]

### 2017-09-22 R/test_gene_transcripts.R

@@ -39,31 +39,27 @@ ggplot(data, aes(n,color=subgenome)) + geom_histogram(bins = 50)
 #--------------------------------------------------------------------------------------------------#
 expressedGenes
 expressedPairs
+getExpressionByExperiment(maize.expression.all, homeologs.pairs, 2)
 
-## Plot frequency of isoform counts for each gene
-data <- expressedGenes %>% inner_join(subgenome, by=c("geneID"="gene2")) %>% select(geneID, FPKM_mean, subgenome) %>% filter(subgenome=="sub1" | subgenome=="sub2")
-ggplot(data, aes(FPKM_mean,color=subgenome)) +
-  geom_density() +
-  scale_x_log10() +
-  labs(y = "Density (FPKM)",
-       x = "log10(FPKM)",
-       title = "Comparison of FPKM Expression between\nGenes in Subgenome 1 and Subgenome 2"
-  )
+data <- getExpressionByExperiment(maize.expression.all, homeologs.pairs, 12)
+data <- data[!is.na(data$FPKM_maize1) & !is.na(data$FPKM_maize2),]
+# data$maize1_Dom <- ifelse(data$FPKM_maize1 > 1*data$FPKM_maize2, TRUE, FALSE)
+# nrow(data[data$maize1_Dom,]) - nrow(data[!data$maize1_Dom,])
+# data$maize1_Dom <- ifelse(data$FPKM_maize1 > 2*data$FPKM_maize2, TRUE, FALSE)
+# nrow(data[data$maize1_Dom,]) - nrow(data[!data$maize1_Dom,])
+data$maize1_Dom <- ifelse(data$FPKM_maize1/data$FPKM_maize2 >= 4, TRUE, FALSE)
+data$maize2_Dom <- ifelse(data$FPKM_maize2/data$FPKM_maize1 >= 4, TRUE, FALSE)
+# data$maize1_Dom <- ifelse((data$FPKM_maize1-data$FPKM_maize2)/data$FPKM_maize2 >= 4, TRUE, FALSE)
+# data$maize2_Dom <- ifelse((data$FPKM_maize2-data$FPKM_maize1)/data$FPKM_maize1 >= 4, TRUE, FALSE)
+nrow(data[data$maize1_Dom,]) / nrow(data[data$maize1_Dom == FALSE & data$maize2_Dom == FALSE,]) * 100
[…]

### 2017-09-28 SubGenome Analysis.Rmd
Improved expression plot
Take mean of replicates for each sample, plot all samples instead of only a subset, replace tracking_ids with experiment names.

@@
 
-```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=4, fig.align='center'}
-plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 1, TRUE)
+```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center'}
+# plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 1, TRUE)
 # plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 4, FALSE)
+plot <- graphGenePairExpressionsByExperiement(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
 plot(plot)

### 2017-09-28 R/do.R

@@ -84,6 +84,33 @@ maize.expression.all <-
   summarise_all(funs(sum))
 
 #==================================================================================================#
+## maize.expression.sample.avg
+#--------------------------------------------------------------------------------------------------#
+maize.expression.sample.avg <- maize.expression.clean
+
+## Convert to v4 ids
+maize.expression.sample.avg <-
+  maize.expression.sample.avg %>%
+  inner_join(maize.genes.v3_to_v4_map.clean, by=c("tracking_id" = "v3_id")) %>%
+  rename(geneID=v4_id)
+
+## Reorder columns
+maize.expression.sample.avg <- maize.expression.sample.avg[,c(70,2:69)]
+
+## convert to sample names, merge replicates in each sample using mean, and output in long form
+maize.expression.sample.avg <-
+  maize.expression.sample.avg %>%
+  gather("tracking_id", "FPKM",-1) %>%
+  left_join(experiment.map, by=c("tracking_id"="tracking_id")) %>%
+  select(geneID, Sample, FPKM) %>%
+  group_by(geneID, Sample) %>%
+  summarise(FPKM_avg=mean(FPKM, na.rm=TRUE)) %>%
+  arrange(geneID)
+
+## When all replicates have NA, mean returns NaN.  Convert it back to NA.
+maize.expression.sample.avg$FPKM_avg[is.nan(maize.expression.sample.avg$FPKM_avg)] <- NA
+
+#==================================================================================================#
 ## parseExpressionData.R
 #--------------------------------------------------------------------------------------------------#
 expressedGenes <- data.frame(ID=maize.expression.clean[,1], Means=rowMeans(maize.expression.clean[,-1], na.rm = TRUE))

### 2017-09-28 R/functions.R

@@ -28,12 +28,68 @@ getExpressionByExperiment <- function(maize.expression.all, homeologs.pairs, exp
     return(homeolog.expression)
 }
 
+## maize.expression.sample.avg = long-form expression data
+## Factor = how many times greater expression has to be to be considered "dominant" -> (i.e. is A >= factor*B)
+## omitNA = if FALSE, NA values in the expression dataset for either homeolog result in a "Not Expressed",
+##          if TRUE, NA values are converted to zero's, so any expression for one homeolog will automatically dominate the other and NA in both results in a "Neither"
+##          this setting does not stop NAs in the case where the gene pair doesn't exist in the expression set
+graphGenePairExpressionsByExperiement <- function(maize.expression.sample.avg, homeologs.pairs, factor, includeNA) {
+  genePairs <-
+    homeologs.pairs %>%
+    subset(Maize1 != "" & Maize2 != "") %>%
+    select(Maize1, Maize2) %>%
+    distinct()
+
+  # nPairs <- nrow(genePairs)
+
+  data <-
+    genePairs %>%
+    inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
+    inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample"))
+  names(data)[4] <- "FPKM_maize1"
+  names(data)[5] <- "FPKM_maize2"
+
+  ## We only count gene pairs where both genes were at least tested in the expression set
+  nPairs <-
+    data %>%
+    select(Maize1, Maize2) %>%
+    distinct() %>%
+    nrow()
+
+  if (includeNA) {
+    data$FPKM_maize1[is.na(data$FPKM_maize1)] <- 0
+    data$FPKM_maize2[is.na(data$FPKM_maize2)] <- 0
+  }
+  data$dominance <- NA
+  data$dominance[is.na(data$FPKM_maize1) | is.na(data$FPKM_maize2)] <- "Not Expressed"
+  data$dominance[data$FPKM_maize1 >= factor*data$FPKM_maize2] <- "Maize1"
+  data$dominance[data$FPKM_maize2 >= factor*data$FPKM_maize1] <- "Maize2"
+  data$dominance[data$FPKM_maize1 <= factor*data$FPKM_maize2 & data$FPKM_maize2 <= factor*data$FPKM_maize1] <- "Neither"
+
+  data <-
+    data %>%
[…]

### 2017-09-28 R/loadData.R

@@ -36,6 +36,7 @@ syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw <- read_delim("./Data/SynMap/so
 
 ## Expression data from the Walley 2016 paper in FPKM over 68 tissues
 maize.expression.raw <- read_delim("./Data/Expression/GSE50191_FPKM.tsv", "\t", trim_ws = TRUE)
+experiment.map <- read_delim("./Data/Expression/tracking_ids.csv", ",", trim_ws = TRUE)
 
 ## Read in GO Annotation data for maize genes
 go.maize.raw <- read_delim("./Data/GO/go_from_ maizecyc.tab", "\t", escape_double = FALSE, trim_ws = TRUE)

### 2017-09-28 R/test_gene_transcripts.R

@@ -6,3 +6,23 @@ library(GenomicFeatures)
 library(tidyr)
 library(dplyr)
 
+
+
+## Fix experiment names
+mm <- match(names(maize.expression.avg), experiment.map$tracking_id)
+names(maize.expression.avg)[!is.na(mm)] <- as.character(experiment.map$Replicate[na.omit(mm)])
+
+
+maize.expression.sample.avg <- maize.expression.all
+maize.expression.sample.avg <-
+  maize.expression.sample.avg %>%
+  gather("tracking_id", "FPKM",-1) %>%
+  left_join(experiment.map, by=c("tracking_id"="tracking_id")) %>%
+  select(geneID, Sample, FPKM) %>%
+  group_by(geneID, Sample) %>%
+  summarise(FPKM_avg=mean(FPKM, na.rm=TRUE)) %>%
+  arrange(geneID)
+maize.expression.sample.avg$FPKM_avg[is.nan(maize.expression.sample.avg$FPKM_avg)] <- NA
+
+plot <- graphGenePairExpressionsByExperiement(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
+plot(plot)

### 2017-09-28 SubGenome Analysis.Rmd
Working on proteins

--- 2017-09-22 SubGenome Analysis.Rmd
@@
 This set includes 69 experiments (3 biological replicates, 23 tissues) for 62,547 mRNAs aligned to Zea mays B73 RefGen_v2 5a WGS annotation set.
+
+Protein abundance data was retrieved from [Walley 2016](http://science.sciencemag.org/content/353/6301/814.full) supplamental table S2.
+This dataset contains 3 to 7 biological replicates over 33 tissues.
+Expression data is based on a subset of this protein data tissues.
+Values are in dNASF, which is summarized [here](http://research.stowers.org/proteomics/Quant.html).
 
@@
 
-```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=4, fig.align='center'}
-plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 1, TRUE)
+```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center'}
+# plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 1, TRUE)
 # plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 4, FALSE)
+plot <- graphGenePairExpressionsByExperiement(maize.expression.sample.avg, homeologs.pairs, 2, TRUE)
 plot(plot)

### 2017-09-28 R/test_gene_transcripts.R

@@ -8,21 +8,42 @@ library(dplyr)
 
 
 
-## Fix experiment names
-mm <- match(names(maize.expression.avg), experiment.map$tracking_id)
-names(maize.expression.avg)[!is.na(mm)] <- as.character(experiment.map$Replicate[na.omit(mm)])
+# ## Fix experiment names
+# mm <- match(names(maize.expression.avg), experiment.map$tracking_id)
+# names(maize.expression.avg)[!is.na(mm)] <- as.character(experiment.map$Replicate[na.omit(mm)])
+#
+#
+# maize.expression.sample.avg <- maize.expression.all
+# maize.expression.sample.avg <-
+#   maize.expression.sample.avg %>%
+#   gather("tracking_id", "FPKM",-1) %>%
+#   left_join(experiment.map, by=c("tracking_id"="tracking_id")) %>%
+#   select(geneID, Sample, FPKM) %>%
+#   group_by(geneID, Sample) %>%
+#   summarise(FPKM_avg=mean(FPKM, na.rm=TRUE)) %>%
+#   arrange(geneID)
+# maize.expression.sample.avg$FPKM_avg[is.nan(maize.expression.sample.avg$FPKM_avg)] <- NA
+#
+# plot <- graphGenePairExpressionsByExperiement(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
+# plot(plot)
 
 
-maize.expression.sample.avg <- maize.expression.all
-maize.expression.sample.avg <-
-  maize.expression.sample.avg %>%
-  gather("tracking_id", "FPKM",-1) %>%
-  left_join(experiment.map, by=c("tracking_id"="tracking_id")) %>%
-  select(geneID, Sample, FPKM) %>%
-  group_by(geneID, Sample) %>%
-  summarise(FPKM_avg=mean(FPKM, na.rm=TRUE)) %>%
-  arrange(geneID)
-maize.expression.sample.avg$FPKM_avg[is.nan(maize.expression.sample.avg$FPKM_avg)] <- NA
+library(readxl)
+maize.protein.abundance.raw <- read_xlsx("~/git/SubGenomes/Data/ProteinAbundance/aag1125_SupportingFile_Table_S2-1.xlsx", sheet = 3, col_names = TRUE)
+maize.protein.abundance.clean <- maize.protein.abundance.raw
+maize.protein.abundance.clean <-
+  maize.protein.abundance.clean %>%
[…]

### 2017-09-29 SubGenome Analysis.Rmd
Completed protein dominance analysis
Also have several small fixes, mostly to comments.  Moved gff file into its own folder and updated load path.

@@
 
+# Dominance
+
+***
+
+In all cases where there is a homeolog on both maize1 and maize2, the we want to look to several measures to see if the genes on one subgenome tend to dominate the other.
+We can define dominance using a factor (i.e.
+4-fold higher expression/abundance etc.).
+
+First we look at expression values.
+We find that maize1 consistently dominates the maize2 pair across all available data, regardless of factor (1,2, or 4), and regardless of whether or not we allow NA values to be treated as '0'.
+
 ```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center'}
-# plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 1, TRUE)
-# plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 4, FALSE)
-plot <- graphGenePairExpressionsByExperiement(maize.expression.sample.avg, homeologs.pairs, 2, TRUE)
+plot <- graphGenePairDominanceByExperiment(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
+plot(plot)
+```
+
+This same dominance patter holds for protein abundance data as well. Here we look at protein abundance data for the tissues that also have expression data.
+
+```{r subgenome_protein_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center'}
+plot <- graphGenePairDominanceByExperiment(maize.protein.abundance.sample.avg, homeologs.pairs, 4, FALSE)
+plot + labs(title = paste0("Which homeolog dominates protein abundance? (factor=",4,")"))
 plot(plot)

### 2017-09-29 R/cleanData.R

@@ -17,6 +17,7 @@ startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
 ##        go.sorghum.raw
 ##        maize.genes.v3_to_v4_map.raw
 ##        txdb
+##        maize.protein.abundance.raw
 ## Output:
 ##        syntelogs.sorghum.v1.maize.v1.clean
 ##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean
@@ -25,6 +26,7 @@ startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
 ##        go.sorghum.clean
 ##        maize.genes.v3_to_v4_map.clean
 ##        geneTranscript.map
+##        maize.protein.abundance.clean
 ##
 ## Date: 2017-08-25
 ## Author: Jesse R. Walsh
@@ -56,8 +58,21 @@ syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene1 <- gsub(syntelogs.sorgh
 #--------------------------------------------------------------------------------------------------#
 maize.expression.clean <- maize.expression.raw
 
-## Remove low FPKM values, calculate row means
+## Remove low FPKM values
 maize.expression.clean[maize.expression.clean < 1] <- NA
+
+#==================================================================================================#
+## maize.protein.abundance.raw
+#--------------------------------------------------------------------------------------------------#
+maize.protein.abundance.clean <- maize.protein.abundance.raw
+
+## Rename the gene id column
+maize.protein.abundance.clean <-
+  maize.protein.abundance.clean %>%
+  rename(v3_id=X__1)
+
+## Remove low dNSAF values
+maize.protein.abundance.clean[maize.protein.abundance.clean < 1] <- NA
 
 #==================================================================================================#
 ## go.maize.raw

### 2017-09-29 R/do.R

@@ -111,6 +111,37 @@ maize.expression.sample.avg <-
 maize.expression.sample.avg$FPKM_avg[is.nan(maize.expression.sample.avg$FPKM_avg)] <- NA
 
 #==================================================================================================#
+## maize.protein.abundance.sample.avg
+#--------------------------------------------------------------------------------------------------#
+maize.protein.abundance.sample.avg <- maize.protein.abundance.clean
+
+## Convert to v4 ids
+maize.protein.abundance.sample.avg <-
+  maize.protein.abundance.sample.avg %>%
+  inner_join(maize.genes.v3_to_v4_map.clean, by=c("v3_id" = "v3_id")) %>%
+  rename(geneID=v4_id)
+
+## Reorder columns
+maize.protein.abundance.sample.avg <- maize.protein.abundance.sample.avg[,c(length(maize.protein.abundance.sample.avg),3:length(maize.protein.abundance.sample.avg)-1)]
+
+## Reduce columns to only the ones comparable to expression data
+colsToKeep <- colnames(maize.protein.abundance.sample.avg) %in% c("geneID",experiment.map.proteins$Replicate[!is.na(experiment.map.proteins$ExpressionSampleName)])
+maize.protein.abundance.sample.avg <- maize.protein.abundance.sample.avg[,colsToKeep]
+
+## convert to sample names, merge replicates in each sample using mean, and output in long form
+maize.protein.abundance.sample.avg <-
+  maize.protein.abundance.sample.avg %>%
+  gather("Replicate", "dNSAF",-1) %>%
+  left_join(experiment.map.proteins, by=c("Replicate"="Replicate")) %>%
+  select(geneID, Sample, dNSAF) %>%
+  group_by(geneID, Sample) %>%
+  summarise(dNSAF_avg=mean(dNSAF, na.rm=TRUE)) %>%
+  arrange(geneID)
+
+## When all replicates have NA, mean returns NaN.  Convert it back to NA.
+maize.protein.abundance.sample.avg$dNSAF_avg[is.nan(maize.protein.abundance.sample.avg$dNSAF_avg)] <- NA
+
+#==================================================================================================#
 ## parseExpressionData.R
 #--------------------------------------------------------------------------------------------------#
 expressedGenes <- data.frame(ID=maize.expression.clean[,1], Means=rowMeans(maize.expression.clean[,-1], na.rm = TRUE))

### 2017-09-29 R/functions.R

@@ -10,46 +10,44 @@
 
 ## Slicing data
 #--------------------------------------------------------------------------------------------------#
-getExpressionByExperiment <- function(maize.expression.all, homeologs.pairs, experiment) {
-  if (experiment <= 1 | experiment >= 70) {
-    return(NA)
-  }
-  expressedGenes <- maize.expression.all[,c(1,experiment)]
-  homeolog.expression <-
-    homeologs.pairs %>%
-    subset(Maize1 != "" & Maize2 != "") %>%
-    select(Maize1, Maize2) %>%
-    inner_join(expressedGenes, by=c("Maize1"="geneID")) %>%
-    inner_join(expressedGenes, by=c("Maize2"="geneID"))
+# getExpressionByExperiment <- function(maize.expression.all, homeologs.pairs, experiment) {
+#   if (experiment <= 1 | experiment >= 70) {
+#     return(NA)
+#   }
+#   expressedGenes <- maize.expression.all[,c(1,experiment)]
+#   homeolog.expression <-
+#     homeologs.pairs %>%
+#     subset(Maize1 != "" & Maize2 != "") %>%
+#     select(Maize1, Maize2) %>%
+#     inner_join(expressedGenes, by=c("Maize1"="geneID")) %>%
+#     inner_join(expressedGenes, by=c("Maize2"="geneID"))
+#
+#     names(homeolog.expression)[3] <- "FPKM_maize1"
+#     names(homeolog.expression)[4] <- "FPKM_maize2"
+#
+#     return(homeolog.expression)
+# }
 
-    names(homeolog.expression)[3] <- "FPKM_maize1"
-    names(homeolog.expression)[4] <- "FPKM_maize2"
-
-    return(homeolog.expression)
-}
-
-## maize.expression.sample.avg = long-form expression data
+## maize.data = long-form expression/protein data
 ## Factor = how many times greater expression has to be to be considered "dominant" -> (i.e. is A >= factor*B)
 ## omitNA = if FALSE, NA values in the expression dataset for either homeolog result in a "Not Expressed",
 ##          if TRUE, NA values are converted to zero's, so any expression for one homeolog will automatically dominate the other and NA in both results in a "Neither"
 ##          this setting does not stop NAs in the case where the gene pair doesn't exist in the expression set
[…]

### 2017-09-29 R/loadData.R

@@ -9,6 +9,9 @@ library(GenomicFeatures)
 ##        syntelogs.sorghum.v1.maize.v1.raw
 ##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw
 ##        maize.expression.raw
+##        experiment.map
+##        maize.expression.raw
+##        experiment.map.proteins
 ##        go.maize.raw
 ##        go.sorghum.raw
 ##        go.goSlim.plant
@@ -34,9 +37,13 @@ library(GenomicFeatures)
 syntelogs.sorghum.v1.maize.v1.raw <- read_delim("./Data/SynMap/sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
 syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw <- read_delim("./Data/SynMap/sorghum_v3.1_vs_maize_v4+rejected.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
 
-## Expression data from the Walley 2016 paper in FPKM over 68 tissues
+## Expression data from the Walley 2016 paper in FPKM for 23 tissues
 maize.expression.raw <- read_delim("./Data/Expression/GSE50191_FPKM.tsv", "\t", trim_ws = TRUE)
 experiment.map <- read_delim("./Data/Expression/tracking_ids.csv", ",", trim_ws = TRUE)
+
+## Protein data from the Walley 2016 paper in dNSAF for 33 tissues
+maize.protein.abundance.raw <- read_xlsx("~/git/SubGenomes/Data/ProteinAbundance/aag1125_SupportingFile_Table_S2-1.xlsx", sheet = 3, col_names = TRUE)
+experiment.map.proteins <- read_delim("./Data/ProteinAbundance/experiment_map_proteins.csv", ",", trim_ws = TRUE)
 
 ## Read in GO Annotation data for maize genes
 go.maize.raw <- read_delim("./Data/GO/go_from_ maizecyc.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
@@ -63,7 +70,7 @@ subgenome.truth <- setNames(data.frame(
 maize.genes.v3_to_v4_map.raw <- read_xlsx("./Data/MaizeGDB_v3_v4.genes.xlsx")
 
 ## Load maize GFF data
-txdb <- makeTxDbFromGFF("./Data/Zea_mays.AGPv4.32.gff3.gz", format="gff3")
[…]

### 2017-09-29 R/test_gene_transcripts.R

@@ -28,22 +28,6 @@ library(dplyr)
 # plot(plot)
 
 
-library(readxl)
-maize.protein.abundance.raw <- read_xlsx("~/git/SubGenomes/Data/ProteinAbundance/aag1125_SupportingFile_Table_S2-1.xlsx", sheet = 3, col_names = TRUE)
-maize.protein.abundance.clean <- maize.protein.abundance.raw
-maize.protein.abundance.clean <-
-  maize.protein.abundance.clean %>%
-  select(1:9, 25:36, 41:50, 55:60, 68:85, 90:116, 121:136, 141:144) %>%
-  rename(v3_id=X__1)
 
-maize.protein.abundance.clean <-
-  maize.protein.abundance.clean %>%
-  inner_join(maize.genes.v3_to_v4_map.clean, by=c("v3_id" = "v3_id")) %>%
-  rename(geneID=v4_id)
-
-maize.protein.abundance.clean <- maize.protein.abundance.clean[,c(103,2:102)]
-
-## Rejected: no match in expression dataset
-names(maize.protein.abundance.clean %>% select(-c(1:9, 25:36, 41:50, 55:60, 68:85, 90:116, 121:136, 141:144)))
 
 

### 2017-09-29 SubGenome Analysis.Rmd
Isoform dominance plot

@@
 
-# source("~/git/SubGenomes/R/parseSubgenomes.R")
-# source("~/git/SubGenomes/R/addGOTerms.R")
-# source("~/git/SubGenomes/R/parseGOSlim.R")
-# source("~/git/SubGenomes/R/createTopGO.R")
-# source("~/git/SubGenomes/R/parseExpressionData.R")
-
-# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\loadData.R")
-# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\cleanData.R")
-# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseSubgenomes.R")
-# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\addGOTerms.R")
-# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseGOSlim.R")
-# # source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\createTopGO.R")
-# source("C:\\Users\\Jesse\\Documents\\git\\SubGenome\\R\\parseExpressionData.R")
-
 library(topGO)
@@
 library(gridExtra)
-
-# topGO::groupGOTerms()
-# MFterms <- ls(GOMFTerm)
-# BPterms <- ls(GOBPTerm)
-# CCterms <- ls(GOCCTerm)
 ```
@@
 
+***
+
 This same dominance patter holds for protein abundance data as well. Here we look at protein abundance data for the tissues that also have expression data.
@@
 ```{r subgenome_protein_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center'}
-plot <- graphGenePairDominanceByExperiment(maize.protein.abundance.sample.avg, homeologs.pairs, 4, FALSE)
-plot + labs(title = paste0("Which homeolog dominates protein abundance? (factor=",4,")"))
+factor <- 4
+plot <- graphGenePairDominanceByExperiment(maize.protein.abundance.sample.avg, homeologs.pairs, factor, FALSE) + 
+  labs(title = paste0("Which homeolog dominates protein abundance? (factor=",factor,")"))
+plot(plot)
+```
+
+***
+
+Isoforms also follows this dominance patter.
+
+```{r subgenome_isoform_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=6, fig.height=4, fig.align='center'}
+## Plot frequency of isoform counts for each gene
+plot <- graphGenePairDominanceByIsoForms(geneTranscript.counts, homeologs.pairs, 4)
 plot(plot)
@@
 ***
-
-The results of this report are available at **`r params$outputFile`**.
-
-```{r writeSubgenomesToFile, echo=FALSE}
-## Now recover which genes are in these sets
-#write.table(subgenome, params$outputFile, sep="\t")
-```

### 2017-09-29 R/functions.R

@@ -82,6 +82,39 @@ graphGenePairDominanceByExperiment <- function(maize.data, homeologs.pairs, fact
   return(plot)
 }
 
+## geneTranscript.counts = # of isoforms per gene
+## Factor = how many times greater expression has to be to be considered "dominant" -> (i.e. is A >= factor*B)
+graphGenePairDominanceByIsoForms <- function(geneTranscript.counts, homeologs.pairs, factor) {
+  data <-
+    homeologs.pairs %>%
+    subset(Maize1 != "" & Maize2 != "") %>%
+    select(Maize1, Maize2) %>%
+    distinct() %>%
+    inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
+    inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
+  names(data)[3] <- "Value_maize1"
+  names(data)[4] <- "Value_maize2"
+  data$dominance <- NA
+  data$dominance[is.na(data$Value_maize1) | is.na(data$Value_maize2)] <- "Not Data"
+  data$dominance[data$Value_maize1 >= factor*data$Value_maize2] <- "Maize1"
+  data$dominance[data$Value_maize2 >= factor*data$Value_maize1] <- "Maize2"
+  data$dominance[data$Value_maize1 <= factor*data$Value_maize2 & data$Value_maize2 <= factor*data$Value_maize1] <- "Neither"
+
+  data <-
+    data %>%
+    count(dominance)
+
+  plot <-
+    ggplot(data, aes(dominance, n)) +
+    geom_bar(aes(fill=dominance), position="dodge", stat="identity") +
+    labs(y = paste0("Number of pairs dominated (counts)"),
+         x = "Dominance",
+         title = paste0("Which homeolog dominates isoform counts? (factor=",factor,")")
+    ) +
+    geom_text(aes(label=n), vjust=-0.3, size=3.5)
+
+  return(plot)
+}
 
 ## Utility Functions
 #--------------------------------------------------------------------------------------------------#

### 2017-09-29 R/test_gene_transcripts.R

@@ -26,8 +26,3 @@ library(dplyr)
 #
 # plot <- graphGenePairExpressionsByExperiement(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
 # plot(plot)
-
-
-
-
-

### 2017-10-04 SubGenome Analysis.Rmd
Minor update to published pages
Grid plot for pair dominance across factors
Separated out data from plot functions for generating the genePairDominanceByExperiment().

@@
 
-# Dominance
+# Compare Homeolog Pairs
 
@@
 
-In all cases where there is a homeolog on both maize1 and maize2, the we want to look to several measures to see if the genes on one subgenome tend to dominate the other.
-We can define dominance using a factor (i.e.
+In all cases where there is a homeolog on both maize1 and maize2, the we want to look to several measures to see if the genes on one subgenome tend to be present in greater abundance compared to the other.
+We can define this comparison as either greater expression, greater protein abundance, or a greater number of isoforms available to the gene.
+We can compare using a factor to define a cutoff for "greater".
+(i.e.
 4-fold higher expression/abundance etc.).
@@
 First we look at expression values.
-We find that maize1 consistently dominates the maize2 pair across all available data, regardless of factor (1,2, or 4), and regardless of whether or not we allow NA values to be treated as '0'.
+We find that maize1 consistently overexpresses the maize2 pair on average across all available data, regardless of factor (1,2, or 4), and regardless of whether or not we allow NA values to be treated as '0'.
 
@@
 
-This same dominance patter holds for protein abundance data as well. Here we look at protein abundance data for the tissues that also have expression data.
+This same pattern holds for protein abundance data as well. Here we look at protein abundance data for the tissues that also have expression data.
 
@@
 plot <- graphGenePairDominanceByExperiment(maize.protein.abundance.sample.avg, homeologs.pairs, factor, FALSE) + 
-  labs(title = paste0("Which homeolog dominates protein abundance? (factor=",factor,")"))
+  labs(title = paste0("Which homeolog has greater protein abundance? (factor=",factor,")"))
+plot(plot)
+```
+
+We can see this same trend for all experiments across all factors (fold-difference of 1 to 9 shown).
+The blue line (neither gene has greater expression) grows as factor increases since we have a stricter and stricter definition of what counts as greater expression.
+Red (maize1) is always above green (maize2) indicating that more maize1 homeologs have greater expression over their maize2 pair than vice versa for all expriments and all factors.
+Red and green lines both tend to zero when factor is large enough.
+
+```{r subgenome_expression_dominance_across_factors, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=12, fig.align='center'}
+plot <- graphGenePairDominanceByExperimentAcrossFactors(maize.expression.sample.avg, homeologs.pairs, FALSE)
 plot(plot)
@@
 
-Isoforms also follows this dominance patter.
+The number of isoforms each gene of the homeolog pairs is thought to express follows a pattern where maize1 tends to have more isoforms than maize2 homeologs.
 
@@
 ***
+Report generated on: `r as.character(format(Sys.Date(), format="%B %d, %Y"))`

### 2017-10-04 R/functions.R

@@ -15,6 +15,39 @@
 ## omitNA = if FALSE, NA values in the expression dataset for either homeolog result in a "Not Expressed",
 ##          if TRUE, NA values are converted to zero's, so any expression for one homeolog will automatically dominate the other and NA in both results in a "Neither"
 ##          this setting does not stop NAs in the case where the gene pair doesn't exist in the expression set
+dataGenePairDominanceByExperiment <- function(maize.data, homeologs.pairs, factor, includeNA) {
+  genePairs <-
+    homeologs.pairs %>%
+    subset(Maize1 != "" & Maize2 != "") %>%
+    select(Maize1, Maize2) %>%
+    distinct()
+
+  data <-
+    genePairs %>%
+    inner_join(maize.data, by=c("Maize1"="geneID")) %>%
+    inner_join(maize.data, by=c("Maize2"="geneID", "Sample"="Sample"))
+  names(data)[4] <- "Value_maize1"
+  names(data)[5] <- "Value_maize2"
+
+  if (includeNA) {
+    data$Value_maize1[is.na(data$Value_maize1)] <- 0
+    data$Value_maize2[is.na(data$Value_maize2)] <- 0
+  }
+  data$dominance <- NA
+  data$dominance[is.na(data$Value_maize1) | is.na(data$Value_maize2)] <- "Not Expressed"
+  data$dominance[!is.na(data$Value_maize1) & !is.na(data$Value_maize2) & data$Value_maize1 >= factor*data$Value_maize2] <- "Maize1"
+  data$dominance[!is.na(data$Value_maize1) & !is.na(data$Value_maize2) & data$Value_maize2 >= factor*data$Value_maize1] <- "Maize2"
+  data$dominance[!is.na(data$Value_maize1) & !is.na(data$Value_maize2) & data$Value_maize1 <= factor*data$Value_maize2 & data$Value_maize2 <= factor*data$Value_maize1] <- "Neither"
+
+  data <-
+    data %>%
+    select(Sample, dominance) %>%
+    group_by(Sample, dominance) %>%
+    count()
+
+  return(data)
+}
+
 graphGenePairDominanceByExperiment <- function(maize.data, homeologs.pairs, factor, includeNA) {
   genePairs <-
     homeologs.pairs %>%
@@ -36,31 +69,65 @@ graphGenePairDominanceByExperiment <- function(maize.data, homeologs.pairs, fact
     distinct() %>%
     nrow()
 
-  if (includeNA) {
[…]

### 2017-10-04 R/staging_testing.R

@@ -7,23 +7,3 @@ library(tidyr)
 library(dplyr)
 
 
-
-# ## Fix experiment names
-# mm <- match(names(maize.expression.avg), experiment.map$tracking_id)
-# names(maize.expression.avg)[!is.na(mm)] <- as.character(experiment.map$Replicate[na.omit(mm)])
-#
-#
-# maize.expression.sample.avg <- maize.expression.all
-# maize.expression.sample.avg <-
-#   maize.expression.sample.avg %>%
-#   gather("tracking_id", "FPKM",-1) %>%
-#   left_join(experiment.map, by=c("tracking_id"="tracking_id")) %>%
-#   select(geneID, Sample, FPKM) %>%
-#   group_by(geneID, Sample) %>%
-#   summarise(FPKM_avg=mean(FPKM, na.rm=TRUE)) %>%
-#   arrange(geneID)
-# maize.expression.sample.avg$FPKM_avg[is.nan(maize.expression.sample.avg$FPKM_avg)] <- NA
-#
-# plot <- graphGenePairExpressionsByExperiement(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
-# plot(plot)
-

### 2017-10-05 SubGenome Analysis.Rmd
Top lists for different comparisons and links to relevant pages

@@
 Red and green lines both tend to zero when factor is large enough.
+Protein data (not shown) follows this same pattern.
 
@@
 
+Looking at a few individual cases (* these are calculated as maize1 > maize2, since maize2 > maize1 makes a negative fold change):
+
+```{r subgenome_expression_top_ten, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
+gene.foldchange <- homeologFoldChanges(maize.expression.sample.avg, homeologs.pairs)
+gene.foldchange$Maize1_url <- apply(gene.foldchange[,1], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
+gene.foldchange$Maize2_url <- apply(gene.foldchange[,2], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
+gene.foldchange[1:10,] %>% 
+  select(Maize1_url, Maize2_url, Sample, Value_maize1, Value_maize2, foldChange) %>%
+  knitr::kable(caption = "Top ten differentially expressed gene pairs (in log2)", align = "llrrr")
+
+protein.foldchange <- homeologFoldChanges(maize.protein.abundance.sample.avg, homeologs.pairs)
+protein.foldchange$Maize1_url <- apply(protein.foldchange[,1], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
+protein.foldchange$Maize2_url <- apply(protein.foldchange[,2], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
+protein.foldchange[1:10,] %>% 
+  select(Maize1_url, Maize2_url, Sample, Value_maize1, Value_maize2, foldChange) %>%
+  knitr::kable(caption = "Top ten differentially abundant protein pairs (in log2)", align = "llrrr")
+
+gene.foldchange[1:50,] %>%
+  inner_join(protein.foldchange[1:50,], by = c("Maize1", "Maize2", "Sample")) %>%
+  rename(foldChange_expr=foldChange.x, foldChange_prot=foldChange.y, Maize1_url=Maize1_url.x, Maize2_url=Maize2_url.x) %>%
+  select(Maize1_url, Maize2_url, Sample, foldChange_expr, foldChange_prot) %>%
+  knitr::kable(caption = "Gene pairs in the top 50 for both differential gene expression and differential protein abundance (in log2).", align = "llrrrr")
+```
+
+Look at the top 10 pairs with highest isoform count diffs (foldChange in log2)
+
+```{r subgenome_isoforms_top_ten, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
+  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
+names(data)[3] <- "Value_maize1"
+names(data)[4] <- "Value_maize2"
+data <- subset(data, !is.na(data$Value_maize1) & !is.na(data$Value_maize2))
+
+nSingleExonGenes.total <- nrow(geneTranscript.counts[geneTranscript.counts$n == 1,])
+nSingleExonGenes.inSet <- unique(c(data$Maize1[data$Value_maize1 == 1], data$Maize2[data$Value_maize2 == 1]))
+
+data <- subset(data, data$Value_maize1 > 1 & data$Value_maize2 > 1) ## Ignore single-exon genes
+data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)
+data <- arrange(data, desc(foldChange))
+data$Maize1_url <- apply(data[,1], 2, function(x) { wrapInHREF(getGBrowsePageURL(x), x) })
+data$Maize2_url <- apply(data[,2], 2, function(x) { wrapInHREF(getGBrowsePageURL(x), x) })
+data[1:10,] %>%
+  select(Maize1_url, Maize2_url, Value_maize1, Value_maize2, foldChange) %>%
+  knitr::kable(caption = "Top 10 pairs with highest isoform count diffs (foldChange in log2)", align = "llrrr")
+```
+
 # Wrapup

### 2017-10-05 R/do.R

@@ -128,14 +128,15 @@ maize.protein.abundance.sample.avg <- maize.protein.abundance.sample.avg[,c(leng
 colsToKeep <- colnames(maize.protein.abundance.sample.avg) %in% c("geneID",experiment.map.proteins$Replicate[!is.na(experiment.map.proteins$ExpressionSampleName)])
 maize.protein.abundance.sample.avg <- maize.protein.abundance.sample.avg[,colsToKeep]
 
-## convert to sample names, merge replicates in each sample using mean, and output in long form
+## convert to sample names using the same terms used in expression set, merge replicates in each sample using mean, and output in long form
 maize.protein.abundance.sample.avg <-
   maize.protein.abundance.sample.avg %>%
   gather("Replicate", "dNSAF",-1) %>%
   left_join(experiment.map.proteins, by=c("Replicate"="Replicate")) %>%
-  select(geneID, Sample, dNSAF) %>%
-  group_by(geneID, Sample) %>%
+  select(geneID, ExpressionSampleName, dNSAF) %>%
+  group_by(geneID, ExpressionSampleName) %>%
   summarise(dNSAF_avg=mean(dNSAF, na.rm=TRUE)) %>%
+  rename(Sample=ExpressionSampleName) %>%
   arrange(geneID)
 
 ## When all replicates have NA, mean returns NaN.  Convert it back to NA.

### 2017-10-05 R/functions.R

@@ -165,6 +165,29 @@ graphGenePairDominanceByIsoForms <- function(geneTranscript.counts, homeologs.pa
   return(plot)
 }
 
+
+homeologFoldChanges <- function(maize.data, homeologs.pairs) {
+  genePairs <-
+    homeologs.pairs %>%
+    subset(Maize1 != "" & Maize2 != "") %>%
+    select(Maize1, Maize2) %>%
+    distinct()
+
+  data <-
+    genePairs %>%
+    inner_join(maize.data, by=c("Maize1"="geneID")) %>%
+    inner_join(maize.data, by=c("Maize2"="geneID", "Sample"="Sample"))
+  names(data)[4] <- "Value_maize1"
+  names(data)[5] <- "Value_maize2"
+
+  ## Remove NAs and calculate foldchange
+  data <- subset(data, !is.na(data$Value_maize1) & !is.na(data$Value_maize2))
+  data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)
+  data <- arrange(data, desc(foldChange))
+  return(data)
+}
+
+
 ## Utility Functions
 #--------------------------------------------------------------------------------------------------#
 saveMyObjects <- function() {
@@ -177,4 +200,20 @@ saveMyObjects <- function() {
 
 loadMyObjects <- function() {
   load("~/git/SubGenomes/Data/SavedObjects/loadedData.RData", .GlobalEnv)
+}
+
+getGenePageURL <- function(geneID) {
+  base <- "https://maizegdb.org/gene_center/gene/"
+  return(paste0(base, geneID))
+}
+
+getGBrowsePageURL <- function(geneID) {
+  base <- "https://www.maizegdb.org/gbrowse/maize_v4/?name="
+  return(paste0(base, geneID))
+}
+
+
+## Example: apply(gene.foldchange[,1], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
+wrapInHREF <- function(url, text) {
+  return(paste0("<a href=\"", url, "\">", text, "</a>"))
 }

### 2017-10-05 R/staging_testing.R
Testing out commandline access to alingment tools.

@@ -1,9 +1,45 @@
-source("https://bioconductor.org/biocLite.R")
-biocLite("GenomicFeatures")
+# source("https://bioconductor.org/biocLite.R")
+# biocLite("GenomicFeatures")
 
-library(readr)
-library(GenomicFeatures)
+# library(readr)
+# library(GenomicFeatures)
 library(tidyr)
 library(dplyr)
 
 
+cmdBlast <- function(query, subject){
+  paste0("blastn -query <(", query, ") -subject <(", subject, ")")
+}
+
+cmdNeedle <- function(query, subject){
+  paste0("needle <(echo ", query, ") <(echo ", subject, ") -gapopen 10 -gapextend 0.5 -outfile >(cat)")
+}
+
+getSequence <- function(geneID){
+
+}
+
+
+
+# ---
+
+graphGenePairDominanceByIsoForms(geneTranscript.counts, homeologs.pairs, 4)
+
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
+  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
+names(data)[3] <- "Value_maize1"
+names(data)[4] <- "Value_maize2"
+data <- subset(data, !is.na(data$Value_maize1) & !is.na(data$Value_maize2))
+
+nSingleExonGenes.total <- nrow(geneTranscript.counts[geneTranscript.counts$n == 1,])
+nSingleExonGenes.inSet <- unique(c(data$Maize1[data$Value_maize1 == 1], data$Maize2[data$Value_maize2 == 1]))
+
+data <- subset(data, data$Value_maize1 > 1 & data$Value_maize2 > 1) ## Ignore single-exon genes
+data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)
+data <- arrange(data, desc(foldChange))
+data[1:10,]

### 2017-10-13 SubGenome Analysis.Rmd
Added scatterplots

@@
 
+Here is a scatter plot of isoform counts across gene pairs.
+
+```{r subgenome_isoforms_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.align='center'}
+# Isoforms
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
+  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
+names(data)[3] <- "Value_maize1"
+names(data)[4] <- "Value_maize2"
+# data <- subset(data, data$Value_maize1 > 1 & data$Value_maize2 > 1) ## Ignore single-exon genes
+data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)
+
+n <-
+  data %>%
+  subset(Value_maize1 > 100 | Value_maize2 > 100) %>%
+  nrow()
+
+plot <-
+  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
+  geom_point(alpha=.1, aes(color = abs(foldChange))) +
+  scale_color_gradient(low = "blue", high = "red") +
+  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  xlim(0,100) +
+  ylim(0,100) +
+  labs(
+    title = paste0("Scatterplot of isoform counts for gene pair(s).\nExcludes ",n," outlier, alpha=.1, foldchange in log2."),
+    x = "Isoform Counts for Maize1 Gene",
+    y = "Isoform Counts for Maize2 Gene"
+  )
+plot(plot)
+```
+
+```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "Value_maize1"
+names(data)[5] <- "Value_maize2"
+# data <- subset(data, Sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))
+
+plot <-
+  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
+  geom_point(alpha=.1, aes(color = Sample)) +
+  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  scale_x_log10() + scale_y_log10() +
+  labs(
+    title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1"),
+    x = "Expression for Maize1 Gene (log scale)",
+    y = "Expression for Maize2 Gene (log scale)"
+  ) +
+  theme(legend.position = "none") + facet_wrap(~Sample, nrow=6)
+plot(plot)
+```
+
+```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+names(data)[4] <- "Value_maize1"
+names(data)[5] <- "Value_maize2"
+# data <- subset(data, Sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))
+
+plot <-
+  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
+  geom_point(alpha=.1, aes(color = Sample)) +
+  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  scale_x_log10() + scale_y_log10() +
+  labs(
+    title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1"),
+    x = "Abundance for Maize1 Gene (log scale)",
+    y = "Abundance for Maize2 Gene (log scale)"
+  ) +
+  theme(legend.position = "none") + facet_wrap(~Sample, nrow=6)
+plot(plot)
+```
 # Wrapup

### 2017-10-13 R/staging_testing.R

@@ -5,7 +5,7 @@
 # library(GenomicFeatures)
 library(tidyr)
 library(dplyr)
-
+library(ggplot2)
 
 cmdBlast <- function(query, subject) {
   paste0("blastn -query <(", query, ") -subject <(", subject, ")")
@@ -31,55 +31,3 @@ subject <- "ATGGAGATCCTCGAATCGACCCTGTTGGGCGAGTTCATCGGCTTCATCAAGGGGAACTGGTCAGCGCA
 cmdNeedle(query, subject)
 
 # ---
-
-graphGenePairDominanceByIsoForms(geneTranscript.counts, homeologs.pairs, 4)
-
-data <-
-  homeologs.pairs %>%
-  subset(Maize1 != "" & Maize2 != "") %>%
-  select(Maize1, Maize2) %>%
-  distinct() %>%
-  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
-  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
-names(data)[3] <- "Value_maize1"
-names(data)[4] <- "Value_maize2"
-data <- subset(data, !is.na(data$Value_maize1) & !is.na(data$Value_maize2))
-
-nSingleExonGenes.total <- nrow(geneTranscript.counts[geneTranscript.counts$n == 1,])
-nSingleExonGenes.inSet <- unique(c(data$Maize1[data$Value_maize1 == 1], data$Maize2[data$Value_maize2 == 1]))
-
-data <- subset(data, data$Value_maize1 > 1 & data$Value_maize2 > 1) ## Ignore single-exon genes
-data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)
-data <- arrange(data, desc(foldChange))
-data[1:10,]
-
-
-
-
-
-#--------------------------------------------------------------------------------------------------#
-# Scatterplots
-#--------------------------------------------------------------------------------------------------#
-# Isoforms
-data <-
-  homeologs.pairs %>%
-  subset(Maize1 != "" & Maize2 != "") %>%
-  select(Maize1, Maize2) %>%
-  distinct() %>%
-  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
-  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
-names(data)[3] <- "Value_maize1"
-names(data)[4] <- "Value_maize2"
[…]

### 2017-10-13 SubGenome Analysis.Rmd
Examining correlations

@@
   labs(
-    title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1"),
+    title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1, reference line"),
     x = "Expression for Maize1 Gene (log scale)",
@@
   labs(
-    title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1"),
+    title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1, reference line"),
     x = "Abundance for Maize1 Gene (log scale)",
@@
 ```
+
+# Expression and Abundance are Weakly Correlated
+
+```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
+## Start with the paired expression data
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "FPKM_maize1"
+names(data)[5] <- "FPKM_maize2"
+
+## Add paired abundance data
+data <-
+  data %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "Sample"="Sample")) %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+names(data)[6] <- "dNSAF_maize1"
+names(data)[7] <- "dNSAF_maize2"
+
+## View as scatter
+ggplot(data, aes(FPKM_maize1, dNSAF_maize1)) +
+  geom_point(alpha=.1) +
+  geom_smooth(method="loess", se=TRUE, fullrange=FALSE, level=0.95) +
+  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  scale_x_log10() + scale_y_log10() +
+  facet_wrap(~Sample, nrow=6) +
+  labs(
+    title = paste0("Scatterplot of Expression in Maize1 vs. Abundance in Maize1\nLog-Log Scale for gene pair(s).\nAlpha=.1, reference line (black), loess fit (blue)"),
+    x = "Expression for Maize1 Gene (log scale)",
+    y = "Abundance for Maize1 Gene (log scale)"
+  ) 
+```
+
+```{r subgenome_expression_protein_foldchange_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=6, fig.height=5, fig.align='center'}
+## Start with the paired expression data
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "FPKM_maize1"
+names(data)[5] <- "FPKM_maize2"
+
+## Add paired abundance data
+data <-
+  data %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "Sample"="Sample")) %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+names(data)[6] <- "dNSAF_maize1"
+names(data)[7] <- "dNSAF_maize2"
+
+data$foldChange_expr <- log2(data$FPKM_maize1) - log2(data$FPKM_maize2)
+data$foldChange_abun <- log2(data$dNSAF_maize1) - log2(data$dNSAF_maize2)
+ggplot(data, aes(foldChange_expr, foldChange_abun)) +
+  geom_point(alpha=.1) +
+  xlim(-10,10) + ylim(-10,10) +
+  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
+  # facet_wrap(~Sample, nrow=6) +
+  labs(
+    title = paste0("Scatterplot of Foldchange in Expression in Maize1 vs.\nFoldchange in Abundance in Maize1\nAlpha=.1, linear fit, n=13413"),
+    x = "log2(Expression Maize1) - log2(Expression Maize2)",
+    y = "log2(Abundance Maize1) - log2(Abundance Maize2)"
+  ) + 
+  annotate("text",x=7,y=-8, label=paste("Pearson Cor = ", round(cor(data$foldChange_expr, data$foldChange_abun, method = c("pearson")),2)))
+```
+
 # Wrapup

### 2017-10-13 R/staging_testing.R

@@ -3,10 +3,15 @@
 
 # library(readr)
 # library(GenomicFeatures)
+library(fitdistrplus)
 library(tidyr)
 library(dplyr)
 library(ggplot2)
+select <- dplyr::select
 
+#--------------------------------------------------------------------------------------------------#
+# Local blast searchs
+#--------------------------------------------------------------------------------------------------#
 cmdBlast <- function(query, subject) {
   paste0("blastn -query <(", query, ") -subject <(", subject, ")")
 }
@@ -23,11 +28,75 @@ parseNeedleOutput <- function() {
 
 }
 
-
 system("perl ./Perl/runNeedle.pl acacgcacca acgcacgca")
-
 query <- "ATGTCGGGGCGCGGCAAGGGCGGGAAGGGCCTGGGCAAGGGCGGAGCGAAGCGCCACCGGAAGGTGCTCCGCGACAACATCCAGGGGATCACGAAGCCGGCGATCCGGCGGCTGGCGCGCAGGGGCGGCGTGAAGCGCATCTCCGGGCTCATCTACGAGGAGACCCGCGGCGTCCTCAAGATCTTCCTGGAGAACGTGATCCGCGACGCCGTCACCTACACCGAGCACGCGCGCCGCAAGACCGTCACAGCCATGGACGTCGTCTACGCGCTCAAGCGCCAGGGCCGCACCCTCTACGGCTTCGGCGGCTAGGCTATGCCGGTCGCGCTCGCGCCTCGCTGTCGTCGCTGCGGTTCTTGTTTCTGCAATTCGCAAGGTGTGGCTTGACGGGGAAAATGCTGGTTAATGTAGCACTAGATCAAGCATGTTGTGCGTGTCTGAAGAAATGCCAATTGTGACTTGCTGTTTGTTCAAACATGTTCCGTTTCTTCACATATCAAC"
[…]

### 2017-10-13 R/staging_testing.R
Examining correlations

@@ -3,9 +3,12 @@
 
 # library(readr)
 # library(GenomicFeatures)
+library(fitdistrplus)
 library(tidyr)
 library(dplyr)
 library(ggplot2)
+select <- dplyr::select
+
 
 cmdBlast <- function(query, subject) {
   paste0("blastn -query <(", query, ") -subject <(", subject, ")")
@@ -31,3 +34,41 @@ subject <- "ATGGAGATCCTCGAATCGACCCTGTTGGGCGAGTTCATCGGCTTCATCAAGGGGAACTGGTCAGCGCA
 cmdNeedle(query, subject)
 
 # ---
+
+
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "FPKM_maize1"
+names(data)[5] <- "FPKM_maize2"
+
+data <-
+  data %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "Sample"="Sample")) %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+names(data)[6] <- "dNSAF_maize1"
+names(data)[7] <- "dNSAF_maize2"
+
+## Fit test
+descdist(data$FPKM_maize1, discrete = FALSE)
+descdist(data$FPKM_maize2, discrete = FALSE)
+descdist(data$dNSAF_maize1, discrete = FALSE)
+descdist(data$dNSAF_maize2, discrete = FALSE)
+fit.gamma <- fitdist(data$Value_maize1, "gamma", method = "mme")
+plot(fit.gamma)
+fit.gamma$aic
+fit <- logspline(stats)
+
+## Correlation test
+cor(data$FPKM_maize1, data$FPKM_maize2, method = "pearson")
+cor(data$dNSAF_maize1, data$dNSAF_maize2, method = "pearson")
[…]

### 2017-10-20 SubGenome Analysis.Rmd
Interactive exploration with ggobi

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, eval.after = 'fig.cap')
 ```
@@
 
-# Import SynMap and GO Term Data
+# 1 Import SynMap and GO Term Data
 ***
@@
 
-# Choosing a KS cutoff value
+# 2 Choosing a KS cutoff value
 ***  
@@
 
-# Major Syntenic Blocks
+# 3 Major Syntenic Blocks
 ***
@@
 
-# SubGenomes
+# 4 SubGenomes
 ***
@@
 
-## Compared to Paper
+## 5 Compared to Paper
 How many chromosomes with synteny to sorghum were correctly placed in subgenome 1 or subgenome 2?
@@
 
-# GO Terms in Maize
+# 6 GO Terms in Maize
 ***
@@
 
-# GO Term Comparison by Subgenome
+# 7 GO Term Comparison by Subgenome
 ***
@@
 
-# Expression
+# 8 Expression
 
@@
 
-# Compare Homeolog Pairs
+# 9 Compare Homeolog Pairs
 
@@
 
-# Expression and Abundance are Weakly Correlated
+# 10 Expression and Abundance are Weakly Correlated
 
-```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
+```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
 ## Start with the paired expression data
@@
   )
+
+cap <- "Fig 10.1: "
 ```
 
-```{r subgenome_expression_protein_foldchange_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=6, fig.height=5, fig.align='center'}
+```{r subgenome_expression_protein_foldchange_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=6, fig.height=5, fig.align='center', fig.cap = cap}
 ## Start with the paired expression data
@@
   annotate("text",x=7,y=-8, label=paste("Pearson Cor = ", round(cor(data$foldChange_expr, data$foldChange_abun, method = c("pearson")),2)))
+
+cap <- "Fig 10.2: Considering only homeologous pairs where both genes were measured for expression and protein abundance, we look at fold change from maize2 expression to maize1 expression vs.
+the fold change from maize2 abundance vs.
+maize2 abundance.
+A slightly positive trend indicates a mild positive correlation between an increase in expression and an increase in abundance."
 ```
 
-# Wrapup
+# 11 Data Exploration
+
+```{r echo=FALSE, out.width='100%'}
+knitr::include_graphics('./Data/Images/High expressing genes tend to vary.png')
+```
+
+***
+
+```{r echo=FALSE, out.width='100%'}
+knitr::include_graphics('./Data/Images/Low Expressing Genes tend to stay low.png')
+```
+
+***
+
+```{r echo=FALSE, out.width='100%'}
+knitr::include_graphics('./Data/Images/Pollen is different.png')
+```
+
+***
+
+```{r echo=FALSE, out.width='100%'}
+knitr::include_graphics('./Data/Images/High-low variation.png')
+```
+
+Red Square = [Zm00001d020592](https://www.maizegdb.org/gene_center/gene/Zm00001d020592) Pericarp/Aleurone 27 DAP vs. [Zm00001d005793](https://www.maizegdb.org/gene_center/gene/Zm00001d005793)
+
+Yellow Square = [Zm00001d003188](https://www.maizegdb.org/gene_center/gene/Zm00001d003188) Pericarp/Aleurone 27 DAP vs. [Zm00001d025753](https://www.maizegdb.org/gene_center/gene/Zm00001d025753)
+
+Yellow X = Zm00001d039040 Mature Leaf 8
+
+Pink Cross = Zm00001d015385 Mature Leaf 8
+
+# 12 Wrapup
 ***

### 2017-10-20 R/do.R

@@ -177,6 +177,54 @@ expressedPairs <-
   rename(FPKM_mean2=FPKM_mean)
 
 #==================================================================================================#
+## Correlation between exp and abundance?
+#--------------------------------------------------------------------------------------------------#
+## Start with the paired expression data
+# data <-
+#   homeologs.pairs %>%
+#   subset(Maize1 != "" & Maize2 != "") %>%
+#   select(Maize1, Maize2) %>%
+#   distinct() %>%
+#   inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
+#   inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+#   subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+# names(data)[4] <- "FPKM_maize1"
+# names(data)[5] <- "FPKM_maize2"
+#
+# ## Add paired abundance data
+# data <-
+#   data %>%
+#   inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "Sample"="Sample")) %>%
+#   inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+#   subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+# names(data)[6] <- "dNSAF_maize1"
+# names(data)[7] <- "dNSAF_maize2"
+#
+# ## Fit test -> This data is not normal!  Log-normal seems like a best fit.
+# descdist(data$FPKM_maize1, discrete = FALSE)
+# descdist(data$FPKM_maize2, discrete = FALSE)
+# descdist(data$dNSAF_maize1, discrete = FALSE)
+# descdist(data$dNSAF_maize2, discrete = FALSE)
+# fit.lnorm <- fitdist(data$FPKM_maize1, "lnorm", method = "mme")
+# plot(fit.lnorm)
+# fit.lnorm$aic
+# fit <- logspline(stats)
+#
+# ## Correlation test
+# cor(data$FPKM_maize1, data$FPKM_maize2, method = "pearson")
+# cor(data$dNSAF_maize1, data$dNSAF_maize2, method = "pearson")
+# cor(data$FPKM_maize1, data$dNSAF_maize1, method = "pearson")
+#
+# cor(data$FPKM_maize1, data$FPKM_maize2, method = "spearman")
+# cor(data$dNSAF_maize1, data$dNSAF_maize2, method = "spearman")
[…]

### 2017-10-20 R/staging_testing.R

@@ -57,39 +57,13 @@ data <-
 names(data)[6] <- "dNSAF_maize1"
 names(data)[7] <- "dNSAF_maize2"
 
-## Fit test -> This data is not normal!  Log-normal seems like a best fit.
-descdist(data$FPKM_maize1, discrete = FALSE)
-descdist(data$FPKM_maize2, discrete = FALSE)
-descdist(data$dNSAF_maize1, discrete = FALSE)
-descdist(data$dNSAF_maize2, discrete = FALSE)
-fit.lnorm <- fitdist(data$FPKM_maize1, "lnorm", method = "mme")
-plot(fit.lnorm)
-fit.lnorm$aic
-fit <- logspline(stats)
-
-## Correlation test
-cor(data$FPKM_maize1, data$FPKM_maize2, method = "pearson")
-cor(data$dNSAF_maize1, data$dNSAF_maize2, method = "pearson")
-cor(data$FPKM_maize1, data$dNSAF_maize1, method = "pearson")
-
-cor(data$FPKM_maize1, data$FPKM_maize2, method = "spearman")
-cor(data$dNSAF_maize1, data$dNSAF_maize2, method = "spearman")
-cor(data$FPKM_maize1, data$dNSAF_maize1, method = "spearman")
-
-fit <- lm(data=data, FPKM_maize1 ~ dNSAF_maize1)
-summary(fit)
-par(mfrow = c(2, 2))
-plot(fit)
-
-
 data$foldChange_expr <- log2(data$FPKM_maize1) - log2(data$FPKM_maize2)
 data$foldChange_abun <- log2(data$dNSAF_maize1) - log2(data$dNSAF_maize2)
+
 ggplot(data, aes(foldChange_expr, foldChange_abun)) +
   geom_point(alpha=.1) +
   xlim(-10,10) + ylim(-10,10) +
   geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
-  # geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
-  # scale_x_log10() + scale_y_log10() +
   # facet_wrap(~Sample, nrow=6) +
   labs(
     title = paste0("Scatterplot of Expression in Maize1 vs. Abundance in Maize1\nLog-Log Scale for gene pair(s).\nAlpha=.1, linear abline, loess smoothing"),

### 2017-10-20 SubGenome Analysis.Rmd
Added fit line

@@
   scale_color_gradient(low = "blue", high = "red") +
+  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
   geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
@@
   labs(
-    title = paste0("Scatterplot of isoform counts for gene pair(s).\nExcludes ",n," outlier, alpha=.1, foldchange in log2."),
+    title = paste0("Scatterplot of isoform counts for gene pair(s).\nExcludes ",n," outlier, alpha=.1, foldchange in log2, blue=lm fit."),
     x = "Isoform Counts for Maize1 Gene",
     y = "Isoform Counts for Maize2 Gene"
-  )
+  ) + 
+  annotate("text",x=85,y=20, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
 plot(plot)
@@
 
+```{r include=FALSE}
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "FPKM_maize1"
+names(data)[5] <- "FPKM_maize2"
+
+## Add paired abundance data
+data <-
+  data %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "Sample"="Sample")) %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+names(data)[6] <- "dNSAF_maize1"
+names(data)[7] <- "dNSAF_maize2"
+
+data$foldChange_expr <- log2(data$FPKM_maize1) - log2(data$FPKM_maize2)
+data$foldChange_abun <- log2(data$dNSAF_maize1) - log2(data$dNSAF_maize2)
+```
+
 ```{r echo=FALSE, out.width='100%'}
@@
 
+***
+
+```{r echo=FALSE, out.width='100%'}
+knitr::include_graphics('./Data/Images/Top fold-change genes.png')
+```
+
 # 12 Wrapup

### 2017-10-26 SubGenome Analysis.Rmd
Minor visual updates

@@
   data %>%
-  subset(Value_maize1 > 100 | Value_maize2 > 100) %>%
+  # subset(Value_maize1 > 100 | Value_maize2 > 100) %>%
+  subset(Value_maize1 > 25 | Value_maize2 > 25) %>%
   nrow()
@@
 plot <-
-  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
-  geom_point(alpha=.1, aes(color = abs(foldChange))) +
+  ggplot(subset(data, Value_maize1 <= 25 & Value_maize2 <= 25), aes(x=Value_maize1, y=Value_maize2)) +
+  geom_point(position = "jitter", alpha=.1, aes(color = abs(foldChange))) +
   scale_color_gradient(low = "blue", high = "red") +
@@
   geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
-  xlim(0,100) +
-  ylim(0,100) +
+  # xlim(0,100) +
+  # ylim(0,100) +
+  xlim(0,25) +
+  ylim(0,25) +
   labs(
@@
   ) + 
-  annotate("text",x=85,y=20, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
+  # annotate("text",x=85,y=20, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
+  annotate("text",x=20,y=3, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
 plot(plot)
@@
 
-```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
-## Start with the paired expression data
-data <-
-  homeologs.pairs %>%
-  subset(Maize1 != "" & Maize2 != "") %>%
-  select(Maize1, Maize2) %>%
-  distinct() %>%
-  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
-  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
-  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
-names(data)[4] <- "FPKM_maize1"
-names(data)[5] <- "FPKM_maize2"
-
-## Add paired abundance data
-data <-
-  data %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "Sample"="Sample")) %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
-  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
-names(data)[6] <- "dNSAF_maize1"
-names(data)[7] <- "dNSAF_maize2"
-
-## View as scatter
-ggplot(data, aes(FPKM_maize1, dNSAF_maize1)) +
-  geom_point(alpha=.1) +
-  geom_smooth(method="loess", se=TRUE, fullrange=FALSE, level=0.95) +
-  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
-  scale_x_log10() + scale_y_log10() +
-  facet_wrap(~Sample, nrow=6) +
-  labs(
-    title = paste0("Scatterplot of Expression in Maize1 vs. Abundance in Maize1\nLog-Log Scale for gene pair(s).\nAlpha=.1, reference line (black), loess fit (blue)"),
-    x = "Expression for Maize1 Gene (log scale)",
-    y = "Abundance for Maize1 Gene (log scale)"
-  )
-
-cap <- "Fig 10.1: "
-```
-
 ```{r subgenome_expression_protein_foldchange_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=6, fig.height=5, fig.align='center', fig.cap = cap}
@@
 
-cap <- "Fig 10.2: Considering only homeologous pairs where both genes were measured for expression and protein abundance, we look at fold change from maize2 expression to maize1 expression vs.
+cap <- "Fig 10.1: Considering only homeologous pairs where both genes were measured for expression and protein abundance, we look at fold change from maize2 expression to maize1 expression vs.
 the fold change from maize2 abundance vs.
@@
 A slightly positive trend indicates a mild positive correlation between an increase in expression and an increase in abundance."
+```
+
+```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
+## Start with the paired expression data
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "FPKM_maize1"
+names(data)[5] <- "FPKM_maize2"
+
+## Add paired abundance data
+data <-
+  data %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "Sample"="Sample")) %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+names(data)[6] <- "dNSAF_maize1"
+names(data)[7] <- "dNSAF_maize2"
+
+## View as scatter
+ggplot(data, aes(FPKM_maize1, dNSAF_maize1)) +
+  geom_point(alpha=.1) +
+  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
+  # geom_smooth(method="loess", se=TRUE, fullrange=FALSE, level=0.95) +
+  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  scale_x_log10() + scale_y_log10() +
+  facet_wrap(~Sample, nrow=6) +
+  labs(
+    title = paste0("Scatterplot of Expression in Maize1 vs. Abundance in Maize1\nLog-Log Scale for gene pair(s).\nAlpha=.1, reference line (black), lm fit (blue)"),
+    x = "Expression for Maize1 Gene (log scale)",
+    y = "Abundance for Maize1 Gene (log scale)"
+  )
+
+cap <- "Fig 10.2: "
 ```

### 2017-11-30 SubGenome Analysis.Rmd
Added some captions

@@
 
-```{r pickKSCutoff, echo=FALSE, fig.width=3, fig.height=3}
+```{r pickKSCutoff, echo=FALSE, fig.width=3, fig.height=3, fig.cap = cap}
 qplot(
@@
 )
+cap <- "Fig 2.1: "
 ```
@@
 
-```{r reportSyntenicRegions, echo=FALSE, results="asis"}
+```{r reportSyntenicRegions, echo=FALSE, results="asis", fig.cap = cap}
 ## Display chromosomes with homeologous genes between org1 and org2
@@
                align = "lccccccccccr")
+cap <- "Table 3.1: "
 ```
@@
 
-```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center'}
+```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
 plot <- graphGenePairDominanceByExperiment(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
 plot(plot)
+cap <- "Fig 9.1:"
 ```
@@
 
-```{r subgenome_protein_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center'}
+```{r subgenome_protein_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
 factor <- 4
@@
 plot(plot)
+cap <- "Fig 9.2:"
 ```
@@
 
-```{r subgenome_expression_dominance_across_factors, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=12, fig.align='center'}
+```{r subgenome_expression_dominance_across_factors, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=12, fig.align='center', fig.cap = cap}
 plot <- graphGenePairDominanceByExperimentAcrossFactors(maize.expression.sample.avg, homeologs.pairs, FALSE)
 plot(plot)
+cap <- "Fig 9.3:"
 ```
@@
 
-```{r subgenome_isoform_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=6, fig.height=4, fig.align='center'}
+```{r subgenome_isoform_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=6, fig.height=4, fig.align='center', fig.cap = cap}
 ## Plot frequency of isoform counts for each gene
@@
 plot(plot)
+cap <- "Fig 9.4: "
 ```
@@
 
-```{r subgenome_isoforms_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.align='center'}
+```{r subgenome_isoforms_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.align='center', fig.cap = cap}
 # Isoforms
@@
 plot(plot)
+
+cap <- "Fig 9.5: "
 ```
 
-```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
+```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
 data <-
@@
 plot(plot)
+
+cap <- "Fig 9.5: "
 ```
 
-```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
+```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
 data <-
@@
 plot(plot)
+
+cap <- "Fig 9.6: "
 ```

### 2017-12-01 SubGenome Analysis.Rmd
Kaeppler expression data
Import and use kaeppler (must replace Walley data at bottom of do.R)

Updated reports

Made the log10_ks_cutoff a variable and not a knitr param.

Fixed a few minor issues in knitr reports.

@@
   subgenome assignments in maize'
-params:
-  log10_ks_cutoff: 0
 output:
@@
 
-You have selected the log10(ks) cutoff value to be **`r params$log10_ks_cutoff`**.  Again, this should ideally separate two peaks in the middle graph.
+You have selected the log10(ks) cutoff value to be **`r log10_ks_cutoff`**.  Again, this should ideally separate two peaks in the middle graph.
 
-```{r pickKSCutoff, echo=FALSE, fig.width=3, fig.height=3, fig.cap = cap}
+```{r pickKSCutoff, echo=FALSE, fig.width=3, fig.height=3}
 qplot(
@@
   data = NULL, 
-  xintercept = 10^params$log10_ks_cutoff, 
+  xintercept = 10^log10_ks_cutoff, 
   na.rm = FALSE, 
@@
   data = NULL, 
-  xintercept = params$log10_ks_cutoff, 
+  xintercept = log10_ks_cutoff, 
   na.rm = FALSE, 
@@
 grid.newpage()
+a1 = expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) > 2) %>% nrow()
+a2 = expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) > 2) %>% nrow()
+a3 = nrow(expressedPairs) - a1 - a2
 g <- draw.pairwise.venn(
-  area1 = expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2) %>% nrow(),
-  area2 = expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2) %>% nrow(),
-  cross.area = intersect(expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2),
-                         expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2)) %>%
-                          nrow(), 
+  area1 = a1+a3,
+  area2 = a2+a3,
+  cross.area = a3, 
   category = c("Sub1", "Sub2"), 
@@
 
-Paired t-test suggests that there is no difference in means between FPKM of homeologs in sub1 vs. sub2 (p-val > 0.05, so we accept null hypothesis of equal means):
+Paired (two-sided) t-test suggests that there is no difference in means between FPKM of homeologs in sub1 vs. sub2 (p-val > 0.05, so we fail to reject null hypothesis of equal means):
 ```{r subgenome_fpkm_ttest, echo=FALSE}

### 2017-12-01 R/cleanData.R

@@ -13,6 +13,7 @@ startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
 ##        syntelogs.sorghum.v1.maize.v1.raw
 ##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw
 ##        maize.expression.raw
+##        maize.kaeppler.expression.raw
 ##        go.maize.raw
 ##        go.sorghum.raw
 ##        maize.genes.v3_to_v4_map.raw
@@ -22,6 +23,7 @@ startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
 ##        syntelogs.sorghum.v1.maize.v1.clean
 ##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean
 ##        maize.expression.clean
+##        maize.kaeppler.expression.clean
 ##        go.maize.clean
 ##        go.sorghum.clean
 ##        maize.genes.v3_to_v4_map.clean
@@ -73,6 +75,19 @@ maize.protein.abundance.clean <-
 
 ## Remove low dNSAF values
 maize.protein.abundance.clean[maize.protein.abundance.clean < 1] <- NA
+
+#==================================================================================================#
+## maize.kaeppler.expression.raw
+#--------------------------------------------------------------------------------------------------#
+maize.kaeppler.expression.clean <- maize.kaeppler.expression.raw
+
+## Select relevant columns (i.e. get rid of chromosome position columns)
+maize.kaeppler.expression.clean <-
+  maize.kaeppler.expression.clean %>%
+  select(-chromosome, -position_left, -position_right)
+
+## Remove low FPKM values
+maize.kaeppler.expression.clean[maize.kaeppler.expression.clean < 1] <- NA
 
 #==================================================================================================#
 ## go.maize.raw

### 2017-12-01 R/do.R

@@ -142,6 +142,49 @@ maize.protein.abundance.sample.avg <-
 ## When all replicates have NA, mean returns NaN.  Convert it back to NA.
 maize.protein.abundance.sample.avg$dNSAF_avg[is.nan(maize.protein.abundance.sample.avg$dNSAF_avg)] <- NA
 
+
+#==================================================================================================#
+## maize.kaeppler.expression.all
+#--------------------------------------------------------------------------------------------------#
+maize.kaeppler.expression.all <- maize.kaeppler.expression.clean
+
+## Convert to v4 ids
+maize.kaeppler.expression.all <-
+  maize.kaeppler.expression.all %>%
+  inner_join(maize.genes.v3_to_v4_map.clean, by=c("Maize_AGPv2_gene" = "v3_id")) %>%
+  rename(geneID=v4_id)
+
+## Reorder columns
+maize.kaeppler.expression.all <- maize.kaeppler.expression.all[,c(81,2:80)]
+
+## Merge duplicate rows by adding FPKM values together
+maize.kaeppler.expression.all <-
+  maize.kaeppler.expression.all %>%
+  group_by(geneID) %>%
+  summarise_all(funs(sum))
+
+#==================================================================================================#
+## maize.kaeppler.expression.sample.avg
+#--------------------------------------------------------------------------------------------------#
+maize.kaeppler.expression.sample.avg <- maize.kaeppler.expression.clean
+
+## Convert to v4 ids
+maize.kaeppler.expression.sample.avg <-
+  maize.kaeppler.expression.sample.avg %>%
+  inner_join(maize.genes.v3_to_v4_map.clean, by=c("Maize_AGPv2_gene" = "v3_id")) %>%
+  rename(geneID=v4_id)
+
+## Reorder columns
+maize.kaeppler.expression.sample.avg <- maize.kaeppler.expression.sample.avg[,c(81,2:80)]
+
+## Output in long form
+maize.kaeppler.expression.sample.avg <-
+  maize.kaeppler.expression.sample.avg %>%
+  gather("Sample", "FPKM",-1) %>%
+  group_by(geneID, Sample) %>%
+  summarise(FPKM_avg=mean(FPKM, na.rm=TRUE)) %>%
+  arrange(geneID)
+
 #==================================================================================================#
 ## parseExpressionData.R
[…]

### 2017-12-01 R/loadData.R

@@ -11,7 +11,8 @@ library(GenomicFeatures)
 ##        maize.expression.raw
 ##        experiment.map
 ##        maize.expression.raw
-##        experiment.map.proteins
+##        maize.protein.abundance.raw
+##        maize.kaeppler.expression.raw
 ##        go.maize.raw
 ##        go.sorghum.raw
 ##        go.goSlim.plant
@@ -45,6 +46,9 @@ experiment.map <- read_delim("./Data/Expression/tracking_ids.csv", ",", trim_ws 
 maize.protein.abundance.raw <- read_xlsx("~/git/SubGenomes/Data/ProteinAbundance/aag1125_SupportingFile_Table_S2-1.xlsx", sheet = 3, col_names = TRUE)
 experiment.map.proteins <- read_delim("./Data/ProteinAbundance/experiment_map_proteins.csv", ",", trim_ws = TRUE)
 
+## Expression data from the Kaeppler 2015 paper in FPKM for 79 tissues
+maize.kaeppler.expression.raw <- read_delim("./Data/Expression/Dataset S1.txt", "\t", trim_ws = TRUE)
+
 ## Read in GO Annotation data for maize genes
 go.maize.raw <- read_delim("./Data/GO/go_from_ maizecyc.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
 
@@ -72,6 +76,8 @@ maize.genes.v3_to_v4_map.raw <- read_xlsx("./Data/MaizeGDB_v3_v4.genes.xlsx")
 ## Load maize GFF data
 txdb <- makeTxDbFromGFF("./Data/MaizeGFF3/Zea_mays.AGPv4.32.gff3.gz", format="gff3")
 
+## log10_ks_cutoff, hand picked based on SynMap reported ks/kn data
+log10_ks_cutoff <- 0
 #--------------------------------------------------------------------------------------------------#
 detach("package:readr", unload=TRUE)
 detach("package:readxl", unload=TRUE)

### 2017-12-01 R/parseSubgenomes.R

@@ -18,7 +18,6 @@ library(dplyr)
 ## Date: 2017-07-26
 ## Author: Jesse R. Walsh
 ####################################################################################################
-log10_ks_cutoff <- params$log10_ks_cutoff
 ks_cutoff <- 10^log10_ks_cutoff
 
 ## Add median and geneCount values (aggregated by block/org_chr1/org_chr2) to each row.  Calculate gene sizes.  Add chromosome ids as numbers.

### 2017-12-14 SubGenome Analysis.Rmd
Migrating to data-as-packages format
Also enabling easier transition between windows/linux platforms by updating file paths.

@@
 ```{r importSynMapData, include=FALSE}
-load("~/git/SubGenomes/Data/SavedObjects/loadedData.RData")
-
-# source("~/git/SubGenomes/R/loadData.R")
-# source("~/git/SubGenomes/R/cleanData.R")
-source("~/git/SubGenomes/R/functions.R")
-source("~/git/SubGenomes/R/do.R")
+source("./R/do.R")
 
@@
 ```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./Data/Images/High expressing genes tend to vary.png')
+knitr::include_graphics('./data/Images/High expressing genes tend to vary.png')
 ```
@@
 ```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./Data/Images/Low Expressing Genes tend to stay low.png')
+knitr::include_graphics('./data/Images/Low Expressing Genes tend to stay low.png')
 ```
@@
 ```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./Data/Images/Pollen is different.png')
+knitr::include_graphics('./data/Images/Pollen is different.png')
 ```
@@
 ```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./Data/Images/High-low variation.png')
+knitr::include_graphics('./data/Images/High-low variation.png')
 ```
@@
 ```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./Data/Images/Top fold-change genes.png')
+knitr::include_graphics('./data/Images/Top fold-change genes.png')
 ```

### 2017-12-14 R/cleanData.R

@@ -2,7 +2,7 @@ library(GenomicFeatures)
 library(topGO)
 library(tidyr)
 library(dplyr)
-startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
+# startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
 ####################################################################################################
 ## Project:         Subgenomes project
 ## Script purpose:  Clean up raw datafiles needed for this project, including renaming column and
@@ -58,79 +58,79 @@ syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$gene1 <- gsub(syntelogs.sorgh
 #==================================================================================================#
 ## maize.expression.raw
 #--------------------------------------------------------------------------------------------------#
-maize.expression.clean <- maize.expression.raw
-
-## Remove low FPKM values
-maize.expression.clean[maize.expression.clean < 1] <- NA
+# maize.expression.clean <- maize.expression.raw
+#
+# # ## Remove low FPKM values
+# maize.expression.clean[maize.expression.clean < 1] <- NA
 
 #==================================================================================================#
 ## maize.protein.abundance.raw
 #--------------------------------------------------------------------------------------------------#
-maize.protein.abundance.clean <- maize.protein.abundance.raw
+# maize.protein.abundance.clean <- maize.protein.abundance.raw
 
-## Rename the gene id column
-maize.protein.abundance.clean <-
-  maize.protein.abundance.clean %>%
-  rename(v3_id=X__1)
-
-## Remove low dNSAF values
-maize.protein.abundance.clean[maize.protein.abundance.clean < 1] <- NA
+# ## Rename the gene id column
+# maize.protein.abundance.clean <-
+#   maize.protein.abundance.clean %>%
+#   rename(v3_id=X__1)
+#
+# ## Remove low dNSAF values
+# maize.protein.abundance.clean[maize.protein.abundance.clean < 1] <- NA
 
[…]

### 2017-12-14 R/do.R

@@ -1,5 +1,3 @@
-library(tidyr)
-library(dplyr)
 ####################################################################################################
 ## Project:         Subgenomes project
 ## Script purpose:  Create and combine objects to use for analysis and plots.
@@ -9,10 +7,17 @@ library(dplyr)
 ## Date: 2017-09-14
 ## Author: Jesse R. Walsh
 ####################################################################################################
+# load("~/git/SubGenomes/Data/SavedObjects/loadedData.RData")
+load("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\SavedObjects\\loadedData.RData")
+# source("./R/loadData.R")
+# source("./R/cleanData.R")
+source("./R/functions.R")
+library(tidyr)
+library(dplyr)
 #==================================================================================================#
 ## parseSubgenomes.R
 #--------------------------------------------------------------------------------------------------#
-source("~/git/SubGenomes/R/parseSubgenomes.R")
+source("./R/parseSubgenomes.R")
 
 #==================================================================================================#
 ## addGOTerms.R

### 2017-12-14 R/loadData.R

@@ -1,6 +1,3 @@
-library(readr)
-library(readxl)
-library(GenomicFeatures)
 ####################################################################################################
 ## Project:         Subgenomes project
 ## Script purpose:  Import raw dataset files relevant to this project
@@ -27,36 +24,33 @@ library(GenomicFeatures)
 # install_github("jrwalsh/MaizeGO")
 # install_github("jrwalsh/MaizeOmics")
 # install_github("jrwalsh/MaizeMap")
-# library(MaizeGO)
-# library(MaizeOmics)
-# library(MaizeMap)
-
-# syntelogs.sorghum.v1.maize.v1.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\SynMap\\sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-# syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\SynMap\\sorghum_v3.1_vs_maize_v4+rejected.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-# maize.expression.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\Expression\\GSE50191_FPKM.tsv", "\t", trim_ws = TRUE)
-# go.maize.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO\\go_from_ maizecyc.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-# go.sorghum.raw <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO\\gramene_sorghumv2_goterms.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
-# go.goSlim.plant <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\GO\\goslim_plant.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-# subgenome.assignments <- read_delim("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\outfile_processed_byHand.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
[…]

### 2017-12-21 SubGenome Analysis.Rmd
Major structure overhall
Moved data from "data" to "data-raw", reformatted data loading to take advantage of packaged data, moved much of the "cleaning" functionality and id mapping in do.R to clean.R.

@@
 
-library(topGO)
-library(readr)
-library(readxl)
 library(ggplot2)
@@
 gene.foldchange[1:10,] %>% 
-  select(Maize1_url, Maize2_url, Sample, Value_maize1, Value_maize2, foldChange) %>%
+  select(Maize1_url, Maize2_url, sample, Value_maize1, Value_maize2, foldChange) %>%
   knitr::kable(caption = "Top ten differentially expressed gene pairs (in log2)", align = "llrrr")
@@
 protein.foldchange[1:10,] %>% 
-  select(Maize1_url, Maize2_url, Sample, Value_maize1, Value_maize2, foldChange) %>%
+  select(Maize1_url, Maize2_url, sample, Value_maize1, Value_maize2, foldChange) %>%
   knitr::kable(caption = "Top ten differentially abundant protein pairs (in log2)", align = "llrrr")
@@
 gene.foldchange[1:50,] %>%
-  inner_join(protein.foldchange[1:50,], by = c("Maize1", "Maize2", "Sample")) %>%
+  inner_join(protein.foldchange[1:50,], by = c("Maize1", "Maize2", "sample")) %>%
   rename(foldChange_expr=foldChange.x, foldChange_prot=foldChange.y, Maize1_url=Maize1_url.x, Maize2_url=Maize2_url.x) %>%
-  select(Maize1_url, Maize2_url, Sample, foldChange_expr, foldChange_prot) %>%
+  select(Maize1_url, Maize2_url, sample, foldChange_expr, foldChange_prot) %>%
   knitr::kable(caption = "Gene pairs in the top 50 for both differential gene expression and differential protein abundance (in log2).", align = "llrrrr")
@@
   inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
-  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
   subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
@@
 names(data)[5] <- "Value_maize2"
-# data <- subset(data, Sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))
+# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))
 
@@
   ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
-  geom_point(alpha=.1, aes(color = Sample)) +
+  geom_point(alpha=.1, aes(color = sample)) +
   geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
@@
   ) +
-  theme(legend.position = "none") + facet_wrap(~Sample, nrow=6)
+  theme(legend.position = "none") + facet_wrap(~sample, nrow=6)
 plot(plot)
@@
   inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID")) %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
   subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
@@
   data %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "Sample"="Sample")) %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "Sample"="Sample")) %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "sample"="sample")) %>%
+  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
   subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
@@
   geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
-  # facet_wrap(~Sample, nrow=6) +
+  # facet_wrap(~sample, nrow=6) +
   labs(
@@
   scale_x_log10() + scale_y_log10() +
-  facet_wrap(~Sample, nrow=6) +
+  facet_wrap(~sample, nrow=6) +
   labs(
@@
 ```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./data/Images/High expressing genes tend to vary.png')
+knitr::include_graphics('./data-raw/Images/High expressing genes tend to vary.png')
 ```
@@
 ```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./data/Images/Low Expressing Genes tend to stay low.png')
+knitr::include_graphics('./data-raw/Images/Low Expressing Genes tend to stay low.png')
 ```
@@
 ```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./data/Images/Pollen is different.png')
+knitr::include_graphics('./data-raw/Images/Pollen is different.png')
 ```
@@
 ```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./data/Images/High-low variation.png')
+knitr::include_graphics('./data-raw/Images/High-low variation.png')
 ```
@@
 ```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./data/Images/Top fold-change genes.png')
+knitr::include_graphics('./data-raw/Images/Top fold-change genes.png')
 ```

### 2017-12-21 R/cleanData.R

@@ -1,8 +1,3 @@
-library(GenomicFeatures)
-library(topGO)
-library(tidyr)
-library(dplyr)
-# startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
 ####################################################################################################
 ## Project:         Subgenomes project
 ## Script purpose:  Clean up raw datafiles needed for this project, including renaming column and
@@ -10,39 +5,158 @@ library(dplyr)
 ##        conversions or modification to data to make it usable.
 ##
 ## Input:
+##        maize.genes.v3_to_v4_map.raw
+##        maize.expression.raw
+##        maize.expression.sample.avg.raw
+##        maize.protein.abundance.sample.avg.raw
+##        maize.kaeppler.expression.raw
+##        maize.kaeppler.expression.sample.avg.raw
+##        go.maize.raw
 ##        syntelogs.sorghum.v1.maize.v1.raw
 ##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw
-##        maize.expression.raw
-##        maize.kaeppler.expression.raw
-##        go.maize.raw
 ##        go.sorghum.raw
-##        maize.genes.v3_to_v4_map.raw
 ##        txdb
-##        maize.protein.abundance.raw
+##
 ## Output:
+##        maize.genes.v3_to_v4_map.clean
+##        maize.expression.clean
+##        maize.expression.sample.avg.clean
+##        maize.protein.abundance.sample.avg.clean
+##        maize.kaeppler.expression.clean
+##        maize.kaeppler.expression.sample.avg.clean
+##        go.maize.clean
 ##        syntelogs.sorghum.v1.maize.v1.clean
 ##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean
-##        maize.expression.clean
-##        maize.kaeppler.expression.clean
-##        go.maize.clean
 ##        go.sorghum.clean
-##        maize.genes.v3_to_v4_map.clean
 ##        geneTranscript.map
-##        maize.protein.abundance.clean
+##        geneTranscript.counts
 ##
 ## Date: 2017-08-25
 ## Author: Jesse R. Walsh
 ####################################################################################################
+library(GenomicFeatures)
+library(topGO)
+library(tidyr)
+library(dplyr)
+# startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
+
+#==================================================================================================#
+## maize.genes.v3_to_v4_map.raw
[…]

### 2017-12-21 R/do.R

@@ -7,13 +7,27 @@
 ## Date: 2017-09-14
 ## Author: Jesse R. Walsh
 ####################################################################################################
-# load("~/git/SubGenomes/Data/SavedObjects/loadedData.RData")
-load("C:\\Users\\Jesse\\Dropbox (Personal)\\Link to Subgenome Data\\SavedObjects\\loadedData.RData")
-# source("./R/loadData.R")
-# source("./R/cleanData.R")
+# library(devtools)
+# install_github("jrwalsh/MaizeGO", force = TRUE)
+# install_github("jrwalsh/MaizeOmics", force = TRUE)
+# install_github("jrwalsh/MaizeMap", force = TRUE)
+# install.packages("readr")
+# install.packages("readxl")
+# install.packages("tidyr")
+# install.packages("dplyr")
+# install.packages("VennDiagram")
+# install.packages("gridExtra")
+## BioConductor Packages
+# source("https://bioconductor.org/biocLite.R")
+# biocLite("GenomicFeatures")
+# biocLite("topGO")
+
 source("./R/functions.R")
+# saveMyObjects() # only need to run this is there is a change in raw data or the cleaning procedures
+loadMyObjects()
 library(tidyr)
 library(dplyr)
+
 #==================================================================================================#
 ## parseSubgenomes.R
 #--------------------------------------------------------------------------------------------------#
@@ -73,122 +87,25 @@ maizeWithSorghumGO <-
 #--------------------------------------------------------------------------------------------------#
 maize.expression.all <- maize.expression.clean
 
-## Convert to v4 ids
-maize.expression.all <-
-  maize.expression.all %>%
-  inner_join(maize.genes.v3_to_v4_map.clean, by=c("tracking_id" = "v3_id")) %>%
-  rename(geneID=v4_id)
-
-## Reorder columns
-maize.expression.all <- maize.expression.all[,c(70,2:69)]
-
-## Merge duplicate rows by adding FPKM values together
-maize.expression.all <-
-  maize.expression.all %>%
-  group_by(geneID) %>%
-  summarise_all(funs(sum))
-
 #==================================================================================================#
 ## maize.expression.sample.avg
[…]

### 2017-12-21 R/functions.R

@@ -25,7 +25,7 @@ dataGenePairDominanceByExperiment <- function(maize.data, homeologs.pairs, facto
   data <-
     genePairs %>%
     inner_join(maize.data, by=c("Maize1"="geneID")) %>%
-    inner_join(maize.data, by=c("Maize2"="geneID", "Sample"="Sample"))
+    inner_join(maize.data, by=c("Maize2"="geneID", "sample"="sample"))
   names(data)[4] <- "Value_maize1"
   names(data)[5] <- "Value_maize2"
 
@@ -41,8 +41,8 @@ dataGenePairDominanceByExperiment <- function(maize.data, homeologs.pairs, facto
 
   data <-
     data %>%
-    select(Sample, dominance) %>%
-    group_by(Sample, dominance) %>%
+    select(sample, dominance) %>%
+    group_by(sample, dominance) %>%
     count()
 
   return(data)
@@ -58,7 +58,7 @@ graphGenePairDominanceByExperiment <- function(maize.data, homeologs.pairs, fact
   data <-
     genePairs %>%
     inner_join(maize.data, by=c("Maize1"="geneID")) %>%
-    inner_join(maize.data, by=c("Maize2"="geneID", "Sample"="Sample"))
+    inner_join(maize.data, by=c("Maize2"="geneID", "sample"="sample"))
   names(data)[4] <- "Value_maize1"
   names(data)[5] <- "Value_maize2"
 
@@ -71,7 +71,7 @@ graphGenePairDominanceByExperiment <- function(maize.data, homeologs.pairs, fact
 
   data <- dataGenePairDominanceByExperiment(maize.data, homeologs.pairs, factor, includeNA)
   plot <-
-    ggplot(data, aes(Sample, n/nPairs*100)) +
+    ggplot(data, aes(sample, n/nPairs*100)) +
     geom_bar(aes(fill=dominance), position="dodge", stat="identity") +
     labs(y = paste0("Percent of gene pairs (n=",nPairs,")"),
          x = "Experiment",
@@ -94,7 +94,7 @@ graphGenePairDominanceByExperimentAcrossFactors <- function(maize.data, homeolog
   data <-
     genePairs %>%
     inner_join(maize.data, by=c("Maize1"="geneID")) %>%
-    inner_join(maize.data, by=c("Maize2"="geneID", "Sample"="Sample"))
+    inner_join(maize.data, by=c("Maize2"="geneID", "sample"="sample"))
   names(data)[4] <- "Value_maize1"
[…]

### 2017-12-21 R/loadData.R

@@ -3,27 +3,27 @@
 ## Script purpose:  Import raw dataset files relevant to this project
 ##
 ## Output:
+##        log10_ks_cutoff
+##        subgenome.truth
+##
+##        maize.genes.v3_to_v4_map.raw
+##        maize.expression.raw
+##        maize.expression.sample.avg.raw
+##        maize.protein.abundance.sample.avg.raw
+##        maize.kaeppler.expression.raw
+##        maize.kaeppler.expression.sample.avg.raw
+##        go.maize.raw
+##
 ##        syntelogs.sorghum.v1.maize.v1.raw
 ##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw
-##        maize.expression.raw
-##        experiment.map
-##        maize.expression.raw
-##        maize.protein.abundance.raw
-##        maize.kaeppler.expression.raw
-##        go.maize.raw
 ##        go.sorghum.raw
 ##        go.goSlim.plant
 ##        subgenome.assignments
-##        subgenome.truth
-##        maize.genes.v3_to_v4_map.raw
 ##        txdb
 ##
 ## Date: 2017-08-25
 ## Author: Jesse R. Walsh
 ####################################################################################################
-# install_github("jrwalsh/MaizeGO")
-# install_github("jrwalsh/MaizeOmics")
-# install_github("jrwalsh/MaizeMap")
 library(readr)
 library(readxl)
 library(GenomicFeatures)
@@ -31,35 +31,10 @@ library(MaizeGO)
 library(MaizeOmics)
 library(MaizeMap)
 
-## Import the raw data from the parsed SynMap output
-syntelogs.sorghum.v1.maize.v1.raw <- read_delim("./Data/SynMap/sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw <- read_delim("./Data/SynMap/sorghum_v3.1_vs_maize_v4+rejected.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-
-## Expression data from the Walley 2016 paper in FPKM for 23 tissues
-data("maize.walley.expression.replicate", package = "MaizeOmics")
-maize.expression.clean <- maize.walley.expression.replicate
-maize.expression.sample.avg <- maize.walley.expression
-
-## Protein data from the Walley 2016 paper in dNSAF for 33 tissues
[…]

### 2017-12-21 R/pathway_analysis.R

@@ -1,61 +1,12 @@
 #--------------------------------------------------------------------------------------------------#
 # Pathways
 #--------------------------------------------------------------------------------------------------#
-## Load
-corncyc.pathways.raw <- read_delim("./Data/Pathways/CornCyc_8.0.1_Pathways.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-corncyc.reactions.raw <- read_delim("./Data/Pathways/CornCyc_8.0.1_Reactions.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
-corncyc.geneids.raw <- read_delim("./Data/Pathways/CornCyc_8.0.1_GeneIDs.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+# ## Load
+corncyc.geneids.clean <- MaizeMap::corncyc.gene.map
+corncyc.reactions.clean <- MaizeMap::corncyc.reaction.gene.map
+corncyc.pathways.clean <- MaizeMap::corncyc.pathway.reaction.map
 
-## Clean
-corncyc.pathways.clean <- corncyc.pathways.raw
-
-# Remove superpathways, remove pathways with no genes, unnest the reaction list
-corncyc.pathways.clean <-
-  corncyc.pathways.clean %>%
-  subset(is.na(SubPathways)) %>%
-  subset(!is.na(Reactions)) %>%
-  select(PathwayID, Reactions) %>%
-  mutate(Reactions = gsub("\"","",Reactions)) %>%
-  mutate(ReactionID = strsplit(as.character(Reactions), " // ")) %>%
-  unnest(ReactionID) %>%
-  select(PathwayID, ReactionID)
-
-corncyc.reactions.clean <- corncyc.reactions.raw
-
-# Remove superpathways, remove pathways with no genes, unnest the reaction list
-corncyc.reactions.clean <-
-  corncyc.reactions.clean %>%
-  subset(!is.na(Genes)) %>%
-  mutate(GeneID = strsplit(as.character(Genes), " // ")) %>%
-  unnest(GeneID) %>%
-  select(ReactionID, GeneID)
-
-corncyc.geneids.clean <- corncyc.geneids.raw
-
-# Collect only GRMZM and Zm names which can be in either the Accession1 or Accession2 column
-corncyc.geneids.clean$v3_id <- NA
-corncyc.geneids.clean$v4_id <- NA
[…]

### 2017-12-21 R/staging_testing.R

@@ -36,6 +36,19 @@ subject <- "ATGGAGATCCTCGAATCGACCCTGTTGGGCGAGTTCATCGGCTTCATCAAGGGGAACTGGTCAGCGCA
 cmdNeedle(query, subject)
 
 #--------------------------------------------------------------------------------------------------#
+# GO
+#--------------------------------------------------------------------------------------------------#
+df1 <- MaizeGO %>%
+  subset(evCode=="EXP" & source=="MaizeCyc") %>%
+  select(geneID, goTerm) %>%
+  distinct()
+
+df2 <- MaizeGO %>%
+  subset(evCode=="EXP") %>%
+  select(geneID, goTerm) %>%
+  distinct()
+
+#--------------------------------------------------------------------------------------------------#
 # Pathways
 #--------------------------------------------------------------------------------------------------#
 

### 2018-05-18 Manuscript Figures.Rmd
New report for making final manuscript figures
Since the work done in the analysis/exploratory phase was hit and miss, filter unnecessary stuff and clean up the plots to be publication ready.

--- 2017-12-21 SubGenome Analysis.Rmd
@@
 ---
-title: 'Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating
-  subgenome assignments in maize'
-output:
-  html_document: default
-  pdf_document: default
+title: "Manuscript Figures"
+output: html_document
 ---
@@
 
-We begin with SynMap output based on the [Schnable 2011](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3053962/) paper (see [my notes](https://www.evernote.com/shard/s303/nl/52987123/1ce74f11-98a0-4854-b363-3172e79eeae3) for details on parameter and download options).
-
-Expression data was retrieved from [NCBI: GSE50191](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE50191) and was reported in [Walley 2016](http://science.sciencemag.org/content/353/6301/814.full).
-This set includes 69 experiments (3 biological replicates, 23 tissues) for 62,547 mRNAs aligned to Zea mays B73 RefGen_v2 5a WGS annotation set.
-
-Protein abundance data was retrieved from [Walley 2016](http://science.sciencemag.org/content/353/6301/814.full) supplamental table S2.
-This dataset contains 3 to 7 biological replicates over 33 tissues.
-Expression data is based on a subset of this protein data tissues.
-Values are in dNASF, which is summarized [here](http://research.stowers.org/proteomics/Quant.html).
-
 # 1 Import SynMap and GO Term Data
 ***
-We parse the SynMap output file with ks/kn values (i.e.
-a DAGChainer file) to a flat format using a perl script (parseKsKnFile.pl) and import the data into R.
-Some ks and kn values are empty, which is to be expected.
-These are ignored for the purposes of finding the median, but included in the results if they are in a block that meets the cutoff criteria.
-GO terms related to Maize gene models are imported as well.
 
@@
 ```
-
-Lets take a look at what we just imported.
-
-***  
- * There were **`r syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean %>% nrow()`** rows of data in the SynMap file.
- * There were **`r syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean %>% select(gene2) %>% distinct() %>% nrow()`** distinct Maize genes with synteny to Sorghum.
- * Summary of the KS values
-     + **`r syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean %>% select(ks) %>% summary()`**
- * Out of the `r syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean %>% select(gene2) %>% distinct() %>% nrow()` distinct Maize genes with synteny to sorghum, we keep **`r homeologs.genes %>% select(gene2) %>% distinct() %>% nrow()`** after applying the KS cutoff.
- * SubGenome 1
-     + **`r subgenome %>% filter(subgenome == "sub1") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 1
-     + **`r subgenome %>% filter(subgenome == "sub1") %>% filter(gene2 %in% expressedGenes$geneID) %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned            to subgenome 1 that are expressed significantly on average over 68 samples
-     + **`r goAnnotations.sub1 %>% nrow()`** GO assignments using **`r goAnnotations.sub1 %>% select(goTerm) %>% distinct() %>% nrow()`** GO Terms
- * SubGenome 2
-     + **`r subgenome %>% filter(subgenome == "sub2") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 2
-     + **`r subgenome %>% filter(subgenome == "sub2") %>% filter(gene2 %in% expressedGenes$geneID) %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned            to subgenome 2 that are expressed significantly on average over 68 samples
-     + **`r goAnnotations.sub2 %>% nrow()`** GO assignments using **`r goAnnotations.sub2 %>% select(goTerm) %>% distinct() %>% nrow()`** GO Terms
- * Rejected
-     + **`r subgenome %>% filter(subgenome == "rejected:small") %>% select(gene2) %>% distinct() %>% nrow()`** genes were rejected during subgenome parsing
-     (Too few syntenic genes across chromosome)
 

### 2018-05-25 Manuscript Figures.Rmd
Cleaning up new report for just reporting manuscript figures
This update leave the original subgenome report broken, but makes the manuscript figures report functional.  Makes use of updated MaizeMap/MaizeGO/MaizeOmics versions in the figures, cleans up a lot of the unnecessary analysis.

@@
 
-# All EV-Codes, GOSlim only
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm), 
-                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm)) %>% 
-              nrow(), 
-  category = c("Sub1", "Sub2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim only)")
+# # All EV-Codes, GOSlim only
+# grid.newpage()
+# g <- draw.pairwise.venn(
+#   area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
+#   area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
+#   cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm), 
+#                          goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm)) %>% 
+#               nrow(), 
+#   category = c("Sub1", "Sub2"), 
+#   lty = rep("blank", 2), 
+#   fill = c("light blue", "green"), 
+#   alpha = rep(0.5, 2), 
+#   cat.pos = c(0, 180), 
+#   scaled = FALSE,
+#   euler.d = FALSE, 
+#   sep.dist = 0.3, 
+#   rotation.degree = 0,
+#   ind = FALSE
+# )
+# grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim only)")
 
@@
 
-#EXP only, GOSlim only
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(evCode=="EXP") %>% select(goTerm), 
-                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm)
-                         ) %>% nrow(), 
-  category = c("Sub1", "Sub2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim only)")
+# #EXP only, GOSlim only
+# grid.newpage()
+# g <- draw.pairwise.venn(
+#   area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+#   area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
+#   cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(evCode=="EXP") %>% select(goTerm), 
+#                          goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm)
+#                          ) %>% nrow(), 
+#   category = c("Sub1", "Sub2"), 
+#   lty = rep("blank", 2), 
+#   fill = c("light blue", "green"), 
+#   alpha = rep(0.5, 2), 
+#   cat.pos = c(0, 180), 
+#   scaled = FALSE,
+#   euler.d = FALSE, 
+#   sep.dist = 0.3, 
+#   rotation.degree = 0,
+#   ind = FALSE
+# )
+# grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim only)")
 
@@
 ```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
-plot <- graphGenePairDominanceByExperiment(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
+maize.walley.v4mapped.expression.data <- 
+  maize.walley.v4mapped.expression %>%
+  select(geneID, sample, FPKM_avg)
+plot <- graphGenePairDominanceByExperiment(maize.walley.v4mapped.expression.data, homeologs.pairs, 4, FALSE)
 plot(plot)
@@
 factor <- 4
-plot <- graphGenePairDominanceByExperiment(maize.protein.abundance.sample.avg, homeologs.pairs, factor, FALSE) + 
+maize.walley.abundance.v4.data <- 
+  maize.walley.abundance.v4 %>%
+  select(geneID, sample, dNSAF_avg)
+plot <- graphGenePairDominanceByExperiment(maize.walley.abundance.v4.data, homeologs.pairs, factor, FALSE) + 
   labs(title = paste0("Which homeolog has greater protein abundance? (factor=",factor,")"))
@@
 ```{r subgenome_expression_dominance_across_factors, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=12, fig.align='center', fig.cap = cap}
-plot <- graphGenePairDominanceByExperimentAcrossFactors(maize.expression.sample.avg, homeologs.pairs, FALSE)
+maize.walley.v4mapped.expression.data <- 
+  maize.walley.v4mapped.expression %>%
+  select(geneID, sample, FPKM_avg)
+plot <- graphGenePairDominanceByExperimentAcrossFactors(maize.walley.v4mapped.expression.data, homeologs.pairs, FALSE)
 plot(plot)
@@
 cap <- "Fig 9.4: "
-```
-
-Looking at a few individual cases (* these are calculated as maize1 > maize2, since maize2 > maize1 makes a negative fold change):
-
-```{r subgenome_expression_top_ten, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-gene.foldchange <- homeologFoldChanges(maize.expression.sample.avg, homeologs.pairs)
-gene.foldchange$Maize1_url <- apply(gene.foldchange[,1], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
-gene.foldchange$Maize2_url <- apply(gene.foldchange[,2], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
-gene.foldchange[1:10,] %>% 
-  select(Maize1_url, Maize2_url, sample, Value_maize1, Value_maize2, foldChange) %>%
-  knitr::kable(caption = "Top ten differentially expressed gene pairs (in log2)", align = "llrrr")
-
-protein.foldchange <- homeologFoldChanges(maize.protein.abundance.sample.avg, homeologs.pairs)
-protein.foldchange$Maize1_url <- apply(protein.foldchange[,1], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
-protein.foldchange$Maize2_url <- apply(protein.foldchange[,2], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
-protein.foldchange[1:10,] %>% 
-  select(Maize1_url, Maize2_url, sample, Value_maize1, Value_maize2, foldChange) %>%
-  knitr::kable(caption = "Top ten differentially abundant protein pairs (in log2)", align = "llrrr")
-
-gene.foldchange[1:50,] %>%
-  inner_join(protein.foldchange[1:50,], by = c("Maize1", "Maize2", "sample")) %>%
-  rename(foldChange_expr=foldChange.x, foldChange_prot=foldChange.y, Maize1_url=Maize1_url.x, Maize2_url=Maize2_url.x) %>%
-  select(Maize1_url, Maize2_url, sample, foldChange_expr, foldChange_prot) %>%
-  knitr::kable(caption = "Gene pairs in the top 50 for both differential gene expression and differential protein abundance (in log2).", align = "llrrrr")
-```
-
-Look at the top 10 pairs with highest isoform count diffs (foldChange in log2)
-
-```{r subgenome_isoforms_top_ten, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-data <-
-  homeologs.pairs %>%
-  subset(Maize1 != "" & Maize2 != "") %>%
-  select(Maize1, Maize2) %>%
-  distinct() %>%
-  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
-  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
-names(data)[3] <- "Value_maize1"
-names(data)[4] <- "Value_maize2"
-data <- subset(data, !is.na(data$Value_maize1) & !is.na(data$Value_maize2))
-
-nSingleExonGenes.total <- nrow(geneTranscript.counts[geneTranscript.counts$n == 1,])
-nSingleExonGenes.inSet <- unique(c(data$Maize1[data$Value_maize1 == 1], data$Maize2[data$Value_maize2 == 1]))
-
-data <- subset(data, data$Value_maize1 > 1 & data$Value_maize2 > 1) ## Ignore single-exon genes
-data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)
-data <- arrange(data, desc(foldChange))
-data$Maize1_url <- apply(data[,1], 2, function(x) { wrapInHREF(getGBrowsePageURL(x), x) })
-data$Maize2_url <- apply(data[,2], 2, function(x) { wrapInHREF(getGBrowsePageURL(x), x) })
-data[1:10,] %>%
-  select(Maize1_url, Maize2_url, Value_maize1, Value_maize2, foldChange) %>%
-  knitr::kable(caption = "Top 10 pairs with highest isoform count diffs (foldChange in log2)", align = "llrrr")
 ```
@@
 ```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
+maize.walley.v4mapped.expression.data <- 
+  maize.walley.v4mapped.expression %>%
+  select(geneID, sample, FPKM_avg)
 data <-
@@
   distinct() %>%
-  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
-  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
   subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
@@
   distinct() %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID")) %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
   subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
@@
 ## Start with the paired expression data
+maize.walley.v4mapped.expression.data <- 
+  maize.walley.v4mapped.expression %>%
+  select(geneID, sample, FPKM_avg)
 data <-
@@
   data %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "sample"="sample")) %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
   subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
@@
 
-# 11 Data Exploration
-
-```{r include=FALSE}
-data <-
-  homeologs.pairs %>%
-  subset(Maize1 != "" & Maize2 != "") %>%
-  select(Maize1, Maize2) %>%
-  distinct() %>%
-  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
-  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
-  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
-names(data)[4] <- "FPKM_maize1"
-names(data)[5] <- "FPKM_maize2"
-
-## Add paired abundance data
-data <-
-  data %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "sample"="sample")) %>%
-  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
-  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
-names(data)[6] <- "dNSAF_maize1"
-names(data)[7] <- "dNSAF_maize2"
-
-data$foldChange_expr <- log2(data$FPKM_maize1) - log2(data$FPKM_maize2)
-data$foldChange_abun <- log2(data$dNSAF_maize1) - log2(data$dNSAF_maize2)
-```
-
-```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./data-raw/Images/High expressing genes tend to vary.png')
-```
-
-***
-
-```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./data-raw/Images/Low Expressing Genes tend to stay low.png')
-```
-
-***
-
-```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./data-raw/Images/Pollen is different.png')
-```
-
-***
-
-```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./data-raw/Images/High-low variation.png')
-```
-
-Red Square = [Zm00001d020592](https://www.maizegdb.org/gene_center/gene/Zm00001d020592) Pericarp/Aleurone 27 DAP vs. [Zm00001d005793](https://www.maizegdb.org/gene_center/gene/Zm00001d005793)
-
-Yellow Square = [Zm00001d003188](https://www.maizegdb.org/gene_center/gene/Zm00001d003188) Pericarp/Aleurone 27 DAP vs. [Zm00001d025753](https://www.maizegdb.org/gene_center/gene/Zm00001d025753)
-
-Yellow X = Zm00001d039040 Mature Leaf 8
-
-Pink Cross = Zm00001d015385 Mature Leaf 8
-
-***
-
-```{r echo=FALSE, out.width='100%'}
-knitr::include_graphics('./data-raw/Images/Top fold-change genes.png')
-```
-
 # 12 Wrapup

### 2018-05-25 R/cleanData.R

@@ -23,7 +23,7 @@
 ####################################################################################################
 library(tidyr)
 library(dplyr)
-# startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
+startsWith = getFromNamespace("startsWith", "backports") # if R version < 3.3.0
 
 #==================================================================================================#
 ## maize.genes.v3_to_v4.map

### 2018-05-25 R/do.R

@@ -25,13 +25,9 @@
 source("./R/functions.R")
 # saveMyObjects() # only need to run this is there is a change in raw data or the cleaning procedures
 loadMyObjects()
+source("./R/parseSubgenomes.R")
 library(tidyr)
 library(dplyr)
-
-#==================================================================================================#
-## parseSubgenomes.R
-#--------------------------------------------------------------------------------------------------#
-source("./R/parseSubgenomes.R")
 
 #==================================================================================================#
 ## addGOTerms.R
@@ -138,12 +134,12 @@ expressedPairs <-
 #==================================================================================================#
 ## analyzeGODiffs.R
 #--------------------------------------------------------------------------------------------------#
-source("~/git/SubGenomes/R/analyzeGODiffs.R")
+# source("~/git/SubGenomes/R/analyzeGODiffs.R")
 
 #==================================================================================================#
 ## createTopGO.R
 #--------------------------------------------------------------------------------------------------#
-source("~/git/SubGenomes/R/createTopGO.R")
+# source("~/git/SubGenomes/R/createTopGO.R")
 
 #--------------------------------------------------------------------------------------------------#
 detach("package:tidyr", unload=TRUE)

### 2018-05-31 Manuscript Figures.Rmd
Attempting to put correlation values in faceted graphs

@@
   geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
   scale_x_log10() + scale_y_log10() +
@@
   theme(legend.position = "none") + facet_wrap(~sample, nrow=6)
+  # + annotate("text", size=2, x=10000,y=10, label=paste("pearson: ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
 plot(plot)
@@
 ggplot(data, aes(FPKM_maize1, dNSAF_maize1)) +
-  geom_point(alpha=.1) +
+  geom_point(alpha=.1, aes(color = sample)) +
   geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
@@
   scale_x_log10() + scale_y_log10() +
-  facet_wrap(~sample, nrow=6) +
+  theme(legend.position = "none") + facet_wrap(~sample, nrow=6)
   labs(

### 2018-06-04 Manuscript Figures.Rmd
Updating manuscript figures

--- 2018-05-25 Manuscript Figures.Rmd
@@
 ```
-
-# 1 Import SynMap and GO Term Data
-***
 
@@
 
-# 2 Choosing a KS cutoff value
-***  
-We need to select a cutoff value for ks (synonymous mutation rate) to differentiate between the most recent alpha duplication event and the previous beta duplication.
-Looking at ks value frequency (left), we should not see any obvious cutoff point.
-By transforming ks values to block median values, we should see two (or more) peaks (center).
-In this (center) graph, the first peak represents orthologs and the second represents homeologs from "pregrass tetraploidy".
-If the vertical line separates the first two peaks sufficently, it is probably a good enough value.
-If not, we need to pick a new cutoff value for ks.
-Viewing the log10 transform of the ks values helps differentiate types of homologs (better version of this graph with color can be generated in CoGe).
+# 1 Chromosome location of Maize1 and Maize2
 
-You have selected the log10(ks) cutoff value to be **`r log10_ks_cutoff`**.  Again, this should ideally separate two peaks in the middle graph.
+```{r chromosome_plot, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=4}
+v4_subg_jw <-
+  subgenome %>%
+  left_join(positions, by=c("gene2"="geneID")) %>%
+  select(Chromosome, Pos_start, Pos_end, subgenome) %>%
+  rename(Subgenome=subgenome) %>%
+  subset(Subgenome %in% c("sub1", "sub2"))
 
-```{r pickKSCutoff, echo=FALSE, fig.width=3, fig.height=3}
-qplot(
-  syntelogs.mutated$ks[syntelogs.mutated$ks < 2.5], 
-  geom="histogram", binwidth=.05, na.rm=TRUE
-) + geom_vline(
-  mapping = NULL, 
-  data = NULL, 
-  xintercept = 10^log10_ks_cutoff, 
-  na.rm = FALSE, 
-  show.legend = NA
-) + labs(
-  title="Frequency of KS values by \nsyntenic block (KS < 2.5)", 
-  x="KS value of syntenic block", 
-  y="Count (# of genes)"
+v4_dom_jw <- expressedPairs
+v4_dom_jw$Subgenome <- "no dominance"
+v4_dom_jw$Subgenome[v4_dom_jw$FPKM_mean1 / v4_dom_jw$FPKM_mean2 > 2] <- "dom"
+v4_dom_jw$Subgenome[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2] <- "dom"
+v4_dom_jw$Maize1[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2] <- v4_dom_jw$Maize2[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2]
+df0 <- v4_dom_jw[v4_dom_jw$Subgenome %in% c("no dominance"),]
+df0$Maize1 <- df0$Maize2
+v4_dom_jw <- bind_rows(v4_dom_jw, df0)
+v4_dom_jw <-
+  v4_dom_jw %>%
+  select(Maize1, Subgenome) %>%
+  left_join(positions, by=c("Maize1"="geneID")) %>%
+  select(Chromosome, Pos_start, Pos_end, Subgenome)
+
+## 6,773 - 3,875 (no map) -~-  1,018 (no positions); split/merge is fine = 1,896 orphan v4_ids with positions based on Zea_mays.AGPv4.36.chr.gff3
+v4_orph_jw <-
+  OrphanGenes %>%
+  select(gene) %>%
+  left_join(maize.genes.v3_to_v4.map, c("gene" = "v3_id"))
+v4_orph_na <- v4_orph_jw[is.na(v4_orph_jw$v4_id),]
+v4_orph_jw <- v4_orph_jw[!is.na(v4_orph_jw$v4_id),]
+v4_orph_jw <-
+  v4_orph_jw %>%
+  select(v4_id) %>%
+  left_join(positions, by=c("v4_id"="geneID"))
+v4_orph_jw <- v4_orph_jw[!is.na(v4_orph_jw$Pos_end),]
+v4_orph_jw$Subgenome <- "orphan"
+v4_orph_jw <-
+  v4_orph_jw %>%
+  select(Chromosome, Pos_start, Pos_end, Subgenome)
+
+centromere.v4.positions <-
+  centromere.v4.positions %>%
+  rename(Chromosome="X__1", Pos_start="X__6", Pos_end="X__7") %>%
+  select(Chromosome, Pos_start, Pos_end)
+centromere.v4.positions
+
+df1 <- centromere.v4.positions
+df1$x <- -.45
+df1$xend <- .45
+df2 <- v4_subg_jw
+df2$x <- -.45
+df2$xend <- -.05
+df3 <- v4_dom_jw
+df3$x <- .05
+df3$xend <- .45
+segment_data <- bind_rows(df2,df3)
+segment_data <- segment_data[!is.na(segment_data$Pos_start),]
+segment_data <-
+  segment_data %>%
+  ungroup()
+segment_data <- segment_data %>% subset(Subgenome %in% c("sub1", "sub2", "orphan", "sub1dom", "sub2dom", "dom"))
+centromere_data <- df1
+
+#--------------------------------------------------------------------------------------------------#
+## Make the plot
+p <-
+  ggplot(data=v4) +
+    geom_bar(aes(v4$chromosome, v4$size), stat="identity", fill="grey70") +
+    # geom_segment(data=segment_data, aes(x=Chromosome+x, xend=Chromosome+xend, y=Pos_start, yend=Pos_end, colour=Subgenome), size=.001) +
+    geom_rect(data=segment_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end, color=Subgenome, fill=Subgenome)) +
+    geom_rect(data=centromere_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end)) +
+    scale_fill_manual(values=c("green", "red", "blue", "black")) +
+    scale_color_manual(values=c("green", "red", "blue", "black")) +
+    scale_x_discrete(limits=c("Chr1","Chr2","Chr3","Chr4","Chr5","Chr6","Chr7","Chr8","Chr9","Chr10")) +
+    scale_y_continuous(labels=format_si())
+p  +
+  labs(
+  # title="Subgenome Locations by Chromosome",
+  x="Chromosome",
+  y="Position on Chromosome"
 )
 
-qplot(
-  syntelogs.mutated$median_ks[syntelogs.mutated$median_ks < 2.5], 
-  geom="histogram", 
-  binwidth=.05, 
-  na.rm=TRUE
-) + geom_vline(
-  mapping = NULL, 
-  data = NULL, 
-  xintercept = 10^log10_ks_cutoff, 
-  na.rm = FALSE, 
-  show.legend = NA
-) + labs(
-  title="Frequency of median KS \nvalues by syntenic block \n(median KS < 2.5)", 
-  x="Median KS value of syntenic block", 
-  y="Count (# of genes)"
-)
-
-qplot(
-  log10(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$ks), 
-  geom="histogram", 
-  binwidth=.01, 
-  na.rm=TRUE
-) + geom_vline(
-  mapping = NULL, 
-  data = NULL, 
-  xintercept = log10_ks_cutoff, 
-  na.rm = FALSE, 
-  show.legend = NA
-) + labs(
-  title="Frequency of log10 KS \nvalues", 
-  x="log10(KS) value of genes", 
-  y="Count (# of genes)"
-)
-cap <- "Fig 2.1: "
 ```
 
-# 3 Major Syntenic Blocks
-***
-Regions of synteny are not confined to a 1:1 relationship between Sorghum and Maize chromosomes.
-Where multiple regions in Maize are syntenic to the same region on Sorghum, we can assume this is a result of the whole genome duplication even.
-These are the regions which will be sorted into subgenome 1 and subgenome 2.
-
-```{r reportSyntenicRegions, echo=FALSE, results="asis", fig.cap = cap}
-## Display chromosomes with homeologous genes between org1 and org2
-homeologs.chromosome %>%
-  select(org_chr1, org_chr2, chromosomeGeneCount) %>%
-  spread(org_chr2, chromosomeGeneCount) -> homeologs.spread
-
-homeologs.spread$total <- rowSums(homeologs.spread[-1], na.rm = TRUE)
-
-# De-emphasize na's by turning them light grey
-homeologs.spread.formattedTable <- data.frame(rbind(homeologs.spread[-c(2),], homeologs.spread[c(2),]))
-homeologs.spread.formattedTable <- cbind(homeologs.spread.formattedTable[,-c(3, 12)], homeologs.spread.formattedTable[,c(3, 12)])
-homeologs.spread.formattedTable <- rename(homeologs.spread.formattedTable, Chromosome=org_chr1)
-homeologs.spread.formattedTable[is.na(homeologs.spread.formattedTable)] <- "<font color=\"lightgrey\">na</font>"
-homeologs.spread.formattedTable %>% 
-  knitr::kable(caption = "Table", 
-               align = "lccccccccccr")
-cap <- "Table 3.1: "
-```
-
-There are **<font color=\"darkgreen\">`r homeologs.spread %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0)`</font>** unaligned sorghum chromosome(s) in the above table.
-If there are any unaligned sorghum chromosomes, consider picking a new ks cutoff.
-
-# 4 SubGenomes
-***
-In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size. 
-Psuedocode of the algorithm is as follows: foreach sorghum chromosome, set the largest syntenic chromosome (by gene) as sub1.
-Foreach additional syntenic chromosome if doesn't overlap what is already in sub1, add it to sub1.
-Else, if it does not overlap anything in sub2, add to sub2.
-Else toss.
-After all syntenic chromosomes are process for this sorghum chromosome, count the number of genes in sub1 and sub2.
-If sub2 is bigger, switch them.
-*Toss out "small" syntenic chromosomes (51 genes or less?), and consider a tolerance for how much overlap must occur to consider it a true overlap.
-Also consider switching to gene level overlap rather than whole-chromosome start/stop values.
-
-```{r reportSubGenomes, echo=FALSE, results="asis"}
-#Positions on each chr with sub 1 or sub 2
-#Total genes for each subgenome by chr and overall
-```
-
-In order to see if one subgenome tends to have more isoforms, we plot the density of isoform counts from each subgenome based on isoforms in the v4 release 32 GFF file.
-We scale the x axis using log10, as we have a very long tail on this data (i.e.
-most counts are 1, but a few range to nearly 600).
-The densities track each other very strongly, so there isn't likely anything to look for here.
-
-```{r subgenome_isoform_counts, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=6, fig.height=4, fig.align='center'}
-## Plot frequency of isoform counts for each gene
-data <- geneTranscript.counts %>% inner_join(subgenome, by=c("gene"="gene2")) %>% select(gene, n, subgenome) %>% filter(subgenome=="sub1" | subgenome=="sub2")
-ggplot(data, aes(n,color=subgenome)) + 
-  geom_density() + 
-  scale_x_log10() +
-  labs(y = "Density (Isoform Counts)", 
-       x = "log10(Isoform Counts)", 
-       title = "Comparison of Density of Isoform Counts between\nGenes in Subgenome 1 and Subgenome 2"
-  )
-```
-
-## 5 Compared to Paper
-How many chromosomes with synteny to sorghum were correctly placed in subgenome 1 or subgenome 2?
-Items in "Paper" only are missed subgenome assignments.
-Items in "Greedy" only are false assignments.
-Items in center are correct.
-(currently sorted by hand)
-
-```{r compareResultsToPaper_chromosomes, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=3, fig.align='center'}
-subgenome.chromosomes.test <- 
-  subgenome.assignments %>% 
-  mutate(chr1=as.numeric(regmatches(org_chr1, regexpr("\\d*$",org_chr1)))) %>%
-  mutate(chr2=as.numeric(regmatches(org_chr2, regexpr("\\d*$",org_chr2)))) %>%
-  select(chr1, chr2, subgenome) %>%
-  subset(subgenome == "sub1" | subgenome == "sub2")
-
-grid.newpage()
-g = draw.pairwise.venn(
-  area1 = subgenome.truth %>% nrow(), 
-  area2 = subgenome.chromosomes.test %>% nrow(), 
-  cross.area = intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow(), 
-  category = c("Paper", "Greedy"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="Overlap of Subgenome Assignments vs. \"Truth\"")
-```
-
-<!-- print("Missed subgenome assignments:") -->
-<!-- setdiff(subgenome.truth, subgenome.chromosomes.test) -->
-
-<!-- print("False positive subgenome assignments:") -->
-<!-- setdiff(subgenome.chromosomes.test, subgenome.truth) -->
-
-# 6 GO Terms in Maize
-***
-
-First we show both how many unique GO terms have been assigned using a computational ev-code vs.
-an experimental ev-code.
-A few terms have been annotated using both or were assigned multiple times based on different publications, which is why n is greater than the number of unique GO terms in our set.
-We also break down the GO terms by type.
-A few terms may not have a type, typically because they have been made obsolete.
-
-Second, we look at the same divisions but this time for all gene annotations (i.e.
-gene + GO Term + Ev-Code).
-Again, since some terms are assigned computationally and experimentally to the same gene or assigned multiple times based on different publications, n is greater than the number of unique GO term assignments in out set.
-The main takeaways here are that computationally assigned GO terms vastly outnumber experimental annotations, most annotations use MF terms, and there are a few terms falling through the cracks most likely due to being outdated.
-
-```{r GOTerms_in_Maize, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=2.3, fig.height=3.6}
-Evidence <- c("EV-EXP","EV-COMP")
-Count <- c(go.maize.clean %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-      go.maize.clean %>% filter(evCode=="COMP") %>% select(goTerm) %>% distinct() %>% nrow())
-df = data.frame(Evidence,Count)
-ggplot(data=df, aes(x=Evidence, y=Count)) +
-  geom_bar(stat="identity", fill="steelblue") +
-  geom_text(aes(label=Count), vjust=-0.3, size=3.5
-) + labs(
-  title=paste("GO Terms by EV Code\n(n=", nrow(unique(go.maize.clean[,c("goTerm", "evCode")])),")", sep = "")
-) + theme(plot.title = element_text(size = rel(.8)))
-
-Type <- c("MF","BP","CC","<NA>")
-Count <- c(go.maize.clean %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
-           go.maize.clean %>% filter(type=="BP") %>% select(goTerm) %>% distinct() %>% nrow(),
-           go.maize.clean %>% filter(type=="CC") %>% select(goTerm) %>% distinct() %>% nrow(),
-           go.maize.clean %>% filter(is.na(type)) %>% select(goTerm) %>% distinct() %>% nrow())
-df = data.frame(Type,Count)
-ggplot(data=df, aes(x=Type, y=Count)) +
-  geom_bar(stat="identity", fill="steelblue") +
-  geom_text(aes(label=Count), vjust=-0.3, size=3.5
-) + labs(
-  title=paste("GO Terms by Type\n(n=", nrow(unique(go.maize.clean[,c("goTerm", "type")])),")", sep = "")
-) + theme(plot.title = element_text(size = rel(.8)))
-
-Evidence <- c("EV-EXP","EV-COMP")
-Count <- c(go.maize.clean %>% filter(evCode=="EXP") %>% nrow(),
-      go.maize.clean %>% filter(evCode=="COMP") %>% nrow())
-df = data.frame(Evidence,Count)
-ggplot(data=df, aes(x=Evidence, y=Count)) +
-  geom_bar(stat="identity", fill="steelblue") +
-  geom_text(aes(label=Count), vjust=-0.3, size=3.5
-) + labs(
-  title=paste("GO Term Annotations by\nEV Code (n=", nrow(go.maize.clean),")", sep = "")
-) + theme(plot.title = element_text(size = rel(.8)))
-
-Type <- c("MF","BP","CC","<NA>")
-Count <- c(go.maize.clean %>% filter(type=="MF") %>% select(geneID, goTerm, evCode) %>% distinct() %>% nrow(),
-           go.maize.clean %>% filter(type=="BP") %>% select(geneID, goTerm, evCode) %>% distinct() %>% nrow(),
-           go.maize.clean %>% filter(type=="CC") %>% select(geneID, goTerm, evCode) %>% distinct() %>% nrow(),
-           go.maize.clean %>% filter(is.na(type)) %>% select(geneID, goTerm, evCode) %>% distinct() %>% nrow())
-df = data.frame(Type,Count)
-ggplot(data=df, aes(x=Type, y=Count)) +
-  geom_bar(stat="identity", fill="steelblue") +
-  geom_text(aes(label=Count), vjust=-0.3, size=3.5
-) + labs(
-  title=paste("GO Term Annotations by\nType (n=", nrow(unique(go.maize.clean[,c("geneID","goTerm","evCode","type")])),")", sep = "")
-) + theme(plot.title = element_text(size = rel(.8)))
-```
-
-# 7 GO Term Comparison by Subgenome
+# 2 GO Term Comparison by Subgenome
 ***
@@
 
-```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=3, fig.height=3}
-# All EV-Codes
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% select(goTerm) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% select(goTerm), goAnnotations.sub2 %>% select(goTerm)) %>% nrow(), 
-  category = c("Sub1", "Sub2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes)")
+GO Term assignments to genes by subgenome, which takes into account the gene-GO Term pairings and considers homeologs equivalent for the purposes of determining overlap in assignments.
 
-# # All EV-Codes, GOSlim only
-# grid.newpage()
-# g <- draw.pairwise.venn(
-#   area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
-#   area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
-#   cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm), 
-#                          goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm)) %>% 
-#               nrow(), 
-#   category = c("Sub1", "Sub2"), 
-#   lty = rep("blank", 2), 
-#   fill = c("light blue", "green"), 
-#   alpha = rep(0.5, 2), 
-#   cat.pos = c(0, 180), 
-#   scaled = FALSE,
-#   euler.d = FALSE, 
-#   sep.dist = 0.3, 
-#   rotation.degree = 0,
-#   ind = FALSE
-# )
-# grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim only)")
+The main takeaways here mirror the assessment from the GO Term presence/absence check above.
+There are many differences in assignments, although we can assume that the same terms are reused very often based on the differenced between unique terms and gene annotations.
 
-# All EV-Codes, MF terms only
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% filter(type=="MF") %>% select(goTerm), 
-                         goAnnotations.sub2 %>% filter(type=="MF") %>% select(goTerm)) %>%
-                          nrow(), 
-  category = c("Sub1", "Sub2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, MF only)")
-
-#EXP only
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-  area2 = goAnnotations.sub2 %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% select(goTerm), 
-                         goAnnotations.sub2 %>% filter(evCode=="EXP") %>% select(goTerm)
-                         ) %>% nrow(), 
-  category = c("Sub1", "Sub2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP only)")
-
-# #EXP only, GOSlim only
-# grid.newpage()
-# g <- draw.pairwise.venn(
-#   area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-#   area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
-#   cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(evCode=="EXP") %>% select(goTerm), 
-#                          goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm)
-#                          ) %>% nrow(), 
-#   category = c("Sub1", "Sub2"), 
-#   lty = rep("blank", 2), 
-#   fill = c("light blue", "green"), 
-#   alpha = rep(0.5, 2), 
-#   cat.pos = c(0, 180), 
-#   scaled = FALSE,
-#   euler.d = FALSE, 
-#   sep.dist = 0.3, 
-#   rotation.degree = 0,
-#   ind = FALSE
-# )
-# grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim only)")
-
+```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
 #EXP only, MF terms only
@@
   area2 = goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm), 
-                         goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm)
+  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct(),  
+                         goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct()
                          ) %>% nrow(), 
-  category = c("Sub1", "Sub2"), 
+  category = c("Maize1", "Maize2"), 
   lty = rep("blank", 2), 
@@
 grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, MF only)")
-```
 
-***
-
-GO Term assignments to genes by subgenome, which takes into account the gene-GO Term pairings and considers homeologs equivalent for the purposes of determining overlap in assignments.
-
-The main takeaways here mirror the assessment from the GO Term presence/absence check above.
-There are many differences in assignments, although we can assume that the same terms are reused very often based on the differenced between unique terms and gene annotations.
-
-```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=3, fig.height=3}
 ## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2.  The converted data.frame will hold maize2 genes with the names replaced with maize1 homologs
@@
 
-# All EV-Codes, all assignments
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% nrow(),
-  area2 = converted %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% select(gene2, goTerm) %>% distinct(), 
-                         converted %>% select(gene2, goTerm) %>% distinct()
-                         ) %>% nrow(),
-  category = c("Sub1", "Sub2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(All EV-Codes)")
-
-#EXP only, all assignments
-grid.newpage()
-g <- draw.pairwise.venn(
-  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% nrow(),
-  area2 = converted %>% filter(evCode=="EXP") %>% nrow(),
-  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% select(gene2, goTerm) %>% distinct(), 
-                         converted %>% filter(evCode=="EXP") %>% select(gene2, goTerm) %>% distinct()
-                         ) %>% nrow(), 
-  category = c("Sub1", "Sub2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(EV-EXP only)")
-
 #EXP only, MF only
@@
                          ) %>% nrow(), 
-  category = c("Sub1", "Sub2"), 
+  category = c("Maize1", "Maize2"), 
   lty = rep("blank", 2), 
@@
 
-# 8 Expression
+# 3 Compare Homeolog Pairs
 
 ***
-
-Cases where the maize1 homolog is more or less expressed than the maize2 homolog
-
-```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4, fig.align='center'}
-# Expression Diffs
-grid.newpage()
-a1 = expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) > 2) %>% nrow()
-a2 = expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) > 2) %>% nrow()
-a3 = nrow(expressedPairs) - a1 - a2
-g <- draw.pairwise.venn(
-  area1 = a1+a3,
-  area2 = a2+a3,
-  cross.area = a3, 
-  category = c("Sub1", "Sub2"), 
-  lty = rep("blank", 2), 
-  fill = c("light blue", "green"), 
-  alpha = rep(0.5, 2), 
-  cat.pos = c(0, 180), 
-  scaled = FALSE,
-  euler.d = FALSE, 
-  sep.dist = 0.3, 
-  rotation.degree = 0,
-  ind = FALSE
-)
-grid.arrange(gTree(children=g), top="Genes with 2-fold higher\nexpression by subgenome")
-```
-
-Overall expression patterns between subgenome 1 and subgenome 2 are highly similar.
-Below we show to views of the same data.
-In the density graph, we can see that subgenome 1 is slightly shifted to the right of subgenome 2 expression, but otherwise follows the same shape.
-Viewed as a boxplot, you can again see that the quartiles in subgenome 1 and 2 are largely the same.
-This also shows a few high expressing outliers on subgenome 1.
-
-Paired (two-sided) t-test suggests that there is no difference in means between FPKM of homeologs in sub1 vs. sub2 (p-val > 0.05, so we fail to reject null hypothesis of equal means):
-```{r subgenome_fpkm_ttest, echo=FALSE}
-t.test(x=expressedPairs$FPKM_mean1, y=expressedPairs$FPKM_mean2, paired = TRUE)
-```
-
-```{r subgenome_expression_frequency, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4.5, fig.height=4}
-## Plot frequency of isoform counts for each gene
-data <- expressedGenes %>% inner_join(subgenome, by=c("geneID"="gene2")) %>% select(geneID, FPKM_mean, subgenome) %>% filter(subgenome=="sub1" | subgenome=="sub2")
-ggplot(data, aes(FPKM_mean,color=subgenome)) +
-  geom_density() +
-  scale_x_log10() +
-  labs(y = "Density (FPKM)",
-       x = "log10(FPKM)",
-       title = "Comparison of FPKM Expression between\nGenes in Subgenome 1 and Subgenome 2"
-  )
-
-ggplot(data, aes(subgenome,log10(FPKM_mean))) +
-  geom_boxplot() +
-  labs(y = "log10(FPKM)",
-       x = "Subgenome",
-       title = "Comparison of FPKM Expression between\nGenes in Subgenome 1 and Subgenome 2"
-  )
-```
-
-I am curious if our definition of sub 1 and sub 2 (i.e.
-the larger syntenic block is defined as sub 1) is really the best way to decide which genes are in sub 1 and sub 2.
-It seems to me that if one could resonably expect that, given 2 identically duplicate genes, if there was no benefit to having duplicated function then it would be a matter of random chance which of the two duplicate genes would loose function over time.
-It also seems to me that the quickest way to "degrade" an unnecessary gene (evolutionarily speaking) is to reduce expression.
-So what happens if we pick the higher expressing gene to be in sub 1 and the lower expressing gene to be in sub 2?
-Not surprising, we can find differences in populations between high and low expressing pairs.
-I wonder if this holds across different maize lines?
-
-```{r subgenome_expression_high_v_low, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4.5, fig.height=4, fig.align='center'}
-data <- expressedPairs
-data$LowExpressingGene <- apply(data[,c(3,4)], 1, FUN=min)
-data$HighExpressingGene <- apply(data[,c(3,4)], 1, FUN=max)
-data <- data %>% gather(key = type, value = value, 5:6)
-ggplot(data, aes(type, log10(value))) +
-  geom_boxplot() +
-  labs(y = "log10(FPKM)",
-       x = "Types",
-       title = "Comparison of FPKM Expression between\nGenes in high express vs. low express groups"
-  )
-```
-
-# 9 Compare Homeolog Pairs
-
-***
-
-In all cases where there is a homeolog on both maize1 and maize2, the we want to look to several measures to see if the genes on one subgenome tend to be present in greater abundance compared to the other.
-We can define this comparison as either greater expression, greater protein abundance, or a greater number of isoforms available to the gene.
-We can compare using a factor to define a cutoff for "greater".
-(i.e.
-4-fold higher expression/abundance etc.).
-
-First we look at expression values.
-We find that maize1 consistently overexpresses the maize2 pair on average across all available data, regardless of factor (1,2, or 4), and regardless of whether or not we allow NA values to be treated as '0'.
-
-```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
-maize.walley.v4mapped.expression.data <- 
-  maize.walley.v4mapped.expression %>%
-  select(geneID, sample, FPKM_avg)
-plot <- graphGenePairDominanceByExperiment(maize.walley.v4mapped.expression.data, homeologs.pairs, 4, FALSE)
-plot(plot)
-cap <- "Fig 9.1:"
-```
-
-***
-
-This same pattern holds for protein abundance data as well. Here we look at protein abundance data for the tissues that also have expression data.
-
-```{r subgenome_protein_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
-factor <- 4
-maize.walley.abundance.v4.data <- 
-  maize.walley.abundance.v4 %>%
-  select(geneID, sample, dNSAF_avg)
-plot <- graphGenePairDominanceByExperiment(maize.walley.abundance.v4.data, homeologs.pairs, factor, FALSE) + 
-  labs(title = paste0("Which homeolog has greater protein abundance? (factor=",factor,")"))
-plot(plot)
-cap <- "Fig 9.2:"
-```
-
-We can see this same trend for all experiments across all factors (fold-difference of 1 to 9 shown).
-The blue line (neither gene has greater expression) grows as factor increases since we have a stricter and stricter definition of what counts as greater expression.
-Red (maize1) is always above green (maize2) indicating that more maize1 homeologs have greater expression over their maize2 pair than vice versa for all expriments and all factors.
-Red and green lines both tend to zero when factor is large enough.
-Protein data (not shown) follows this same pattern.
-
-```{r subgenome_expression_dominance_across_factors, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=12, fig.align='center', fig.cap = cap}
-maize.walley.v4mapped.expression.data <- 
-  maize.walley.v4mapped.expression %>%
-  select(geneID, sample, FPKM_avg)
-plot <- graphGenePairDominanceByExperimentAcrossFactors(maize.walley.v4mapped.expression.data, homeologs.pairs, FALSE)
-plot(plot)
-cap <- "Fig 9.3:"
-```
-
-***
-
-The number of isoforms each gene of the homeolog pairs is thought to express follows a pattern where maize1 tends to have more isoforms than maize2 homeologs.
-
-```{r subgenome_isoform_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=6, fig.height=4, fig.align='center', fig.cap = cap}
-## Plot frequency of isoform counts for each gene
-plot <- graphGenePairDominanceByIsoForms(geneTranscript.counts, homeologs.pairs, 4)
-plot(plot)
-cap <- "Fig 9.4: "
-```
 
@@
   labs(
-    title = paste0("Scatterplot of isoform counts for gene pair(s).\nExcludes ",n," outlier, alpha=.1, foldchange in log2, blue=lm fit."),
+    # title = paste0("Scatterplot of isoform counts for gene pair(s).\nExcludes ",n," outlier, alpha=.1, foldchange in log2, blue=lm fit."),
     x = "Isoform Counts for Maize1 Gene",
@@
 
-cap <- "Fig 9.5: "
+cap <- "Fig 3.1: "
+```
+
+```{r subgenome_expression_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=4, fig.align='center', fig.cap = cap}
+maize.walley.v4mapped.expression.data <- 
+  maize.walley.v4mapped.expression %>%
+  select(geneID, sample, FPKM_avg)
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "Value_maize1"
+names(data)[5] <- "Value_maize2"
+# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))
+
+data <- 
+  data %>%
+  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode"))
+
+plot <-
+  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
+  geom_point(alpha=.1, aes(color = sample)) +
+  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
+  scale_x_log10() + scale_y_log10() +
+  labs(
+    # title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1, reference line"),
+    x = "Expression for Maize1 Gene (log scale)",
+    y = "Expression for Maize2 Gene (log scale)"
+  ) +
+  theme(legend.position = "none") + facet_wrap(~sample, nrow=1)
+plot(plot)
+
+cap <- "Fig 3.2: "
 ```
@@
   geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
   scale_x_log10() + scale_y_log10() +
   labs(
-    title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1, reference line"),
+    # title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1, reference line"),
     x = "Expression for Maize1 Gene (log scale)",
@@
 
-cap <- "Fig 9.5: "
+cap <- "Fig 3.3: "
+```
+
+```{r subgenome_protein_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=4, fig.align='center', fig.cap = cap}
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+names(data)[4] <- "Value_maize1"
+names(data)[5] <- "Value_maize2"
+# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))
+data <- 
+  data %>%
+  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode"))
+
+plot <-
+  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
+  geom_point(alpha=.1, aes(color = sample)) +
+  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
+  scale_x_log10() + scale_y_log10() +
+  labs(
+    # title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1, reference line"),
+    x = "Abundance for Maize1 Gene (log scale)",
+    y = "Abundance for Maize2 Gene (log scale)"
+  ) +
+  theme(legend.position = "none") + facet_wrap(~sample, nrow=1)
+  # + annotate("text", size=2, x=10000,y=10, label=paste("pearson: ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
+plot(plot)
+
+cap <- "Fig 3.4: "
 ```
@@
   geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
   scale_x_log10() + scale_y_log10() +
   labs(
-    title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1, reference line"),
+    # title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1, reference line"),
     x = "Abundance for Maize1 Gene (log scale)",
@@
   theme(legend.position = "none") + facet_wrap(~sample, nrow=6)
+  # + annotate("text", size=2, x=10000,y=10, label=paste("pearson: ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
 plot(plot)
 
-cap <- "Fig 9.6: "
+cap <- "Fig 3.5: "
 ```
 
-# 10 Expression and Abundance are Weakly Correlated
+# 4 Expression and Abundance are Weakly Correlated
 
@@
   xlim(-10,10) + ylim(-10,10) +
+  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
   geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
@@
   labs(
-    title = paste0("Scatterplot of Foldchange in Expression in Maize1 vs.\nFoldchange in Abundance in Maize1\nAlpha=.1, linear fit, n=13413"),
+    # title = paste0("Scatterplot of Foldchange in Expression in Maize1 vs.\nFoldchange in Abundance in Maize1\nAlpha=.1, linear fit, n=13413"),
     x = "log2(Expression Maize1) - log2(Expression Maize2)",
@@
   scale_x_log10() + scale_y_log10() +
-  facet_wrap(~sample, nrow=6) +
+  theme(legend.position = "none") + facet_wrap(~sample, nrow=6) +
   labs(
-    title = paste0("Scatterplot of Expression in Maize1 vs. Abundance in Maize1\nLog-Log Scale for gene pair(s).\nAlpha=.1, reference line (black), lm fit (blue)"),
+    # title = paste0("Scatterplot of Expression in Maize1 vs. Abundance in Maize1\nLog-Log Scale for gene pair(s).\nAlpha=.1, reference line (black), lm fit (blue)"),
     x = "Expression for Maize1 Gene (log scale)",
@@
 
-# 12 Wrapup
+# 5 Expression pattern types
+functionalDivergence.R
+
+# 6 Wrapup
 ***

### 2018-06-04 R/analysis/chromosomePlots.R

@@ -15,10 +15,7 @@ library(dplyr)
 library(readr)
 library(readxl)
 
-v4 <- read.csv("./data-raw/ChrImages/Zea_mays.AGPv4.36.sizes.csv")
-v4_subg <- read_csv("data-raw/ChrImages/v4_subg.csv")
-positions <- v4_subg <- read.csv("./data-raw/ChrImages/gene_positions.tab", sep = "\t")
-OrphanGenes <- read_excel("data-raw/ChrImages/OrphanGenes.xlsx")
+
 
 v4_subg_jw <-
   subgenome %>%
@@ -154,29 +151,3 @@ plot(p+labs(
          x="Chromosome",
          y="Position on Chromosome"))
 dev.off()
-
-#--------------------------------------------------------------------------------------------------#
-# Add SI units
-format_si <- function(...) {
-  # Based on code by Ben Tupper
-  # https://stat.ethz.ch/pipermail/r-help/2012-January/299804.html
-
-  function(x) {
-    limits <- c(1e0, 1e3, 1e6, 1e9, 1e12)
-    prefix <- c("b", "kb", "Mb", "Gb", "Tb")
-    # prefix <- c("y",   "z",   "a",   "f",   "p",
-    #             "n",   "µ",   "m",   " ",   "k",
-    #             "M",   "G",   "T",   "P",   "E",
-    #             "Z",   "Y")
-
-    # Vector with array indices according to position in intervals
-    i <- findInterval(abs(x), limits)
-
-    # Set prefix to " " for very small values < 1e-24
-    i <- ifelse(i==0, which(limits == 1e0), i)
-
-    paste(format(round(x/limits[i], 1),
-                 trim=TRUE, scientific=FALSE, ...),
-          prefix[i])
-  }
-}

### 2018-06-04 R/functions.R

@@ -217,3 +217,28 @@ getGBrowsePageURL <- function(geneID) {
 wrapInHREF <- function(url, text) {
   return(paste0("<a href=\"", url, "\">", text, "</a>"))
 }
+
+# Add SI units (for plots to convert 10000000 basepairs into 10Mb basepairs, etc)
+format_si <- function(...) {
+  # Based on code by Ben Tupper
+  # https://stat.ethz.ch/pipermail/r-help/2012-January/299804.html
+
+  function(x) {
+    limits <- c(1e0, 1e3, 1e6, 1e9, 1e12)
+    prefix <- c("b", "kb", "Mb", "Gb", "Tb")
+    # prefix <- c("y",   "z",   "a",   "f",   "p",
+    #             "n",   "µ",   "m",   " ",   "k",
+    #             "M",   "G",   "T",   "P",   "E",
+    #             "Z",   "Y")
+
+    # Vector with array indices according to position in intervals
+    i <- findInterval(abs(x), limits)
+
+    # Set prefix to " " for very small values < 1e-24
+    i <- ifelse(i==0, which(limits == 1e0), i)
+
+    paste(format(round(x/limits[i], 1),
+                 trim=TRUE, scientific=FALSE, ...),
+          prefix[i])
+  }
+}

### 2018-06-04 R/loadData.R

@@ -79,6 +79,12 @@ subgenome.assignments <- read_delim("./data-raw/outfile_processed_byHand.tab", "
 chromosome.v4.size <- read.csv("./data-raw/ChrImages/Zea_mays.AGPv4.36.sizes.csv")
 gene.positions.v4.chromosome <- v4_subg <- read.csv("./data-raw/ChrImages/gene_positions.tab", sep = "\t") # this should really be in MaizeMap
 
+v4 <- read.csv("./data-raw/ChrImages/Zea_mays.AGPv4.36.sizes.csv")
+v4_subg <- read_csv("data-raw/ChrImages/v4_subg.csv")
+positions <- v4_subg <- read.csv("./data-raw/ChrImages/gene_positions.tab", sep = "\t")
+OrphanGenes <- read_excel("data-raw/ChrImages/OrphanGenes.xlsx")
+centromere.v4.positions <- read_excel("data-raw/ChrImages/centromere_positions.xlsx", col_names = FALSE)
+
 #--------------------------------------------------------------------------------------------------#
 detach("package:readr", unload=TRUE)
 detach("package:readxl", unload=TRUE)

### 2018-06-04 R/staging_testing.R

@@ -115,4 +115,22 @@ View(CSHLPairs)
 View(MyPairs)
 
 #--------------------------------------------------------------------------------------------------#
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "Value_maize1"
+names(data)[5] <- "Value_maize2"
+
+data$m1 <- 0
+data$m2 <- 0
+data$m1[data$Value_maize1 > data$Value_maize2] <- 1
+data$m2[data$Value_maize1 < data$Value_maize2] <- 1
+data.temp <- data %>% select(sample, m1, m2) %>% group_by(sample) %>% summarise(vm1=sum(m1), vm2=sum(m2)) %>% ungroup()
+data.temp$per <- data.temp$vm1/(data.temp$vm1 + data.temp$vm2)
+#--------------------------------------------------------------------------------------------------#
 detach("package:fitdistrplus", unload=TRUE)

### 2018-06-06 Manuscript Figures.Rmd
Updated figures

@@
 
-```{r chromosome_plot, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=4}
+***
+
+```{r chromosome_plot, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6}
 v4_subg_jw <-
@@
   ungroup()
-segment_data <- segment_data %>% subset(Subgenome %in% c("sub1", "sub2", "orphan", "sub1dom", "sub2dom", "dom"))
+segment_data$Subgenome[segment_data$Subgenome %in% c("sub1")] <- "Maize1"
+segment_data$Subgenome[segment_data$Subgenome %in% c("sub2")] <- "Maize2"
+# segment_data$Subgenome[segment_data$Subgenome %in% c("dom")] <- "Dominant Homeolog"
+segment_data <- segment_data %>% subset(Subgenome %in% c("Maize1", "Maize2", "orphan", "sub1dom", "sub2dom", "dom"))
 centromere_data <- df1
@@
 ## Make the plot
-p <-
+chromosome_plot <-
   ggplot(data=v4) +
@@
     scale_y_continuous(labels=format_si())
-p  +
+chromosome_plot  +
   labs(
@@
 # 2 GO Term Comparison by Subgenome
+
 ***
-
-GO Terms by subgenome.  If the GO Term is assigned to any gene in the subgenome, it is counted once for that subgenome.  This is a presence/absence test.
-
-The main takeaways here are that 1) there are differences in functional assignments between each subgenome, regardless of which filters I apply.
-2) the GOSlim for plants is extremely restrictive, and probably not useful.
-3) Sub1 has more functional assignments than sub2, although this may be a direct correlation to the larger size (by definition) of sub1.
-
-GO Term assignments to genes by subgenome, which takes into account the gene-GO Term pairings and considers homeologs equivalent for the purposes of determining overlap in assignments.
-
-The main takeaways here mirror the assessment from the GO Term presence/absence check above.
-There are many differences in assignments, although we can assume that the same terms are reused very often based on the differenced between unique terms and gene annotations.
 
@@
 grid.newpage()
-g <- draw.pairwise.venn(
+GOTerms_by_subgenome_a <- draw.pairwise.venn(
   area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
@@
 )
-grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, MF only)")
+grid.arrange(gTree(children=GOTerms_by_subgenome_a), top="GO Term by Subgenome\n(EV-EXP, MF only)")
 
@@
 grid.newpage()
-g <- draw.pairwise.venn(
+GOTerms_by_subgenome_b <- draw.pairwise.venn(
   area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% nrow(),
@@
 )
-grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(EV-EXP, MF only)")
+grid.arrange(gTree(children=GOTerms_by_subgenome_b), top="GO Term Assignments by Subgenome\n(EV-EXP, MF only)")
+
 ```
 
-# 3 Compare Homeolog Pairs
+# 3 Isoforms
 
@@
 
-```{r subgenome_isoforms_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.align='center', fig.cap = cap}
+```{r subgenome_isoforms_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.align='center'}
 # Isoforms
@@
 
-plot <-
+subgenome_isoforms_scatter <-
   ggplot(subset(data, Value_maize1 <= 25 & Value_maize2 <= 25), aes(x=Value_maize1, y=Value_maize2)) +
@@
   annotate("text",x=20,y=3, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
-plot(plot)
+plot(subgenome_isoforms_scatter)
 
-cap <- "Fig 3.1: "
 ```
 
-```{r subgenome_expression_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=4, fig.align='center', fig.cap = cap}
+# 4 Expression
+
+***
+
+```{r subgenome_expression_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=2.5, fig.align='center'}
 maize.walley.v4mapped.expression.data <- 
@@
   data %>%
-  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode"))
+  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode", "B73 Mature Pollen"))
 
-plot <-
+subgenome_expression_scatter_rep <-
   ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
@@
   theme(legend.position = "none") + facet_wrap(~sample, nrow=1)
-plot(plot)
+plot(subgenome_expression_scatter_rep)
 
-cap <- "Fig 3.2: "
 ```
 
-```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
+```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
 maize.walley.v4mapped.expression.data <- 
@@
 
-plot <-
+subgenome_expression_scatter <-
   ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
@@
   theme(legend.position = "none") + facet_wrap(~sample, nrow=6)
-plot(plot)
-
-cap <- "Fig 3.3: "
+plot(subgenome_expression_scatter)
 ```
 
-```{r subgenome_protein_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=4, fig.align='center', fig.cap = cap}
+# 5 Protein Abundance
+
+***
+
+```{r subgenome_protein_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=2.5, fig.align='center'}
 data <-
@@
   data %>%
-  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode"))
+  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode", "B73 Mature Pollen"))
 
-plot <-
+subgenome_protein_scatter_rep <-
   ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
@@
   # + annotate("text", size=2, x=10000,y=10, label=paste("pearson: ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
-plot(plot)
-
-cap <- "Fig 3.4: "
+plot(subgenome_protein_scatter_rep)
 ```
 
-```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
+```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
 data <-
@@
 
-plot <-
+subgenome_protein_scatter <-
   ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
@@
   # + annotate("text", size=2, x=10000,y=10, label=paste("pearson: ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
-plot(plot)
-
-cap <- "Fig 3.5: "
+plot(subgenome_protein_scatter)
 ```
 
-# 4 Expression and Abundance are Weakly Correlated
+# 6 Expression and Abundance are Weakly Correlated
 
-```{r subgenome_expression_protein_foldchange_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=6, fig.height=5, fig.align='center', fig.cap = cap}
+***
+
+```{r subgenome_expression_protein_foldchange_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=6, fig.height=5, fig.align='center'}
 ## Start with the paired expression data
@@
 
+# Remove pollen
+data <-
+  data %>%
+  subset(!sample %in% c("B73 Mature Pollen"))
+
 data$foldChange_expr <- log2(data$FPKM_maize1) - log2(data$FPKM_maize2)
 data$foldChange_abun <- log2(data$dNSAF_maize1) - log2(data$dNSAF_maize2)
-ggplot(data, aes(foldChange_expr, foldChange_abun)) +
+subgenome_expression_protein_foldchange_scatter <- ggplot(data, aes(foldChange_expr, foldChange_abun)) +
   geom_point(alpha=.1) +
@@
   ) + 
-  annotate("text",x=7,y=-8, label=paste("Pearson Cor = ", round(cor(data$foldChange_expr, data$foldChange_abun, method = c("pearson")),2)))
-
-cap <- "Fig 10.1: Considering only homeologous pairs where both genes were measured for expression and protein abundance, we look at fold change from maize2 expression to maize1 expression vs.
-the fold change from maize2 abundance vs.
-maize2 abundance.
-A slightly positive trend indicates a mild positive correlation between an increase in expression and an increase in abundance."
+  annotate("text",x=7,y=-8, label=paste("Pearson Cor = ", round(cor(data$foldChange_expr, data$foldChange_abun, method = c("pearson")),2))) + 
+  annotate("text",x=7,y=-10, label=paste("Spearman Cor = ", round(cor(data$foldChange_expr, data$foldChange_abun, method = c("spearman")),2)))
+plot(subgenome_expression_protein_foldchange_scatter)
 ```
 
-```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
+```{r subgenome_expression_protein_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=2.5, fig.align='center'}
+## Start with the paired expression data
+maize.walley.v4mapped.expression.data <- 
+  maize.walley.v4mapped.expression %>%
+  select(geneID, sample, FPKM_avg)
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "FPKM_maize1"
+names(data)[5] <- "FPKM_maize2"
+
+## Add paired abundance data
+data <-
+  data %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+names(data)[6] <- "dNSAF_maize1"
+names(data)[7] <- "dNSAF_maize2"
+
+# Pollen behaves much differently than the rest, lets see
+data <- 
+  data %>%
+  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode", "B73 Mature Pollen"))
+
+## View as scatter
+subgenome_expression_protein_scatter_rep <- ggplot(data, aes(FPKM_maize1, dNSAF_maize1)) +
+  geom_point(alpha=.1, aes(color = sample)) +
+  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
+  # geom_smooth(method="loess", se=TRUE, fullrange=FALSE, level=0.95) +
+  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
+  scale_x_log10() + scale_y_log10() +
+  theme(legend.position = "none") + facet_wrap(~sample, nrow=1) +
+  labs(
+    # title = paste0("Scatterplot of Expression in Maize1 vs. Abundance in Maize1\nLog-Log Scale for gene pair(s).\nAlpha=.1, reference line (black), lm fit (blue)"),
+    x = "Expression for Maize1 Gene (log scale)",
+    y = "Abundance for Maize1 Gene (log scale)"
+  ) 
+plot(subgenome_expression_protein_scatter_rep)
+```
+
+```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
 ## Start with the paired expression data
@@
 ## View as scatter
-ggplot(data, aes(FPKM_maize1, dNSAF_maize1)) +
+subgenome_expression_protein_scatter <- ggplot(data, aes(FPKM_maize1, dNSAF_maize1)) +
   geom_point(alpha=.1, aes(color = sample)) +
@@
   ) 
-
-cap <- "Fig 10.2: "
+plot(subgenome_expression_protein_scatter)
 ```
 
-# 5 Expression pattern types
+# 7 Expression pattern types
 functionalDivergence.R
 
-# 6 Wrapup
+# 8 Misc
+
 ***
+
+```{r extra1, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
+#Pollen has far fewer cases with exp
+## Start with the paired expression data
+maize.walley.v4mapped.expression.data <- 
+  maize.walley.v4mapped.expression %>%
+  select(geneID, sample, FPKM_avg)
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "FPKM_maize1"
+names(data)[5] <- "FPKM_maize2"
+
+case_counts <- data %>% group_by(sample) %>% count() %>% arrange(n)
+ggplot(case_counts, aes(reorder(sample, n), n)) + geom_col() + theme(axis.text.x=element_text(angle=90,hjust=1))
+```
+
+```{r extra2, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
+#Pollen has far fewer cases with abun
+## Start with the abundance data
+maize.walley.v4mapped.expression.data <- 
+  maize.walley.v4mapped.expression %>%
+  select(geneID, sample, FPKM_avg)
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+names(data)[4] <- "dNSAF_maize1"
+names(data)[5] <- "dNSAF_maize2"
+
+case_counts <- data %>% group_by(sample) %>% count() %>% arrange(n)
+ggplot(case_counts, aes(reorder(sample, n), n)) + geom_col() + theme(axis.text.x=element_text(angle=90,hjust=1))
+```
+
+```{r extra3, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
+#Pollen has far fewer cases with both exp and abun
+## Start with the paired expression data
+maize.walley.v4mapped.expression.data <- 
+  maize.walley.v4mapped.expression %>%
+  select(geneID, sample, FPKM_avg)
+data <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  distinct() %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
+  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
+names(data)[4] <- "FPKM_maize1"
+names(data)[5] <- "FPKM_maize2"
+
+## Add paired abundance data
+data <-
+  data %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
+  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
+names(data)[6] <- "dNSAF_maize1"
+names(data)[7] <- "dNSAF_maize2"
+
+case_counts <- data %>% group_by(sample) %>% count() %>% arrange(n)
+ggplot(case_counts, aes(reorder(sample, n), n)) + geom_col() + theme(axis.text.x=element_text(angle=90,hjust=1))
+```
+
+***
+
+# 9 Wrapup
+
+***
+
+```{r print, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
+ggsave("Output/Figure 1.png", plot = chromosome_plot)
+
+png("Output/Figure 2a.png");
+grid.draw(GOTerms_by_subgenome_a)
+dev.off()
+png("Output/Figure 2b.png");
+grid.draw(GOTerms_by_subgenome_b)
+dev.off()
+
+ggsave("Output/Figure 3.png", plot = subgenome_isoforms_scatter)
+ggsave("Output/Figure 4.png", plot = subgenome_expression_scatter_rep)
+ggsave("Output/Figure 5.png", plot = subgenome_protein_scatter_rep)
+ggsave("Output/Figure 6.png", plot = subgenome_expression_protein_foldchange_scatter)
+ggsave("Output/Figure 7.png", plot = subgenome_expression_protein_scatter)
+ggsave("Output/Figure 8.png", plot = subgenome_expression_protein_scatter_rep)
+ggsave("Output/Figure S1.png", plot = subgenome_expression_scatter)
+ggsave("Output/Figure S2.png", plot = subgenome_protein_scatter)
+ggsave("Output/Figure S3.png", plot = subgenome_expression_protein_scatter)
+```
+
 Report generated on: `r as.character(format(Sys.Date(), format="%B %d, %Y"))`

### 2018-06-07 Manuscript Figures.Rmd
Add protein dominance to chromosome plot
And also clean up code and move load/clean functions to proper scripts.  Remove Maggie subgenomes and Orphan Genes lists

@@
 
-```{r chromosome_plot, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6}
-v4_subg_jw <-
+```{r chromosome_plot_exp, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6}
+subgenome.positions <-
   subgenome %>%
-  left_join(positions, by=c("gene2"="geneID")) %>%
+  left_join(B73.v4.gene.positions, by=c("gene2"="geneID")) %>%
   select(Chromosome, Pos_start, Pos_end, subgenome) %>%
@@
 
-v4_dom_jw <- expressedPairs
-v4_dom_jw$Subgenome <- "no dominance"
-v4_dom_jw$Subgenome[v4_dom_jw$FPKM_mean1 / v4_dom_jw$FPKM_mean2 > 2] <- "dom"
-v4_dom_jw$Subgenome[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2] <- "dom"
-v4_dom_jw$Maize1[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2] <- v4_dom_jw$Maize2[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2]
-df0 <- v4_dom_jw[v4_dom_jw$Subgenome %in% c("no dominance"),]
+dominant_homeolog_exp <- expressedPairs
+dominant_homeolog_exp$Subgenome <- "no dominance"
+dominant_homeolog_exp$Subgenome[dominant_homeolog_exp$FPKM_mean1 / dominant_homeolog_exp$FPKM_mean2 > 2] <- "Dominant Expression"
+dominant_homeolog_exp$Subgenome[dominant_homeolog_exp$FPKM_mean2 / dominant_homeolog_exp$FPKM_mean1 > 2] <- "Dominant Expression"
+dominant_homeolog_exp$Maize1[dominant_homeolog_exp$FPKM_mean2 / dominant_homeolog_exp$FPKM_mean1 > 2] <- dominant_homeolog_exp$Maize2[dominant_homeolog_exp$FPKM_mean2 / dominant_homeolog_exp$FPKM_mean1 > 2]
+df0 <- dominant_homeolog_exp[dominant_homeolog_exp$Subgenome %in% c("no dominance"),]
 df0$Maize1 <- df0$Maize2
-v4_dom_jw <- bind_rows(v4_dom_jw, df0)
-v4_dom_jw <-
-  v4_dom_jw %>%
+dominant_homeolog_exp <- bind_rows(dominant_homeolog_exp, df0)
+dominant_homeolog_exp <-
+  dominant_homeolog_exp %>%
   select(Maize1, Subgenome) %>%
-  left_join(positions, by=c("Maize1"="geneID")) %>%
+  left_join(B73.v4.gene.positions, by=c("Maize1"="geneID")) %>%
   select(Chromosome, Pos_start, Pos_end, Subgenome)
 
-## 6,773 - 3,875 (no map) -~-  1,018 (no positions); split/merge is fine = 1,896 orphan v4_ids with positions based on Zea_mays.AGPv4.36.chr.gff3
-v4_orph_jw <-
-  OrphanGenes %>%
-  select(gene) %>%
-  left_join(maize.genes.v3_to_v4.map, c("gene" = "v3_id"))
-v4_orph_na <- v4_orph_jw[is.na(v4_orph_jw$v4_id),]
-v4_orph_jw <- v4_orph_jw[!is.na(v4_orph_jw$v4_id),]
-v4_orph_jw <-
-  v4_orph_jw %>%
-  select(v4_id) %>%
-  left_join(positions, by=c("v4_id"="geneID"))
-v4_orph_jw <- v4_orph_jw[!is.na(v4_orph_jw$Pos_end),]
-v4_orph_jw$Subgenome <- "orphan"
-v4_orph_jw <-
-  v4_orph_jw %>%
-  select(Chromosome, Pos_start, Pos_end, Subgenome)
-
-centromere.v4.positions <-
-  centromere.v4.positions %>%
-  rename(Chromosome="X__1", Pos_start="X__6", Pos_end="X__7") %>%
-  select(Chromosome, Pos_start, Pos_end)
-centromere.v4.positions
-
-df1 <- centromere.v4.positions
+df1 <- B73.v4.centromere.positions
 df1$x <- -.45
 df1$xend <- .45
-df2 <- v4_subg_jw
+df2 <- subgenome.positions
 df2$x <- -.45
 df2$xend <- -.05
-df3 <- v4_dom_jw
+df3 <- dominant_homeolog_exp
 df3$x <- .05
@@
 # segment_data$Subgenome[segment_data$Subgenome %in% c("dom")] <- "Dominant Homeolog"
-segment_data <- segment_data %>% subset(Subgenome %in% c("Maize1", "Maize2", "orphan", "sub1dom", "sub2dom", "dom"))
+segment_data <- segment_data %>% subset(Subgenome %in% c("Maize1", "Maize2", "Dominant Expression"))
 centromere_data <- df1
 
-#--------------------------------------------------------------------------------------------------#
 ## Make the plot
-chromosome_plot <-
-  ggplot(data=v4) +
-    geom_bar(aes(v4$chromosome, v4$size), stat="identity", fill="grey70") +
+chromosome_plot_exp <-
+  ggplot(data=B73.v4.chr.size) +
+  geom_bar(aes(B73.v4.chr.size$chromosome, B73.v4.chr.size$size), stat="identity", fill="grey70") +
   # geom_segment(data=segment_data, aes(x=Chromosome+x, xend=Chromosome+xend, y=Pos_start, yend=Pos_end, colour=Subgenome), size=.001) +
@@
   scale_y_continuous(labels=format_si())
-chromosome_plot  +
+chromosome_plot_exp  +
   labs(
@@
   )
+```
 
+```{r chromosome_plot_abundance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6}
+subgenome.positions <-
+  subgenome %>%
+  left_join(B73.v4.gene.positions, by=c("gene2"="geneID")) %>%
+  select(Chromosome, Pos_start, Pos_end, subgenome) %>%
+  rename(Subgenome=subgenome) %>%
+  subset(Subgenome %in% c("sub1", "sub2"))
+
+dominant_homeolog_abun <- expressedProteinsPairs
+dominant_homeolog_abun$Subgenome <- "no dominance"
+dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean1 / dominant_homeolog_abun$dNSAF_mean2 > 2] <- "Dominant Abundance"
+dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- "Dominant Abundance"
+dominant_homeolog_abun$Maize1[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- dominant_homeolog_abun$Maize2[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2]
+df0 <- dominant_homeolog_abun[dominant_homeolog_abun$Subgenome %in% c("no dominance"),]
+df0$Maize1 <- df0$Maize2
+dominant_homeolog_abun <- bind_rows(dominant_homeolog_abun, df0)
+dominant_homeolog_abun <-
+  dominant_homeolog_abun %>%
+  select(Maize1, Subgenome) %>%
+  left_join(B73.v4.gene.positions, by=c("Maize1"="geneID")) %>%
+  select(Chromosome, Pos_start, Pos_end, Subgenome)
+
+df1 <- B73.v4.centromere.positions
+df1$x <- -.45
+df1$xend <- .45
+df2 <- subgenome.positions
+df2$x <- -.45
+df2$xend <- -.05
+df3 <- dominant_homeolog_abun
+df3$x <- .05
+df3$xend <- .45
+segment_data <- bind_rows(df2,df3)
+segment_data <- segment_data[!is.na(segment_data$Pos_start),]
+segment_data <-
+  segment_data %>%
+  ungroup()
+segment_data$Subgenome[segment_data$Subgenome %in% c("sub1")] <- "Maize1"
+segment_data$Subgenome[segment_data$Subgenome %in% c("sub2")] <- "Maize2"
+# segment_data$Subgenome[segment_data$Subgenome %in% c("dom")] <- "Dominant Homeolog"
+segment_data <- segment_data %>% subset(Subgenome %in% c("Maize1", "Maize2", "Dominant Abundance"))
+centromere_data <- df1
+
+## Make the plot
+chromosome_plot_abundance <-
+  ggplot(data=B73.v4.chr.size) +
+  geom_bar(aes(B73.v4.chr.size$chromosome, B73.v4.chr.size$size), stat="identity", fill="grey70") +
+  # geom_segment(data=segment_data, aes(x=Chromosome+x, xend=Chromosome+xend, y=Pos_start, yend=Pos_end, colour=Subgenome), size=.001) +
+  geom_rect(data=segment_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end, color=Subgenome, fill=Subgenome)) +
+  geom_rect(data=centromere_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end)) +
+  scale_fill_manual(values=c("green", "red", "blue", "black")) +
+  scale_color_manual(values=c("green", "red", "blue", "black")) +
+  scale_x_discrete(limits=c("Chr1","Chr2","Chr3","Chr4","Chr5","Chr6","Chr7","Chr8","Chr9","Chr10")) +
+  scale_y_continuous(labels=format_si())
+chromosome_plot_abundance  +
+  labs(
+    # title="Subgenome Locations by Chromosome",
+    x="Chromosome",
+    y="Position on Chromosome"
+  )
 ```
@@
 
-# Remove pollen
-data <-
-  data %>%
-  subset(!sample %in% c("B73 Mature Pollen"))
+# # Remove pollen
+# data <-
+#   data %>%
+#   subset(!sample %in% c("B73 Mature Pollen"))
 
@@
 ```{r print, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
-ggsave("Output/Figure 1.png", plot = chromosome_plot)
-
-png("Output/Figure 2a.png");
-grid.draw(GOTerms_by_subgenome_a)
-dev.off()
-png("Output/Figure 2b.png");
-grid.draw(GOTerms_by_subgenome_b)
-dev.off()
-
-ggsave("Output/Figure 3.png", plot = subgenome_isoforms_scatter)
-ggsave("Output/Figure 4.png", plot = subgenome_expression_scatter_rep)
-ggsave("Output/Figure 5.png", plot = subgenome_protein_scatter_rep)
-ggsave("Output/Figure 6.png", plot = subgenome_expression_protein_foldchange_scatter)
-ggsave("Output/Figure 7.png", plot = subgenome_expression_protein_scatter)
-ggsave("Output/Figure 8.png", plot = subgenome_expression_protein_scatter_rep)
-ggsave("Output/Figure S1.png", plot = subgenome_expression_scatter)
-ggsave("Output/Figure S2.png", plot = subgenome_protein_scatter)
-ggsave("Output/Figure S3.png", plot = subgenome_expression_protein_scatter)
+# ggsave("Output/Figure 1a.png", plot = chromosome_plot_exp)
+# ggsave("Output/Figure 1b.png", plot = chromosome_plot_abundance)
+# 
+# png("Output/Figure 2a.png");
+# grid.draw(GOTerms_by_subgenome_a)
+# dev.off()
+# png("Output/Figure 2b.png");
+# grid.draw(GOTerms_by_subgenome_b)
+# dev.off()
+# 
+# ggsave("Output/Figure 3.png", plot = subgenome_isoforms_scatter)
+# ggsave("Output/Figure 4.png", plot = subgenome_expression_scatter_rep)
+# ggsave("Output/Figure 5.png", plot = subgenome_protein_scatter_rep)
+# ggsave("Output/Figure 6.png", plot = subgenome_expression_protein_foldchange_scatter)
+# ggsave("Output/Figure 7.png", plot = subgenome_expression_protein_scatter)
+# ggsave("Output/Figure 8.png", plot = subgenome_expression_protein_scatter_rep)
+# ggsave("Output/Figure S1.png", plot = subgenome_expression_scatter)
+# ggsave("Output/Figure S2.png", plot = subgenome_protein_scatter)
+# ggsave("Output/Figure S3.png", plot = subgenome_expression_protein_scatter)
 ```

### 2018-06-07 R/cleanData.R

@@ -10,6 +10,7 @@
 ##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.raw
 ##        go.sorghum.raw
 ##        gene.transcript.map
+##        B73.v4.centromere.positions.raw
 ##
 ## Output:
 ##        maize.walley.abundance.v4
@@ -17,6 +18,7 @@
 ##        syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean
 ##        go.sorghum.clean
 ##        geneTranscript.counts
+##        B73.v4.centromere.positions
 ##
 ## Date: 2017-08-25
 ## Author: Jesse R. Walsh
@@ -104,6 +106,20 @@ geneTranscript.counts <-
   select(gene) %>%
   group_by(gene) %>%
   summarise(n=n())
+
+#==================================================================================================#
+## B73.v4.centromere.positions.raw
+#--------------------------------------------------------------------------------------------------#
+B73.v4.centromere.positions <- B73.v4.centromere.positions.raw
+B73.v4.centromere.positions <-
+  B73.v4.centromere.positions %>%
+  rename(Chromosome="X__1", Pos_start="X__6", Pos_end="X__7") %>%
+  select(Chromosome, Pos_start, Pos_end)
+
+# Remove the extra region on chromosome 9, that is just confusing
+B73.v4.centromere.positions <- B73.v4.centromere.positions[-c(10),]
+
+B73.v4.centromere.positions$Chromosome <- as.integer(B73.v4.centromere.positions$Chromosome)
 
 #--------------------------------------------------------------------------------------------------#
 ## Clean up raw files

### 2018-06-07 R/do.R

@@ -84,6 +84,31 @@ expressedPairs <-
   rename(FPKM_mean2=FPKM_mean)
 
 #==================================================================================================#
+## parseAbundanceData.R
+#--------------------------------------------------------------------------------------------------#
+## Get means for each protein across all tissues and samples
+maize.walley.abundance.v4.replicate <-
+  maize.walley.abundance.v4 %>%
+  spread(key=sample, value=dNSAF_avg)
+expressedProteins <- data.frame(ID=maize.walley.abundance.v4.replicate[,1], Means=rowMeans(maize.walley.abundance.v4.replicate[,-1], na.rm = TRUE))
+
+## Remove rows with NA
+expressedProteins <-
+  expressedProteins %>%
+  rename(dNSAF_mean = Means) %>%
+  subset(!is.na(dNSAF_mean))
+
+## Attach log2(FPKM_mean) values to homeologous pairs
+expressedProteinsPairs <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>%
+  select(Maize1, Maize2) %>%
+  inner_join(expressedProteins, by=c("Maize1"="geneID")) %>%
+  rename(dNSAF_mean1=dNSAF_mean) %>%
+  inner_join(expressedProteins, by=c("Maize2"="geneID")) %>%
+  rename(dNSAF_mean2=dNSAF_mean)
+
+#==================================================================================================#
 ## Correlation between exp and abundance?
 #--------------------------------------------------------------------------------------------------#
 ## Start with the paired expression data

### 2018-06-07 R/loadData.R

@@ -79,11 +79,9 @@ subgenome.assignments <- read_delim("./data-raw/outfile_processed_byHand.tab", "
 chromosome.v4.size <- read.csv("./data-raw/ChrImages/Zea_mays.AGPv4.36.sizes.csv")
 gene.positions.v4.chromosome <- v4_subg <- read.csv("./data-raw/ChrImages/gene_positions.tab", sep = "\t") # this should really be in MaizeMap
 
-v4 <- read.csv("./data-raw/ChrImages/Zea_mays.AGPv4.36.sizes.csv")
-v4_subg <- read_csv("data-raw/ChrImages/v4_subg.csv")
-positions <- v4_subg <- read.csv("./data-raw/ChrImages/gene_positions.tab", sep = "\t")
-OrphanGenes <- read_excel("data-raw/ChrImages/OrphanGenes.xlsx")
-centromere.v4.positions <- read_excel("data-raw/ChrImages/centromere_positions.xlsx", col_names = FALSE)
+B73.v4.chr.size <- read.csv("./data-raw/ChrImages/Zea_mays.AGPv4.36.sizes.csv")
+B73.v4.gene.positions <- read_delim("~/git/SubGenomes/data-raw/ChrImages/gene_positions.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
+B73.v4.centromere.positions.raw <- read_excel("data-raw/ChrImages/centromere_positions.xlsx", col_names = FALSE)
 
 #--------------------------------------------------------------------------------------------------#
 detach("package:readr", unload=TRUE)

### 2018-06-07 Manuscript Figures.Rmd
Average FPKMs by subgenome favors maize2 often

@@
 
+```{r extra4, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
+# Average expression for genes in each subgenome, biased by sample size.  While other plots show that maize1 dominates maize2 more often in all tissue, maize2 
+# for pollen and endosperm crown the average expression is higher in maize2.  Basically, when maize2 wins, it wins by more raw FPKMs.  If we account for all
+# genes in the subgenome, not just the pairs, maize2 often has higher average expression.  This is probably due to many very low expressed genes in maize1.
+data <- 
+  homeologs.pairs %>%
+  # subset(Maize1 != "" & Maize2 != "") %>%
+  # select(Maize1, Maize2) %>%
+  distinct() %>%
+  select(Maize1, Maize2) %>%
+  left_join(maize.walley.v4mapped.expression, by=c("Maize1"="geneID")) %>%
+  rename(FPKM_avg1=FPKM_avg) %>%
+  left_join(maize.walley.v4mapped.expression, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  rename(FPKM_avg2=FPKM_avg) %>%
+  select(Maize1, Maize2, sample, FPKM_avg1, FPKM_avg2) %>%
+  # subset(sample %in% c("B73 Mature Pollen")) %>%
+  group_by(sample) %>%
+  summarize(Mean1=mean(FPKM_avg1, na.rm=TRUE), Mean2=mean(FPKM_avg2, na.rm=TRUE))
+data <- 
+  data %>%
+  gather(subgenome, FPKM, -1)
+
+ggplot(data, aes(x=sample, y=FPKM, fill=subgenome)) + geom_col(position = position_dodge()) + theme(axis.text.x=element_text(angle=90,hjust=1))
+```
+
 ***

### 2018-06-07 Manuscript Figures.Rmd
Why does maize1 average lower FPKM than maize2

@@
   homeologs.pairs %>%
-  # subset(Maize1 != "" & Maize2 != "") %>%
-  # select(Maize1, Maize2) %>%
+  # subset(Maize1 != "" & Maize2 != "") %>% #only matching
+  # subset(!(Maize1 %in% c("") & Maize2 %in% c(""))) %>% #only nonmatching
   distinct() %>%
@@
 ggplot(data, aes(x=sample, y=FPKM, fill=subgenome)) + geom_col(position = position_dodge()) + theme(axis.text.x=element_text(angle=90,hjust=1))
+```
+
+```{r extra5, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
+## In maize1, there are 3672 matching genes and 7927 nonmatching.  Since they tend towards 1 (despite cutoff at 1), more will pull average down.
+data <-
+  homeologs.pairs %>%
+  subset(Maize2 %in% c("")) %>% #only nonmatching
+  select(Maize1) %>%
+  left_join(maize.walley.v4mapped.expression, by=c("Maize1"="geneID")) %>%
+  select(Maize1, sample, FPKM_avg)
+temp <-
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>% #only matching
+  select(Maize1) %>%
+  left_join(maize.walley.v4mapped.expression, by=c("Maize1"="geneID")) %>%
+  select(Maize1, sample, FPKM_avg)
+
+data$status <- "nonMatch"
+temp$status <- "match"
+data <-
+  data %>%
+  bind_rows(temp)
+
+ggplot(data, aes(status,log10(FPKM_avg))) +
+  geom_boxplot() +
+  labs(y = "log10(FPKM)",
+       x = "Subgenome",
+       title = "Comparison of FPKM Expression between\nGenes in Subgenome 1 and Subgenome 2"
+  )
 ```

### 2018-06-08 Manuscript Figures.Rmd
Methods to sort pattern types in expression and abundance

@@
 
+```{r extra6, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=4, fig.align='center'}
+## Not really a plot yet... just sorting types
+source("./R/analysis/functionalDivergence_func.R")
+"gene"
+gene.pair.types <- getTypeSort(homeologs.pairs, maize.walley.v4mapped.expression, 2, .95)
+summary(as.factor(gene.pair.types$type))
+"protein"
+protein.pair.types <- getTypeSortProteins(homeologs.pairs, maize.walley.abundance.v4, 2, .95)
+summary(as.factor(protein.pair.types$type))
+"agree"
+summary(as.factor(intersect(gene.pair.types, protein.pair.types)$type))
+
+# intersect(gene.pair.types[!gene.pair.types$type %in% c("ambiguous"),], protein.pair.types[!protein.pair.types$type %in% c("ambiguous"),])
+# #dead expression and still dead abundance (17/297)
+# protein.pair.types[protein.pair.types$Maize1 %in% gene.pair.types$Maize1[gene.pair.types$type %in% c("1dead","2dead")],] %>% subset(type %in% c("1dead","2dead"))
+```
+
 ***

### 2018-06-08 R/analysis/functionalDivergence.R

@@ -23,11 +23,18 @@ library(dplyr)
 library(ggplot2)
 source("./R/scripts/functionalDivergence_func.R")
 
-maize.expression.sample.avg
-expressedPairs
-retained.duplicates <-
-  homeologs.pairs %>%
-  subset(Maize1 != "" & Maize2 != "")
+# retained.duplicates <-
+#   homeologs.pairs %>%
+#   subset(Maize1 != "" & Maize2 != "")
+#
+# data <-
+#   maize.walley.v4mapped.expression %>%
+#   select(geneID, sample, FPKM_avg) %>%
+#   subset(geneID %in% retained.duplicates$Maize1 | geneID %in% retained.duplicates$Maize2)
+#
+# data$FPKM_avg[is.na(data$FPKM_avg)] <- 0
+
+# expressedPairs
 
 which(retained.duplicates=="Zm00001d033174", arr.ind = TRUE)
 which(retained.duplicates=="Zm00001d034368", arr.ind = TRUE)
@@ -35,14 +42,33 @@ which(retained.duplicates=="Zm00001d034429", arr.ind = TRUE)
 which(retained.duplicates=="Zm00001d034517", arr.ind = TRUE)
 index <- 120
 test.cutoff <- 2
-plotRetainedDuplicatesExpression(retained.duplicates, maize.expression.sample.avg, index)
-plotRetainedDuplicatesAbundance(retained.duplicates, maize.protein.abundance.sample.avg, index)
+plotRetainedDuplicatesExpression(retained.duplicates, data, index)
+plotRetainedDuplicatesAbundance(retained.duplicates, maize.walley.abundance.v4, index)
 index
 index <- index + 1
 
-getType3 <- function() {
+gene.pair.types <- getTypeSort(homeologs.pairs, maize.walley.v4mapped.expression, 2, .95)
+summary(as.factor(getTypeSort(homeologs.pairs, maize.walley.v4mapped.expression, 2, .95)$type))
+protein.pair.types <- getTypeSortProteins(homeologs.pairs, maize.walley.abundance.v4, 2, .95)
+summary(as.factor(getTypeSortProteins(homeologs.pairs, maize.walley.abundance.v4, 2, .95)$type))
+intersect(gene.pair.types, protein.pair.types)
+intersect(gene.pair.types[!gene.pair.types$type %in% c("ambiguous"),], protein.pair.types[!protein.pair.types$type %in% c("ambiguous"),])
[…]

### 2018-06-08 R/analysis/functionalDivergence_func.R

@@ -1,3 +1,193 @@
+#==================================================================================================#
+## getTypeSort
+#--------------------------------------------------------------------------------------------------#
+getTypeSort <- function(homeologs.pairs, maize.walley.v4mapped.expression, foldCutoff, correlationCutoff) {
+  ## Build dataframes necessary for sorting types
+  # Retained duplicates are gene pairs where one is in maize1 and the other in maize2
+  retained.duplicates <-
+    homeologs.pairs %>%
+    subset(Maize1 != "" & Maize2 != "") %>%
+    select(Maize1, Maize2) %>%
+    distinct()
+
+  # Expression data for all genes in retained duplicates list as an avg of biological replicates
+  data <-
+    maize.walley.v4mapped.expression %>%
+    select(geneID, sample, FPKM_avg) %>%
+    subset(geneID %in% retained.duplicates$Maize1 | geneID %in% retained.duplicates$Maize2)
+
+  # NA's mess with some of the aggregating later on, make them zeros
+  data$FPKM_avg[is.na(data$FPKM_avg)] <- 0
+
+  # A dead gene is a gene with less than 1 FPKM in all tissues. Find these first so the "zero" FPKMs don't mess with foldchanges calculations later on.
+  dead.genes <-
+    data %>%
+    group_by(geneID) %>%
+    summarise(max=max(FPKM_avg)) %>%
+    ungroup() %>%
+    subset(max < 1) %>%
+    select(geneID)
+
+  retained.duplicates.2dead <- retained.duplicates[retained.duplicates$Maize1 %in% dead.genes$geneID & retained.duplicates$Maize2 %in% dead.genes$geneID,]
+  retained.duplicates <- setdiff(retained.duplicates, retained.duplicates.2dead)
+  retained.duplicates.1dead <- retained.duplicates[retained.duplicates$Maize1 %in% dead.genes$geneID | retained.duplicates$Maize2 %in% dead.genes$geneID,]
+  retained.duplicates <- setdiff(retained.duplicates, retained.duplicates.1dead)
+
+  # Find genes with reciprocal dominance. (BED genes in the Pophaly paper)
+  bed.genes <-
+    retained.duplicates %>%
+    inner_join(data, by=c("Maize1"="geneID")) %>%
[…]

### 2018-06-08 R/staging_testing.R
Where are the dead genes on the chromosomes

@@ -133,56 +133,60 @@ data$m2[data$Value_maize1 < data$Value_maize2] <- 1
 data.temp <- data %>% select(sample, m1, m2) %>% group_by(sample) %>% summarise(vm1=sum(m1), vm2=sum(m2)) %>% ungroup()
 data.temp$per <- data.temp$vm1/(data.temp$vm1 + data.temp$vm2)
 #--------------------------------------------------------------------------------------------------#
-plot <- graphGenePairDominanceByExperiment(maize.walley.v4mapped.expression %>% select(geneID, sample, FPKM_avg), homeologs.pairs, 4, FALSE)
-plot(plot)
+subgenome.positions <-
+  subgenome %>%
+  left_join(B73.v4.gene.positions, by=c("gene2"="geneID")) %>%
+  select(Chromosome, Pos_start, Pos_end, subgenome) %>%
+  rename(Subgenome=subgenome) %>%
+  subset(Subgenome %in% c("sub1", "sub2"))
 
-data <-
-  homeologs.pairs %>%
-  # subset(Maize1 != "" & Maize2 != "") %>% #only matching
-  subset(!(Maize1 %in% c("") & Maize2 %in% c(""))) %>% #only nonmatching
-  distinct() %>%
-  select(Maize1, Maize2) %>%
-  left_join(maize.walley.v4mapped.expression, by=c("Maize1"="geneID")) %>%
-  rename(FPKM_avg1=FPKM_avg) %>%
-  left_join(maize.walley.v4mapped.expression, by=c("Maize2"="geneID", "sample"="sample")) %>%
-  rename(FPKM_avg2=FPKM_avg) %>%
-  select(Maize1, Maize2, sample, FPKM_avg1, FPKM_avg2) %>%
-  # subset(sample %in% c("B73 Mature Pollen")) %>%
-  group_by(sample) %>%
-  summarize(Mean1=mean(FPKM_avg1, na.rm=TRUE), Mean2=mean(FPKM_avg2, na.rm=TRUE))
-data <-
-  data %>%
-  gather(subgenome, FPKM, -1)
+gene.pair.types <- getTypeSort(homeologs.pairs, maize.walley.v4mapped.expression, 2, .95)
+# dominant_homeolog_exp <- data.frame(geneID=append(gene.pair.types$Maize1[gene.pair.types$type %in% c("1dead")], gene.pair.types$Maize2[gene.pair.types$type %in% c("1dead")]))
[…]

### 2018-06-08 Manuscript Figures.Rmd
Higher quality exports of figures

@@
 
-```{r extra4, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
+```{r extra4, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
 # Average expression for genes in each subgenome, biased by sample size.  While other plots show that maize1 dominates maize2 more often in all tissue, maize2 
@@
   homeologs.pairs %>%
-  # subset(Maize1 != "" & Maize2 != "") %>% #only matching
+  subset(Maize1 != "" & Maize2 != "") %>% #only matching
   # subset(!(Maize1 %in% c("") & Maize2 %in% c(""))) %>% #only nonmatching
@@
 ```{r print, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
-# ggsave("Output/Figure 1a.png", plot = chromosome_plot_exp)
-# ggsave("Output/Figure 1b.png", plot = chromosome_plot_abundance)
+# # ggsave("Output/Figure 1a.png", plot = chromosome_plot_exp)
+# png(filename = "Output/Figure 1a.png", type = "cairo", units = "in", width = 8, height = 8, pointsize = 12, res = 96)
+# plot(chromosome_plot_exp)
+# dev.off()
+# 
+# # ggsave("Output/Figure 1b.png", plot = chromosome_plot_abundance)
+# png(filename = "Output/Figure 1b.png", type = "cairo", units = "in", width = 8, height = 8, pointsize = 12, res = 96)
+# plot(chromosome_plot_abundance)
+# dev.off()
 # 
@@
 # 
-# ggsave("Output/Figure 3.png", plot = subgenome_isoforms_scatter)
-# ggsave("Output/Figure 4.png", plot = subgenome_expression_scatter_rep)
-# ggsave("Output/Figure 5.png", plot = subgenome_protein_scatter_rep)
-# ggsave("Output/Figure 6.png", plot = subgenome_expression_protein_foldchange_scatter)
-# ggsave("Output/Figure 7.png", plot = subgenome_expression_protein_scatter)
-# ggsave("Output/Figure 8.png", plot = subgenome_expression_protein_scatter_rep)
+# # ggsave("Output/Figure 3.png", plot = subgenome_isoforms_scatter)
+# png(filename = "Output/Figure 3.png", type = "cairo", units = "in", width = 8, height = 8, pointsize = 12, res = 96)
+# plot(subgenome_isoforms_scatter)
+# dev.off()
+# 
+# # ggsave("Output/Figure 4.png", plot = subgenome_expression_scatter_rep)
+# png(filename = "Output/Figure 4.png", type = "cairo", units = "in", width = 12, height = 4, pointsize = 12, res = 96)
+# plot(subgenome_expression_scatter_rep)
+# dev.off()
+# 
+# # ggsave("Output/Figure 5.png", plot = subgenome_protein_scatter_rep)
+# png(filename = "Output/Figure 5.png", type = "cairo", units = "in", width = 12, height = 4, pointsize = 12, res = 96)
+# plot(subgenome_protein_scatter_rep)
+# dev.off()
+# 
+# # ggsave("Output/Figure 6.png", plot = subgenome_expression_protein_scatter_rep)
+# png(filename = "Output/Figure 6.png", type = "cairo", units = "in", width = 12, height = 4, pointsize = 12, res = 96)
+# plot(subgenome_expression_protein_scatter_rep)
+# dev.off()
+# 
 # ggsave("Output/Figure S1.png", plot = subgenome_expression_scatter)
@@
 # ggsave("Output/Figure S3.png", plot = subgenome_expression_protein_scatter)
+# ggsave("Output/Figure S4.png", plot = subgenome_expression_protein_foldchange_scatter)
 ```

### 2018-06-14 Manuscript Figures.Rmd
Includes basic text output for several counts used in manuscript

@@
 source("./R/do.R")
+source("./R/pathway_analysis.R")
 
@@
 ```
+
+Some direct numbers for reporting in various parts of the manuscript.
+
+***
+
+Overlap in reactions and pathways annotated to retained pairs.
+
+```{r numbers1, echo=FALSE, results='asis'}
+knitr::kable(getPathwayReactionOverlap(homeologs.pairs) %>% select(Description, Value))
+```
+
+How often does the maize1 gene express higher than maize2?  Only count when both genes have expression.
+
+```{r numbers2, echo=FALSE, results='asis'}
+data <- 
+  homeologs.pairs %>%
+  subset(Maize1 != "" & Maize2 != "") %>% #only matching
+  select(Maize1, Maize2) %>%
+  inner_join(maize.walley.v4mapped.expression, by=c("Maize1"="geneID")) %>%
+  rename(FPKM_avg1=FPKM_avg) %>%
+  inner_join(maize.walley.v4mapped.expression, by=c("Maize2"="geneID", "sample"="sample")) %>%
+  rename(FPKM_avg2=FPKM_avg) %>%
+  select(Maize1, Maize2, sample, FPKM_avg1, FPKM_avg2) %>%
+  count(sample, sign=sign(FPKM_avg1-FPKM_avg2)) %>%
+  spread(key = sign, value = n) %>%
+  mutate(percent_favor_m1=`1`/(`-1`+`1`)) %>%
+  rename(m2_dom=`-1`, m1_dom=`1`) %>%
+  select(sample, m1_dom, m2_dom, percent_favor_m1) %>%
+  arrange(percent_favor_m1)
+knitr::kable(data)
+```
+
+ * There are **`r nrow(homeologs.pairs %>% subset(Maize1 != "" & Maize2 != ""))`** retained duplicate pairs.
+ * 
+ <!-- * Summary of the KS values -->
+ <!--     + **`r 1+2`** -->
 
@@
 
-```{r chromosome_plot_abundance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6}
-subgenome.positions <-
-  subgenome %>%
-  left_join(B73.v4.gene.positions, by=c("gene2"="geneID")) %>%
-  select(Chromosome, Pos_start, Pos_end, subgenome) %>%
-  rename(Subgenome=subgenome) %>%
-  subset(Subgenome %in% c("sub1", "sub2"))
-
-dominant_homeolog_abun <- expressedProteinsPairs
-dominant_homeolog_abun$Subgenome <- "no dominance"
-dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean1 / dominant_homeolog_abun$dNSAF_mean2 > 2] <- "Dominant Abundance"
-dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- "Dominant Abundance"
-dominant_homeolog_abun$Maize1[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- dominant_homeolog_abun$Maize2[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2]
-df0 <- dominant_homeolog_abun[dominant_homeolog_abun$Subgenome %in% c("no dominance"),]
-df0$Maize1 <- df0$Maize2
-dominant_homeolog_abun <- bind_rows(dominant_homeolog_abun, df0)
-dominant_homeolog_abun <-
-  dominant_homeolog_abun %>%
-  select(Maize1, Subgenome) %>%
-  left_join(B73.v4.gene.positions, by=c("Maize1"="geneID")) %>%
-  select(Chromosome, Pos_start, Pos_end, Subgenome)
-
-df1 <- B73.v4.centromere.positions
-df1$x <- -.45
-df1$xend <- .45
-df2 <- subgenome.positions
-df2$x <- -.45
-df2$xend <- -.05
-df3 <- dominant_homeolog_abun
-df3$x <- .05
-df3$xend <- .45
-segment_data <- bind_rows(df2,df3)
-segment_data <- segment_data[!is.na(segment_data$Pos_start),]
-segment_data <-
-  segment_data %>%
-  ungroup()
-segment_data$Subgenome[segment_data$Subgenome %in% c("sub1")] <- "Maize1"
-segment_data$Subgenome[segment_data$Subgenome %in% c("sub2")] <- "Maize2"
-# segment_data$Subgenome[segment_data$Subgenome %in% c("dom")] <- "Dominant Homeolog"
-segment_data <- segment_data %>% subset(Subgenome %in% c("Maize1", "Maize2", "Dominant Abundance"))
-centromere_data <- df1
-
-## Make the plot
-chromosome_plot_abundance <-
-  ggplot(data=B73.v4.chr.size) +
-  geom_bar(aes(B73.v4.chr.size$chromosome, B73.v4.chr.size$size), stat="identity", fill="grey70") +
-  # geom_segment(data=segment_data, aes(x=Chromosome+x, xend=Chromosome+xend, y=Pos_start, yend=Pos_end, colour=Subgenome), size=.001) +
-  geom_rect(data=segment_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end, color=Subgenome, fill=Subgenome)) +
-  geom_rect(data=centromere_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end)) +
-  scale_fill_manual(values=c("green", "red", "blue", "black")) +
-  scale_color_manual(values=c("green", "red", "blue", "black")) +
-  scale_x_discrete(limits=c("Chr1","Chr2","Chr3","Chr4","Chr5","Chr6","Chr7","Chr8","Chr9","Chr10")) +
-  scale_y_continuous(labels=format_si())
-chromosome_plot_abundance  +
-  labs(
-    # title="Subgenome Locations by Chromosome",
-    x="Chromosome",
-    y="Position on Chromosome"
-  )
-```
+<!-- # ```{r chromosome_plot_abundance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6} -->
+<!-- # subgenome.positions <- -->
+<!-- #   subgenome %>% -->
+<!-- #   left_join(B73.v4.gene.positions, by=c("gene2"="geneID")) %>% -->
+<!-- #   select(Chromosome, Pos_start, Pos_end, subgenome) %>% -->
+<!-- #   rename(Subgenome=subgenome) %>% -->
+<!-- #   subset(Subgenome %in% c("sub1", "sub2")) -->
+<!-- #  -->
+<!-- # dominant_homeolog_abun <- expressedProteinsPairs -->
+<!-- # dominant_homeolog_abun$Subgenome <- "no dominance" -->
+<!-- # dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean1 / dominant_homeolog_abun$dNSAF_mean2 > 2] <- "Dominant Abundance" -->
+<!-- # dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- "Dominant Abundance" -->
+<!-- # dominant_homeolog_abun$Maize1[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- dominant_homeolog_abun$Maize2[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] -->
+<!-- # df0 <- dominant_homeolog_abun[dominant_homeolog_abun$Subgenome %in% c("no dominance"),] -->
+<!-- # df0$Maize1 <- df0$Maize2 -->
+<!-- # dominant_homeolog_abun <- bind_rows(dominant_homeolog_abun, df0) -->
+<!-- # dominant_homeolog_abun <- -->
+<!-- #   dominant_homeolog_abun %>% -->
+<!-- #   select(Maize1, Subgenome) %>% -->
+<!-- #   left_join(B73.v4.gene.positions, by=c("Maize1"="geneID")) %>% -->
+<!-- #   select(Chromosome, Pos_start, Pos_end, Subgenome) -->
+<!-- #  -->
+<!-- # df1 <- B73.v4.centromere.positions -->
+<!-- # df1$x <- -.45 -->
+<!-- # df1$xend <- .45 -->
+<!-- # df2 <- subgenome.positions -->
+<!-- # df2$x <- -.45 -->
+<!-- # df2$xend <- -.05 -->
+<!-- # df3 <- dominant_homeolog_abun -->
+<!-- # df3$x <- .05 -->
+<!-- # df3$xend <- .45 -->
+<!-- # segment_data <- bind_rows(df2,df3) -->
+<!-- # segment_data <- segment_data[!is.na(segment_data$Pos_start),] -->
+<!-- # segment_data <- -->
+<!-- #   segment_data %>% -->
+<!-- #   ungroup() -->
+<!-- # segment_data$Subgenome[segment_data$Subgenome %in% c("sub1")] <- "Maize1" -->
+<!-- # segment_data$Subgenome[segment_data$Subgenome %in% c("sub2")] <- "Maize2" -->
+<!-- # # segment_data$Subgenome[segment_data$Subgenome %in% c("dom")] <- "Dominant Homeolog" -->
+<!-- # segment_data <- segment_data %>% subset(Subgenome %in% c("Maize1", "Maize2", "Dominant Abundance")) -->
+<!-- # centromere_data <- df1 -->
+<!-- #  -->
+<!-- # ## Make the plot -->
+<!-- # chromosome_plot_abundance <- -->
+<!-- #   ggplot(data=B73.v4.chr.size) + -->
+<!-- #   geom_bar(aes(B73.v4.chr.size$chromosome, B73.v4.chr.size$size), stat="identity", fill="grey70") + -->
+<!-- #   # geom_segment(data=segment_data, aes(x=Chromosome+x, xend=Chromosome+xend, y=Pos_start, yend=Pos_end, colour=Subgenome), size=.001) + -->
+<!-- #   geom_rect(data=segment_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end, color=Subgenome, fill=Subgenome)) + -->
+<!-- #   geom_rect(data=centromere_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end)) + -->
+<!-- #   scale_fill_manual(values=c("green", "red", "blue", "black")) + -->
+<!-- #   scale_color_manual(values=c("green", "red", "blue", "black")) + -->
+<!-- #   scale_x_discrete(limits=c("Chr1","Chr2","Chr3","Chr4","Chr5","Chr6","Chr7","Chr8","Chr9","Chr10")) + -->
+<!-- #   scale_y_continuous(labels=format_si()) -->
+<!-- # chromosome_plot_abundance  + -->
+<!-- #   labs( -->
+<!-- #     # title="Subgenome Locations by Chromosome", -->
+<!-- #     x="Chromosome", -->
+<!-- #     y="Position on Chromosome" -->
+<!-- #   ) -->
+<!-- # ``` -->
 
@@
 
-```{r extra6, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=4, fig.align='center'}
+```{r extra6, echo=FALSE, results='asis', warning=FALSE, message=FALSE, fig.width=4, fig.height=4, fig.align='center'}
 ## Not really a plot yet... just sorting types
@@
 
+data <- 
+  data.frame(type=c("gene","protein","agree"),
+             values=bind_rows(summary(as.factor(gene.pair.types$type)), 
+                              summary(as.factor(protein.pair.types$type)),
+                              summary(as.factor(intersect(gene.pair.types, protein.pair.types)$type))
+                              )
+             ) %>% rename(one_dead=values.1dead, both_dead=values.2dead, ambiguous=values.ambiguous, bed=values.bed, true=values.true)
+
+knitr::kable(data)
 # intersect(gene.pair.types[!gene.pair.types$type %in% c("ambiguous"),], protein.pair.types[!protein.pair.types$type %in% c("ambiguous"),])
@@
 # ggsave("Output/Figure S4.png", plot = subgenome_expression_protein_foldchange_scatter)
+# 
+# list.of.subgenome.assignments <- subgenome %>% select(gene2, subgenome) %>% rename(B73.v4.ID=gene2, Subgenome=subgenome)
+# write.csv(list.of.subgenome.assignments, "list.of.subgenome.assignments.csv")
+# 
+# write.csv(homeologs.pairs %>% subset(Maize1 != "" & Maize2 != "") %>% select(Maize1, Maize2) %>% rename(B73.v4.ID.Maize1=Maize1, B73.v4.ID.Maize2=Maize2), "list.of.retained.duplicate.pairs")
 ```

### 2018-07-08 R/analysis/functionalDivergence_func.R
Comment fixed

@@ -105,7 +105,7 @@ getTypeSortProteins <- function(homeologs.pairs, maize.walley.abundance.v4, fold
     select(Maize1, Maize2) %>%
     distinct()
 
-  # Expression data for all genes in retained duplicates list as an avg of biological replicates
+  # Abundance data for all proteins in retained duplicates list as an avg of biological replicates
   data <-
     maize.walley.abundance.v4 %>%
     select(geneID, sample, dNSAF_avg) %>%

### 2018-07-08 Manuscript Figures.Rmd
Updates to graph

@@
 
-```{r chromosome_plot_exp, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6}
+```{r chromosome_plot_exp, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=5}
 subgenome.positions <-
@@
 df1 <- B73.v4.centromere.positions
-df1$x <- -.45
-df1$xend <- .45
+df1$x <- -.5
+df1$xend <- .05
 df2 <- subgenome.positions
@@
 df3$x <- .05
-df3$xend <- .45
+df3$xend <- .25
 segment_data <- bind_rows(df2,df3)
@@
     y="Position on Chromosome"
-  )
+  ) + coord_flip()
 ```
-
-<!-- # ```{r chromosome_plot_abundance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6} -->
-<!-- # subgenome.positions <- -->
-<!-- #   subgenome %>% -->
-<!-- #   left_join(B73.v4.gene.positions, by=c("gene2"="geneID")) %>% -->
-<!-- #   select(Chromosome, Pos_start, Pos_end, subgenome) %>% -->
-<!-- #   rename(Subgenome=subgenome) %>% -->
-<!-- #   subset(Subgenome %in% c("sub1", "sub2")) -->
-<!-- #  -->
-<!-- # dominant_homeolog_abun <- expressedProteinsPairs -->
-<!-- # dominant_homeolog_abun$Subgenome <- "no dominance" -->
-<!-- # dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean1 / dominant_homeolog_abun$dNSAF_mean2 > 2] <- "Dominant Abundance" -->
-<!-- # dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- "Dominant Abundance" -->
-<!-- # dominant_homeolog_abun$Maize1[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- dominant_homeolog_abun$Maize2[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] -->
-<!-- # df0 <- dominant_homeolog_abun[dominant_homeolog_abun$Subgenome %in% c("no dominance"),] -->
-<!-- # df0$Maize1 <- df0$Maize2 -->
-<!-- # dominant_homeolog_abun <- bind_rows(dominant_homeolog_abun, df0) -->
-<!-- # dominant_homeolog_abun <- -->
-<!-- #   dominant_homeolog_abun %>% -->
-<!-- #   select(Maize1, Subgenome) %>% -->
-<!-- #   left_join(B73.v4.gene.positions, by=c("Maize1"="geneID")) %>% -->
-<!-- #   select(Chromosome, Pos_start, Pos_end, Subgenome) -->
-<!-- #  -->
-<!-- # df1 <- B73.v4.centromere.positions -->
-<!-- # df1$x <- -.45 -->
-<!-- # df1$xend <- .45 -->
-<!-- # df2 <- subgenome.positions -->
-<!-- # df2$x <- -.45 -->
-<!-- # df2$xend <- -.05 -->
-<!-- # df3 <- dominant_homeolog_abun -->
-<!-- # df3$x <- .05 -->
-<!-- # df3$xend <- .45 -->
-<!-- # segment_data <- bind_rows(df2,df3) -->
-<!-- # segment_data <- segment_data[!is.na(segment_data$Pos_start),] -->
-<!-- # segment_data <- -->
-<!-- #   segment_data %>% -->
-<!-- #   ungroup() -->
-<!-- # segment_data$Subgenome[segment_data$Subgenome %in% c("sub1")] <- "Maize1" -->
-<!-- # segment_data$Subgenome[segment_data$Subgenome %in% c("sub2")] <- "Maize2" -->
-<!-- # # segment_data$Subgenome[segment_data$Subgenome %in% c("dom")] <- "Dominant Homeolog" -->
-<!-- # segment_data <- segment_data %>% subset(Subgenome %in% c("Maize1", "Maize2", "Dominant Abundance")) -->
-<!-- # centromere_data <- df1 -->
-<!-- #  -->
-<!-- # ## Make the plot -->
-<!-- # chromosome_plot_abundance <- -->
-<!-- #   ggplot(data=B73.v4.chr.size) + -->
-<!-- #   geom_bar(aes(B73.v4.chr.size$chromosome, B73.v4.chr.size$size), stat="identity", fill="grey70") + -->
-<!-- #   # geom_segment(data=segment_data, aes(x=Chromosome+x, xend=Chromosome+xend, y=Pos_start, yend=Pos_end, colour=Subgenome), size=.001) + -->
-<!-- #   geom_rect(data=segment_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end, color=Subgenome, fill=Subgenome)) + -->
-<!-- #   geom_rect(data=centromere_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end)) + -->
-<!-- #   scale_fill_manual(values=c("green", "red", "blue", "black")) + -->
-<!-- #   scale_color_manual(values=c("green", "red", "blue", "black")) + -->
-<!-- #   scale_x_discrete(limits=c("Chr1","Chr2","Chr3","Chr4","Chr5","Chr6","Chr7","Chr8","Chr9","Chr10")) + -->
-<!-- #   scale_y_continuous(labels=format_si()) -->
-<!-- # chromosome_plot_abundance  + -->
-<!-- #   labs( -->
-<!-- #     # title="Subgenome Locations by Chromosome", -->
-<!-- #     x="Chromosome", -->
-<!-- #     y="Position on Chromosome" -->
-<!-- #   ) -->
-<!-- # ``` -->
 
@@
 ```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
+# #EXP only, MF terms only
+# grid.newpage()
+# GOTerms_by_subgenome_a <- draw.pairwise.venn(
+#   area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
+#   area2 = goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
+#   cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct(),  
+#                          goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct()
+#                          ) %>% nrow(), 
+#   category = c("Maize1", "Maize2"), 
+#   lty = rep("blank", 2), 
+#   fill = c("light blue", "green"), 
+#   alpha = rep(0.5, 2), 
+#   cat.pos = c(0, 180), 
+#   scaled = FALSE,
+#   euler.d = FALSE, 
+#   sep.dist = 0.3, 
+#   rotation.degree = 0,
+#   ind = FALSE
+# )
+# grid.arrange(gTree(children=GOTerms_by_subgenome_a), top="GO Term by Subgenome\n(EV-EXP, MF only)")
+# 
+# ## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2.  The converted data.frame will hold maize2 genes with the names replaced with maize1 homologs
+# converted <-
+#   homeologs.pairs %>%
+#   subset(Maize1 != "" & Maize2 != "") %>%
+#   right_join(., goAnnotations.sub2,by=c("Maize2"="gene2"))
+# converted$Maize2[!is.na(converted$Maize1)] <- as.character(converted$Maize1[!is.na(converted$Maize1)])
+# converted <- 
+#   converted %>%
+#   select(Maize2, goTerm, evCode, type) %>%
+#   rename(gene2=Maize2)
+# 
+# #EXP only, MF only
+# grid.newpage()
+# GOTerms_by_subgenome_b <- draw.pairwise.venn(
+#   area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% nrow(),
+#   area2 = converted %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% nrow(),
+#   cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(gene2, goTerm) %>% distinct(), 
+#                          converted %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(gene2, goTerm) %>% distinct()
+#                          ) %>% nrow(), 
+#   category = c("Maize1", "Maize2"), 
+#   lty = rep("blank", 2), 
+#   fill = c("light blue", "green"), 
+#   alpha = rep(0.5, 2), 
+#   cat.pos = c(0, 180), 
+#   scaled = FALSE,
+#   euler.d = FALSE, 
+#   sep.dist = 0.3, 
+#   rotation.degree = 0,
+#   ind = FALSE
+# )
+# grid.arrange(gTree(children=GOTerms_by_subgenome_b), top="GO Term Assignments by Subgenome\n(EV-EXP, MF only)")
+
 #EXP only, MF terms only
@@
 grid.arrange(gTree(children=GOTerms_by_subgenome_b), top="GO Term Assignments by Subgenome\n(EV-EXP, MF only)")
-
 ```
